### 2024-01-20 bio_dynamics/bio274_hw1.Rmd
bio274 hw1 q1 through 3

---
title: "BIOL 274 Homework 1"
author: "Ruby Krasnow"
date: "2024-01-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1
>Assume that once a drug is administered to a patient, it leaves the body following exponential decay. After administration of a dose, the concentration of the drug in the body decreases by 50% in 30 hours. Approximately how long does it take for the drug to decrease to 1% of its initial value (round to the nearest hour)?

We can represent the concentration of drug in the body at time $t$ as $x(t)=x_0e^{kt}$. Since the drug leaves the body according to an exponential decay model, that means $k$ should be negative.

We are given $\frac{x}{x_0}=0.5$ at $t=30$ hours. Plugging into our equation, we get
$$0.5=e^{30k}$$
$$\ln(0.5)=\ln(e^{30k})$$
$$\ln(0.5)=30k$$
$$k=\frac{\ln(0.5)}{30}$$
```{r solve k}
k = log(0.5)/30
k
```

<!-- So $k$ = `r k`. -->
Now, we want to know how long it will take for $x$ to be 1% of $x_0$, or at what time $t$ we will have $\frac{x(t)}{x(0)}=0.01$. 
$$0.01=e^{kt}$$
$$t=\frac{\ln{(0.01)}}{k}$$

```{r solve t}
t = log(0.01)/k
round(t, 0)
```

It will take approximately 199 hours for the drug to decrease to 1% of its initial value.

## Question 2
> Consider the population model: $\frac{dP}{dt}=0.3 \left( 1-\frac{P}{200} \right) \left(\frac{P}{50}-1 \right)P$.

>a. For what values of P is the population increasing?
b. For what values of P is the population decreasing?

If the population is increasing, that means the rate of change $\frac{dP}{dt}$ is positive, while a decreasing population means that $\frac{dP}{dt}$ is negative. First we need to find the equilibrium points, or the values of $P$ for which $\frac{dP}{dt}=0.3 \left( 1-\frac{P}{200} \right) \left(\frac{P}{50}-1 \right)P=0$.

We can see that this will clearly occur if $P=0$; i.e., the population has no organisms in it. The other equilibrium points occur when
$$1-\frac{P}{200}=0$$
$$1=\frac{P}{200}$$
$$P=200$$

and similarly when $1=\frac{P}{50}$ or $P=50$.
So we need to determine the sign of $\frac{dP}{dt}$ on the intervals $(0,50),(50,200),$ and $(200,\infty)$.
Let's evaluate some convenient values within these intervals: 25, 100, and 400.

```{r}
eqn <- function(P){
  0.3*(1-(P/200))*((P/50-1))*P
} 
eqn(25)
eqn(100)
eqn(400)
```

So the population is decreasing when $0<P<50$ or $P>200$ and the population is increasing when $50<P<200$. Let's confirm this with a quick plot:
```{r}
x=c(1:250)

ggplot()+
  geom_line(aes(x=x, y=0.3*(1-(x/200))*((x/50-1))*x))+
  geom_vline(xintercept = c(50, 200), linetype="dashed")+
  geom_hline(yintercept = 0, linetype="dashed", color="red")+
  theme_bw()+
  labs(x="P", y="dP/dt")
```

## Question 3
Find the general solution to each of the following DEs:

### a
$$ \frac{dy}{dt}=\frac{3t+1}{2y}$$
$$ 2y \, dy=(3t+1)dt$$
$$\int 2y \, dy=\int(3t+1)dt$$
$$y^2=\frac{3t^2}{2}+t+C$$
$$y=\sqrt{\frac{3t^2}{2}+t+C}$$

### b
$$\frac{dy}{dx}=x\sqrt[3]{y}$$

$$y^{-\frac{1}{3}} \, dy=x \, dx$$
$$\int y^{-\frac{1}{3}} \, dy= \int x \, dx$$
$$\frac{3}{2}y^{\frac{2}{3}}=\frac{x^2}{2}+C_1$$
$$y^{\frac{2}{3}}=\frac{x^2+C_2}{3}$$
$$y=\left(\frac{x^2+C_2}{3} \right)^\frac{3}{2}=\frac{(x^2+C_2)^{\frac{3}{2}}}{3\sqrt3}$$

### c
$$\frac{dp}{dt}=\frac{1}{3tp^2}$$
$$3p^2\,dp=\frac{1}{t}dt$$
$$\int 3p^2\,dp=\int \frac{1}{t}dt$$
$$p^3=\ln(t)+C$$
$$p=\sqrt[3]{\ln(t)+C}$$