### 2024-01-26 bio_dynamics/bio274_filing_cabinet.Rmd
started hw2

```{r}
x=c(100:5000)

fish1 <- tibble(x=x,
                orig=0.3*(1-(x/2500)),
                constant=0.3*(1-(x/2500))-100,
                third=0.3*(1-(x/2500))-x/3,
                prop=0.3*(1-(x/2500))-0.5*sqrt(x))
fish1 <- fish1 %>% pivot_longer(cols=2:5, names_to = "eqn", values_to = "dP_dt")

ggplot()+
 # geom_line(data=fish1, aes(x=x, y=dP_dt, color=eqn))+
  geom_line(data=fish1 %>% filter(eqn=="orig"), aes(x=x, y=dP_dt, color=eqn))+
  theme_bw()+
  labs(x="P", y="dP/dt", color="type")
```