### 2024-01-27 bio_dynamics/bio274_hw2.Rmd
problem 2 hw2

---
title: "BIOL 274 Homework 2"
author: "Ruby Krasnow"
date: "2024-01-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1
Suppose a species of fish in a particular lake has a population that is modeled by the Logistic
Growth model with $r = 0.3, K = 2500$, and $P_0 = 2500$. Adjust the model and write down an
IVP to represent each of the following situations.

>a. One hundred fish are harvested each year.
>b. One-third of the fish population is harvested annually.
>c. The number of fish harvested each year is proportional to the square root of the number
of fish in the lake.

We will assume that, in the absence of harvesting, the fish population at time t, $p(t)$, can be modeled by a discrete logistic growth equation with a time step of $\Delta t=1$ year. Given an initial population size of $P(0)=P_0 = 2500$, we have the following IVP:

\begin{equation}
    \begin{cases}
      \frac{dP}{dt}=0.3P \left( 1- \frac{P}{2500} \right) \\
      P(0)=2500
    \end{cases}       
\end{equation}

Although since we are working in discrete time, it might be more accurate to write the change in population as a difference equation, $P_{n+1}-P_n=\Delta P= 0.3P \left( 1- \frac{P}{2500} \right)$. For now we will continue using differential equation notation and apologize to Leibniz.

The initial conditions will remain the same for each of the three situations, but we can adjust the differential equation for each scenario as follows:

> **a.** One hundred fish are harvested each year:

$$\frac{dP}{dt}=0.3P \left( 1- \frac{P}{2500} \right)-100$$

> **b.** One-third of the fish population is harvested annually:

$$\frac{dP}{dt}=0.3P \left( 1- \frac{P}{2500} \right)-\frac{P}{3}$$

> **c.** The number of fish harvested each year is proportional to the square root of the number
of fish in the lake:

$$\frac{dP}{dt}=0.3P \left( 1- \frac{P}{2500} \right)-\alpha \sqrt{P}$$
Where $\alpha$ is a constant of proportionality.

Now we're going to have some fun and see how the population would change over a 100-year period if no harvesting occurred and under each of these three harvesting strategies! We will use the numerical ODE solver from the `deSolve` package.

```{r}

library(deSolve) # best package for solving ODEs in R

#First we define our state variables
state_fish <- c(Po = 2500, #original - no harvest
                Pc = 2500, #constant harvest (100 fish/year)
                P3 = 2500, #1/3 of the fish are harvested every year
                Ps = 2500) #fish are harvested prop. to the square root of P

#How many time steps (years) do we want the model to run for?
times_fish <- seq(0, 100, 1)

#Provide values for our parameters. This makes it easy to identify 
#how the population might change if we changed the carrying capacity 
#or had a different harvesting intensity in scenario C.
params_fish <- c(K=2500, #carrying capacity
                 a=0.5) #constant that scales the harvest intensity in scenario C
```

```{r}
#Define the model structure and the differential equations
fish_model <- function(t, state, parameters) {

  with(as.list(c(state, parameters)), { 
    dPo_dt <- 0.3*Po*(1-(Po/K)) #original - no harvest
    dPc_dt <- 0.3*Pc*(1-(Pc/K))-100 #constant harvest (100 fish/year)
    dP3_dt <- 0.3*P3*(1-(P3/K))-P3/3 #1/3 of the fish are harvested every year
    dPs_dt <- 0.3*Ps*(1-(Ps/K))-a*sqrt(Ps) #fish are harvested prop. to the square root of P

  #output
    return(list(c(dPo_dt, dPc_dt, dP3_dt, dPs_dt)))
  })
}

#Run the model
ode_output_fish <- ode(y = state_fish, t = times_fish, func = fish_model, parms = params_fish)

#Convert model output to a better format for visualization
fish_output <- ode_output_fish %>% 
  as.data.frame() %>% 
  pivot_longer(cols=2:5, names_to = "eqn", values_to = "P") %>% 
  mutate(names_full = case_when(
    eqn =="Po" ~ "No harvesting",
    eqn =="Pc" ~ "100 fish",
    eqn =="P3" ~ "1/3 of P",
    eqn =="Ps"~"0.5*sqrt(P)")
  )
```

Let's see what we got!
```{r}
ggplot()+
  geom_line(data=fish_output, aes(x=time, y=P, color=names_full))+
  theme_bw()+
  labs(x="Time (years)", y="Pop. size", color=NULL)
```

```{r}
#How many fish are left after 100 years?
fish_output %>% slice_tail(n=4)
```

Cool! So in the original formulation, the population doesn't change because our initial conditions had P equal to our carrying capacity, so the $\left( 1- \frac{P}{K} \right)$ term was equal to zero and therefore $\frac{dP}{dt}=0$.

Harvesting 1/3 of the population each year is not a sustainable management strategy, because we have decreased the population to almost zero.

Harvest scenario 3 actually did pretty well, and the impact on the population would be even smaller if we changed the values of $\alpha$. If I was doing a whole project on these equations, it would be fun to test different values of $K$, $P_0$, and $\alpha$ and test the impacts on the population, dive into how the bifurcation points and stability of the solutions might change, and explore how we could use the model to identify an optimal harvest strategy.

## Question 2
Using Separation of Variables, find a solution to the IVP representing the situation in #1b.

Recall that the IVP for the situation in #1b is:
\begin{equation}\label{IVP}
    \begin{cases}
      \frac{dP}{dt}=0.3P \left( 1- \frac{2500}{P} \right)-\frac{P}{3} \\
      P(0)=2500
    \end{cases}       
\end{equation}

$$\frac{dP}{dt}=\frac{3P}{10} \left( 1- \frac{P}{2500} \right)-\frac{P}{3}$$
$$dP\, \frac{1}{\frac{3P}{10} \left( 1- \frac{P}{2500} \right)-\frac{P}{3}}=dt$$
$$dP\, \frac{1}{\frac{3P}{10} - \frac{3P^2}{25000}-\frac{P}{3}}=dt$$
$$dP\, \frac{1}{P \left( \frac{3}{10} - \frac{3P}{25000}- \frac{1}{3} \right)} = dt $$
$$dP\, \frac{1}{P \left( \frac{9}{30} - \frac{3P}{25000}- \frac{10}{30} \right)} = dt $$
$$dP\, \frac{1}{P \left(- \frac{3P}{25000}- \frac{1}{30} \right)} = dt $$

Now we use partial fraction decomposition to rewrite the fraction, assuming there exist some $A, B$ such that
$$\frac{1}{P \left(- \frac{3P}{25000}- \frac{1}{30} \right)}=
\frac{A}{- \frac{3P}{25000}- \frac{1}{30}} + \frac{B}{P} $$

$$AP+B\left(- \frac{3P}{25000}- \frac{1}{30} \right)=1 $$
$$AP-\frac{3BP}{25000}-\frac{B}{30}=1 $$
$-\frac{B}{30}=1$, so $B=-30$.
$$P\left( A-\frac{3B}{25000} \right) = P\left( A+\frac{9}{2500} \right)=0 $$
$A=-\frac{9}{2500}$, so now
$$\frac{A}{- \frac{3P}{25000}- \frac{1}{30}} + \frac{B}{P}=$$
$$\frac{-9}{- \frac{3P}{10}- \frac{250}{3}} + \frac{-30}{P}$$
$$ \left( \frac{9}{ \frac{3P}{10}+ \frac{250}{3}} - \frac{30}{P} \right) dP= dt$$
$$ \left( \frac{9}{ \frac{9P+2500}{30}} - \frac{30}{P} \right) dP= dt$$
$$270 \int \frac{1}{ 9P+2500} dP-30\int  \frac{1}{P}\,dP= \int dt$$
$$30\ln{(9P+2500)}-30\ln(P)= t+C_1$$
$$\ln{(9P+2500)}-\ln(P)= \frac{t+C_1}{30}$$
$$\ln{\left(\frac{9P+2500}{P}\right)}= \frac{t+C_1}{30}$$
$$\frac{9P+2500}{P}= e^{\frac{t}{30}+\frac{C_1}{30}}=C_2e^{\frac{t}{30}}$$

$$9P+2500=PC_2e^{\frac{t}{30}}$$
$$9P-PC_2e^{\frac{t}{30}}=-2500$$
$$P\left (9-C_2e^{\frac{t}{30}}\right)=-2500$$

$$P=\frac{2500}{C_2e^{\frac{t}{30}}-9}$$

Now using our initial conditions,
$$2500=\frac{2500}{C_2-9}$$
$C_2-9=1$, so $C_2=10$

Our final solution is therefore:
\begin{equation}\label{solution}P(t)=\frac{2500}{10e^{\frac{t}{30}}-9}\end{equation}
Here is what it looks like:
```{r}
t=c(1:250)
y=2500/(10*exp(t/30)-9)

ggplot()+
  geom_line(aes(x=t, y=y, color="P(t)"))+
  theme_bw()+
  labs(x="time", y="P")
```

Let's check and make sure it is actually a solution and satisfies our initial conditions:

We want to show that the derivative with respect to $t$ of (\ref{solution}) is equivalent to $\frac{3P}{10} \left( 1- \frac{P}{2500} \right)-\frac{P}{3}$.

Using the chain rule, we differentiate (\ref{solution}) to get:
$$P'=2500(-1) \left( 10e^{\frac{t}{30}}-9 \right)^{-2} \left( 10e^{\frac{t}{30}} \right) \left(\frac{1}{30}\right)=\frac{-2500e^{\frac{t}{30}}}{3 \left( 10e^{\frac{t}{30}}-9 \right)^{2}}$$
Since there are no $t$'s in the original $\frac{dP}{dt}$, we can use our solution to rewrite them as a function of $P$:
$$P\left( 10e^{\frac{t}{30}}-9\right)=2500$$
$$10e^{\frac{t}{30}}=\frac{2500}{P}+9$$
$$e^{\frac{t}{30}}=\frac{250}{P}+\frac{9}{10}$$
Now substituting this into our equation for $P'$, we get
$$\frac{-2500 \left( \frac{250}{P}+\frac{9}{10} \right)}
{3 \left( 10\left( \frac{250}{P}+\frac{9}{10} \right)-9 \right)^{2}} = 
\frac{-2500 \left( \frac{250}{P}+\frac{9}{10} \right)}
{3 \left( \frac{2500}{P}+9-9 \right)^{2}} = 
\frac{-2500 \left( \frac{250}{P}+\frac{9}{10} \right)}
{3 \left( \frac{2500^2}{P^2} \right)} = 
\frac{ \frac{-250}{P}-\frac{9}{10}}
{3 \left( \frac{2500}{P^2} \right)}$$
Now multiply the numerator and denominator by $P$:

$$\frac{-250-\frac{9P}{10}}
{\frac{7500}{P}} = \left(-250-\frac{9P}{10}\right)\left( \frac{P}{7500}\right)=
\frac{-250P}{7500}-\frac{9P^2}{75000}=P \left(- \frac{3P}{25000}- \frac{1}{30} \right)$$

And see that our initial $\frac{dP}{dt}$,
$$\frac{3P}{10} \left( 1- \frac{P}{2500} \right)-\frac{P}{3} = 
\frac{3P}{10} - \frac{3P^2}{2500}-\frac{P}{3} =
P \left(- \frac{3P}{25000}- \frac{1}{30} \right)$$

So indeed, our solution satisfies the differential equation.

And checking the initial conditions:
$$P(0)=\frac{2500}{10e^{0}-9}=\frac{2500}{1}=2500$$
We can also confirm by using R to calculate the derivative of our solution and plotting it alongside the original expression for $\frac{dP}{dt}$:

```{r}
dy=D(expression(2500/(10*exp(t/30)-9)), "t")

ggplot()+
  geom_line(aes(x=y, y=eval(dy), color="calculated"), linewidth=3, alpha=0.5)+
  geom_line(aes(x=y, y= 0.3*y*(1-(y/2500))-y/3, color="original dP/dt"))+
  scale_color_manual(values=c("lightblue", "black"))+
  theme_bw()+
  labs(x="P", y="dP/dt")
```

## Question 2
Using Bernoulli's method, find a solution to the IVP representing the situation in #1b.

$$\frac{dP}{dt} = \frac{3P}{10} - \frac{3P^2}{25000}-\frac{1P}{30}$$

Bernoulli's method review:
If the differential equation is in the form $y'(x)+p(x)y=g(x)y^n$,
we can divide by $y^n$ and then use the substitution $z=y^{1-n}$.

Here,
$$P'(t)+\frac{1P}{30}=-\frac{3P^2}{25000}$$
Divide by $P^2$:
$$P'(P^{-2})+\frac{1}{30P}=-\frac{3}{25000}$$
Use the substitution $z=P^{-1}$, $z'=-P^{-2}P'$:
$$-z'+\frac{z}{30}=-\frac{3}{25000}$$
$$z'-\frac{z}{30}=\frac{3}{25000}$$
$$\mu=e^{\int \frac{-1}{30} dt}=e^{\frac{-t}{30}}$$

$$\int \left( e^{\frac{-t}{30}}z \right)'dt=\int \frac{3e^{\frac{-t}{30}}}{25000} dt $$

$$ e^{\frac{-t}{30}}z=\frac{-9e^{\frac{-t}{30}}}{2500}+C_1$$

$$z=\frac{-9}{2500}+\frac{C_1}{e^{\frac{-t}{30}}}$$

$$P=1/z= \frac{1}{\frac{-9}{2500}+C_1e^{\frac{t}{30}}}$$

$$P(0)= 2500=\frac{1}{\frac{-9}{2500}+C_1}$$

$$2500 \left( \frac{-9}{2500}+C_1 \right)=1$$

$$-9+2500 C_1=1$$

Then $2500 C_1=10$, so $C_1=\frac{1}{250}$

$$P(t)= \frac{1} {\frac{-9}{2500}+\left( \frac{1}{250}\right) e^{\frac{t}{30}}}=
\frac{2500} {10e^{\frac{t}{30}}-9}$$

So we did get the same answer as when we used separation of variables.