### 2024-01-29 PDEs/math245_hw1.Rmd
PDE hw1

---
title: "MATH 245 Homework 1"
author: "Ruby Krasnow"
date: "2024-01-28"
output:
  pdf_document:
    extra_dependencies: ["derivative"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1
Show the function is a solution of the PDE:  
(a) $u_{xx}+u_{yy}=0$   
(i) $u(x,y)=e^x\sin(y)$  
$$u_x=e^x\sin(y), \, u_{xx}=e^x\sin(y), \, u_y=e^x\cos(y),\, u_{yy}=-e^x\sin(y)$$
Which means that $u_{xx}+u_{yy}=e^x\sin(y)+\left(-e^x\sin(y)\right)=0$ and so $u(x,y)=e^x\sin(y)$ is
a solution to the PDE.  
(ii) $u(x,y)=\log{\sqrt{x^2+y^2}}$  
Assuming this is meant to be $u(x,y)=\ln\sqrt{{{x^2+y^2}}}$,
$$u_x=\left({{x^2+y^2}}\right)^{-1/2}
\left(\frac{1}{2}\right)\left({{x^2+y^2}}\right)^{-1/2}
\left(2x\right)=x\left({{x^2+y^2}}\right)^{-1}$$

$$u_{xx}=x(-1)\left({{x^2+y^2}}\right)^{-2}\left(2x\right)+\left({{x^2+y^2}}\right)^{-1}$$

$$u_y=\left({{x^2+y^2}}\right)^{-1/2}
\left(\frac{1}{2}\right)\left({{x^2+y^2}}\right)^{-1/2}
\left(2y\right)=y\left({{x^2+y^2}}\right)^{-1}$$

$$u_{yy}=y(-1)\left({{x^2+y^2}}\right)^{-2}\left(2y\right)+\left({{x^2+y^2}}\right)^{-1}$$

$$u_{xx}+u_{yy}=\frac{-2x^2}{\left({{x^2+y^2}}\right)^{2}}+
\frac{-2y^2}{\left({{x^2+y^2}}\right)^{2}}+
\frac{2}{\left({{x^2+y^2}}\right)}=$$

$$\frac{-2\left({{x^2+y^2}}\right)}{\left({{x^2+y^2}}\right)^{2}}+ \frac{2}{\left({{x^2+y^2}}\right)}=0$$
Which means that $u(x,y)=\log{\sqrt{x^2+y^2}}$ is a solution to the PDE.

(b) $bu_x+au_y+u=0, \, u(x,y)=\exp{\left(\frac{-x}{b}\right)}f(ax-by)$ for arbitrary differentiable function $f$.
$$u_x=e^{\frac{-x}{b}} \left(\frac{-1}{b}\right)f(ax-by)+f'(ax-by)(a)e^{\frac{-x}{b}}$$
$$u_y=e^{\frac{-x}{b}}f'(ax-by)(-b)$$
Then,
$$bu_x+au_y+u=-e^{\frac{-x}{b}}f(ax-by)+(ab)\left(e^{\frac{-x}{b}}\right)f'(ax-by)+
e^{\frac{-x}{b}}f'(ax-by)(-ab)+e^{\frac{-x}{b}}f(ax-by)=0$$

So $u(x,y)=\exp{\left(\frac{-x}{b}\right)}f(ax-by)$ is a solution to the PDE for an arbitrary differentiable function $f$.

(c) $u_{xx}-\frac{1}{x}u_x-x^2u_{yy}=0, \, u(x,y)=f\left(2y+x^2\right)+g\left(2y-x^2\right)$ for arbitrary twice-differentiable functions $f$ and $g$.

$$u_x=\left(f'\left(2y+x^2\right)\right)(2x)+\left(g'\left(2y-x^2\right)\right)(-2x)$$
$$u_y=2 \left(f'\left(2y+x^2\right)\right) +2\left(g'\left(2y-x^2\right)\right)$$
$$u_{xx}=2f'+(f'')(4x^2)-2g'+(g'')(4x^2)$$
$$u_{yy}=4f''+4g''$$
$$u_{xx}-\frac{1}{x}u_x-x^2u_{yy}=$$ $$2f'+4x^2f''-2g'+4x^2g''-\frac{1}{x}\left(2xf'-2xg'\right)-x^2\left(4f''+4g''\right)=$$
$$2f'-2f'+4x^2f''-4x^2f''-2g'+2g'+4x^2g''-4x^2g''=0 $$

So $u(x,y)=f\left(2y+x^2\right)+g\left(2y-x^2\right)$ is a solution to the PDE for arbitrary twice-differentiable functions $f$ and $g$.

### Question 2
(a) 2nd-order linear homogenous  
(b) 4th-order linear inhomogenous  
(c) 2nd-order quasi-linear homogeneous  
(d) We can rewrite this as $u^{-5}u_{xx}+u^{-5}u_{yy}+u^{-4}f(x,y)=g(x,y)$ to see that it is a 2nd-order quasi-linear inhomogeneous PDE.

### Question 3
Use separation of variables to solve the following problems:  
#### (a)
$u_x+u=u_y, \, u(x,0)=4x^{-3x}$, use $u(x,y)=f(x)g(y)$

#### (b)
$x^2u_{xy}+9y^2u=0, \, u(x,0)=\exp{\left(\frac{1}{x}\right)}$, use $u(x,y)=f(x)g(y)$

#### (c)
$u_x^2+u_y^2=1$, use $u(x,y)=f(x)+g(y)$

#### (d)
$x^2u_x^2+y^2u_y^2=u^2$, use $u(x,y)=e^{f(x)}e^{g(y)}$

### Question 4
For each of the following IVPs, (i) find and plot the characteristic lines (curves), (ii) solve the IVP, and (iii) plot the solution of (a)-(b) for indicated time.  
#### (a)
$u_t+\left(1+x^2\right)u_x=0, \, u(0,x)=\arctan{(x)}, \, t=1,2,3$, and
what is $\lim\limits_{x \to \infty}u(t,x)$?

$$\frac{dx}{dt}=1+x^2$$
$$\frac{1}{1+x^2}dx=dt$$

$$\arctan{(x)}=t+C$$
So $x=\tan{(t+C)}$ are the characteristic curves of $u_t+\left(1+x^2\right)u_x=0$.  
On each of the curves, $u(x,t)$ is constant because $\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x} \sec^2{(t+c)}$.  
And since $1+\tan^2\theta=\sec^2\theta$, this means $\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x}\left( 1+\tan^2{(t+c)}\right)=u_t+\left(1+x^2\right)u_x$, which we know is 0.

Thus $u(t, \tan{(t+C)})=u(0, \tan{(0+C)})=u(0, C)$ is independent of $t$. Putting $x=\tan{(t+C)}$ and $C=\arctan{(x)}-t$, we have
$$u(t, x)=u(0,\arctan{(x)}-t)$$
It follows that $u(t,x)=f(\arctan{(x)}-t)$. And since we are given that $u(0,x)=\arctan{(x)}$, we have $f(\arctan{(x)}-0)=\arctan{(x)}$ so that $f(w)=w$ for any $w$, yielding our solution of 
$$u(t,x)=\arctan{(x)}-t$$

```{r}
C=1
df_4a <- tibble(x = seq(-2*pi, 2*pi, length.out = 1000),
                y=tan(x+C)) %>% 
  mutate(y = ifelse(abs(y) < 20, y, NA))

ggplot(df_4a, aes(x=x, y=y)) + 
  geom_line()+
  coord_cartesian(ylim=c(-10,10), xlim = c(-5,5))+
  theme_bw()+
  labs(x="t", y="x")+
  ggtitle("Characteristic curves for Problem 4a", subtitle = "with C=1")
```

Since the range of $\arctan$ is limited to $\left(\frac{-\pi}{2},\frac{\pi}{2}\right)$, as $\lim\limits_{t \to \infty}$ that means $u(t,x)=\arctan(x)-t$ will go to $-\infty$.

```{r}

x=seq(-10,10,length.out = 1000)
ggplot() + 
  geom_line(aes(x=x, y=atan(x)-1, color="1"))+
  geom_line(aes(x=x, y=atan(x)-2, color="2"))+
  geom_line(aes(x=x, y=atan(x)-3, color="3"))+
  theme_bw()+
  labs(x="x", y="u(x,t)", color="Value of t")+
  ggtitle("Solution curves for Problem 4a")
```