### 2024-01-30 PDEs/math245_hw1.Rmd
problem 3, colored text boxes

---
title: "MATH 245 Homework 1"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-01-30"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1
Show the function is a solution of the PDE:  
(a) $u_{xx}+u_{yy}=0$   
(i) $u(x,y)=e^x\sin(y)$  
$$u_x=e^x\sin(y), \, u_{xx}=e^x\sin(y), \, u_y=e^x\cos(y),\, u_{yy}=-e^x\sin(y)$$
Which means that $u_{xx}+u_{yy}=e^x\sin(y)+\left(-e^x\sin(y)\right)=0$ and so $u(x,y)=e^x\sin(y)$ is
a solution to the PDE.  
(ii) $u(x,y)=\log{\sqrt{x^2+y^2}}$  
Assuming this is meant to be $u(x,y)=\ln\sqrt{{{x^2+y^2}}}$,
$$u_x=\left({{x^2+y^2}}\right)^{-1/2}
\left(\frac{1}{2}\right)\left({{x^2+y^2}}\right)^{-1/2}
\left(2x\right)=x\left({{x^2+y^2}}\right)^{-1}$$

$$u_{xx}=x(-1)\left({{x^2+y^2}}\right)^{-2}\left(2x\right)+\left({{x^2+y^2}}\right)^{-1}$$

$$u_y=\left({{x^2+y^2}}\right)^{-1/2}
\left(\frac{1}{2}\right)\left({{x^2+y^2}}\right)^{-1/2}
\left(2y\right)=y\left({{x^2+y^2}}\right)^{-1}$$

$$u_{yy}=y(-1)\left({{x^2+y^2}}\right)^{-2}\left(2y\right)+\left({{x^2+y^2}}\right)^{-1}$$

$$u_{xx}+u_{yy}=\frac{-2x^2}{\left({{x^2+y^2}}\right)^{2}}+
\frac{-2y^2}{\left({{x^2+y^2}}\right)^{2}}+
\frac{2}{\left({{x^2+y^2}}\right)}=$$

$$\frac{-2\left({{x^2+y^2}}\right)}{\left({{x^2+y^2}}\right)^{2}}+ \frac{2}{\left({{x^2+y^2}}\right)}=0$$
Which means that $u(x,y)=\log{\sqrt{x^2+y^2}}$ is a solution to the PDE.

(b) $bu_x+au_y+u=0, \, u(x,y)=\exp{\left(\frac{-x}{b}\right)}f(ax-by)$ for arbitrary differentiable function $f$.
$$u_x=e^{\frac{-x}{b}} \left(\frac{-1}{b}\right)f(ax-by)+f'(ax-by)(a)e^{\frac{-x}{b}}$$
$$u_y=e^{\frac{-x}{b}}f'(ax-by)(-b)$$
Then,
$$bu_x+au_y+u=-e^{\frac{-x}{b}}f(ax-by)+(ab)\left(e^{\frac{-x}{b}}\right)f'(ax-by)+
e^{\frac{-x}{b}}f'(ax-by)(-ab)+e^{\frac{-x}{b}}f(ax-by)=0$$

So $u(x,y)=\exp{\left(\frac{-x}{b}\right)}f(ax-by)$ is a solution to the PDE for an arbitrary differentiable function $f$.

(c) $u_{xx}-\frac{1}{x}u_x-x^2u_{yy}=0, \, u(x,y)=f\left(2y+x^2\right)+g\left(2y-x^2\right)$ for arbitrary twice-differentiable functions $f$ and $g$.

$$u_x=\left(f'\left(2y+x^2\right)\right)(2x)+\left(g'\left(2y-x^2\right)\right)(-2x)$$
$$u_y=2 \left(f'\left(2y+x^2\right)\right) +2\left(g'\left(2y-x^2\right)\right)$$
$$u_{xx}=2f'+(f'')(4x^2)-2g'+(g'')(4x^2)$$
$$u_{yy}=4f''+4g''$$
$$u_{xx}-\frac{1}{x}u_x-x^2u_{yy}=$$ $$2f'+4x^2f''-2g'+4x^2g''-\frac{1}{x}\left(2xf'-2xg'\right)-x^2\left(4f''+4g''\right)=$$
$$2f'-2f'+4x^2f''-4x^2f''-2g'+2g'+4x^2g''-4x^2g''=0 $$

So $u(x,y)=f\left(2y+x^2\right)+g\left(2y-x^2\right)$ is a solution to the PDE for arbitrary twice-differentiable functions $f$ and $g$.

### Question 2
(a) 2nd-order linear homogeneous  
(b) 4th-order linear inhomogeneous  
(c) 2nd-order quasi-linear homogeneous  
(d) We can rewrite this as $u_{xx}+u_{yy}+f(x,y)u-g(x,y)u^5=0$, making it clear that this is a 2nd-order semi-linear homogeneous PDE.

### Question 3
Use separation of variables to solve the following problems:

#### (a)
$u_x+u=u_y, \, u(x,0)=4x^{-3x}$, use $u(x,y)=f(x)g(y)$
$$u_x=f'g, \, u_y=fg'$$
$$u_x+u-u_y=f'g+fg-fg'=f'g+(g-g')f=0$$
$$f'g=-(g-g')f$$
$$\frac{f'}{f}=\frac{g'-g}{g}=\lambda$$
$$\frac{f'}{f}=\lambda, \,\frac{g'}{g}-1=\lambda $$
Integrating the first ODE with respect to $x$,
$$\ln(f)=\lambda x+C_1, \, f(x)=C_2e^{\lambda x}$$
$$\frac{g'}{g}=\lambda+1 $$
$$\ln(g)=\lambda y+y+C_3,\, g(y)=C_4e^{y(\lambda +1)}$$
$$u(x,y)=f(x)g(y)=C_2e^{\lambda x}C_4e^{y(\lambda +1)}$$
$$u(x,y)=Ce^{\lambda x+y(\lambda +1)}$$
$$u(x,0)=4e^{-3x}=Ce^{\lambda x}$$

Which means $C=4, \lambda=-3$ and our final solution is
\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=4e^{-3x-2y}$$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$u_x+u=-12e^{-3x-2y}+4e^{-3x-2y}=-8e^{-3x-2y}=u_y$$
\end{tcolorbox}

#### (b)
$x^2u_{xy}+9y^2u=0, \, u(x,0)=\exp{\left(\frac{1}{x}\right)}$, use $u(x,y)=f(x)g(y)$

$$u_{xy}=f'g'$$
$$x^2f'g'+9y^2fg=0$$

$$x^2f'g'=-9y^2fg$$
$$x^2\frac{f'}{f}=-9y^2\frac{g}{g'}=\lambda$$
$$x^2\frac{f'}{f}=\lambda, \, -9y^2\frac{g}{g'}=\lambda$$
$$x^2\frac{f'}{f}=\lambda \quad \longrightarrow \quad 
\frac{f'}{f}=\lambda x^{-2} \quad \longrightarrow \quad
\ln(f)=\frac{-\lambda}{x}+C_1 \quad \longrightarrow \quad 
f(x)=C_2e^{\frac{-\lambda}{x}}$$

$$-9y^2\frac{g}{g'}=\lambda \quad \longrightarrow \quad
\frac{g}{g'}= \frac{\lambda}{-9y^2} \quad \longrightarrow \quad 
\frac{g'}{g}= \frac{-9y^2}{\lambda} \quad \longrightarrow \quad 
\ln(g)=\frac{-3y^3}{\lambda}+C_3 \quad \longrightarrow \quad 
g(y)=C_4e^{\frac{-3}{\lambda}y^3}$$

$$u(x,y)=f(x)g(y)=C_2e^{\frac{-\lambda}{x}}C_4e^{\frac{-3}{\lambda}y^3}=Ce^{\frac{-\lambda}{x}+\frac{-3}{\lambda}y^3}$$
$$u(x,0)=e^{\frac{1}{x}}=Ce^{\frac{-\lambda}{x}}$$
Which means $C=1, \lambda=-1$, and our final solution is
\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=e^{\frac{1}{x}+3y^3}$$
\end{tcolorbox}
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$x^2u_{xy}+9y^2u=x^2e^{\frac{1}{x}+3y^3}\left(-x^{-2} \right) \left(9y^2 \right)+9y^2e^{\frac{1}{x}+y^3}=0$$
\end{tcolorbox}

#### (c)
$u_x^2+u_y^2=1$, use $u(x,y)=f(x)+g(y)$
$$u_x=f', u_y=g'$$
$$(f')^2+(g')^2=1\quad \longrightarrow \quad (f')^2=1-(g')^2=\lambda^2$$
$$(f')^2=\lambda^2\quad \longrightarrow \quad f'=\lambda\quad \longrightarrow \quad f(x)=\lambda x+C_1$$
$$(g')^2=1-\lambda^2 \quad \longrightarrow \quad g'=\pm\sqrt{1-\lambda^2}\quad \longrightarrow \quad g(y)=\pm y\sqrt{1-\lambda^2} + C_2$$
\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=f(x)+g(y)=\lambda x \pm y\sqrt{1-\lambda^2}+C$$
\end{tcolorbox}
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$u_x=\lambda, u_y=\pm \sqrt{1-\lambda^2},\quad u_x^2+u_y^2= \lambda ^2 + 1 - \lambda^2 = 1$$
\end{tcolorbox}

#### (d)
$x^2u_x^2+y^2u_y^2=u^2$, use $u(x,y)=e^{f(x)}e^{g(y)}$

$$u_x= f'e^{f+g}, u_y= g'e^{f+g}$$
$$x^2u_x^2+y^2u_y^2-u^2=x^2(f')^2e^{2f+2g}+y^2(g')^2e^{2f+2g}-e^{2f+2g}=0$$
$$x^2(f')^2+y^2(g')^2-1=0 \quad \longrightarrow \quad x^2(f')^2=1-y^2(g')^2=\lambda^2$$
$$x^2(f')^2= \lambda^2 \quad \longrightarrow \quad f'=\frac{\lambda}{x} \quad \longrightarrow \quad f(x)=\lambda \ln(x)+C_1$$
$$y^2(g')^2= 1-\lambda^2 \quad \longrightarrow \quad g'=\frac{\sqrt{1-\lambda^2}}{y} \quad \longrightarrow \quad g(y)=\sqrt{1-\lambda^2}\ln(y) +C_2$$

$$u(x,y)=e^{f(x)}e^{g(y)}=C_3e^{\lambda \ln(x)}C_4e^{\sqrt{1-\lambda^2}\ln(y)}$$

\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=C x^\lambda y^{\sqrt{1-\lambda^2}}$$
\end{tcolorbox}
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$u_x=\lambda C x^{\lambda-1} y^{\sqrt{1-\lambda^2}},u_y= \sqrt{1-\lambda^2} C x^\lambda y^{\sqrt{1-\lambda^2}-1}$$
$$x^2u_x^2+y^2u_y^2=x^2\left(\lambda C x^{\lambda-1} y^{\sqrt{1-\lambda^2}} \right)^2+y^2 \left(\sqrt{1-\lambda^2} C x^\lambda y^{\sqrt{1-\lambda^2}-1}\right)^2 =$$

$$x^2\lambda^2C^2 x^{2\lambda-2}y^{2\sqrt{1-\lambda^2}}+y^2(1-\lambda^2)C^2x^{2\lambda} y^{2\sqrt{1-\lambda^2}-2}=$$
$$\lambda^2 C^2 x^{2\lambda}y^{2\sqrt{1-\lambda^2}}+(1-\lambda^2)C^2x^{2\lambda} y^{2\sqrt{1-\lambda^2}} = C^2x^{2\lambda} y^{2\sqrt{1-\lambda^2}}=u^2$$
\end{tcolorbox}

### Question 4
For each of the following IVPs, (i) find and plot the characteristic lines (curves), (ii) solve the IVP, and (iii) plot the solution of (a)-(b) for indicated time.

#### (a)
$u_t+\left(1+x^2\right)u_x=0, \, u(0,x)=\arctan{(x)}, \, t=1,2,3$, and
what is $\lim\limits_{t \to \infty}u(t,x)$?

$$\frac{dx}{dt}=1+x^2$$
$$\frac{1}{1+x^2}dx=dt$$

$$\arctan{(x)}=t+C$$
So $x=\tan{(t+C)}$ are the characteristic curves of $u_t+\left(1+x^2\right)u_x=0$.  
On each of the curves, $u(x,t)$ is constant because $$\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x} \sec^2{(t+c)}$$  
And since $1+\tan^2\theta=\sec^2\theta$, this means $\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x}\left( 1+\tan^2{(t+c)}\right)=u_t+\left(1+x^2\right)u_x$, which we know is 0.

Thus $u(t, \tan{(t+C)})=u(0, \tan{(0+C)})=u(0, C)$ is independent of $t$.  
Putting $x=\tan{(t+C)}$ and $C=\arctan{(x)}-t$, we have
$$u(t, x)=u(0,\arctan{(x)}-t)$$
It follows that $u(t,x)=f(\arctan{(x)}-t)$.  
And since we are given that $u(0,x)=\arctan{(x)}$, we have $f(\arctan{(x)}-0)=\arctan{(x)}$ so that $f(w)=w$ for any $w$, yielding our solution of

\begin{tcolorbox}[colback=white, title=Solution]
$$u(t,x)=\arctan{(x)}-t$$
\end{tcolorbox}

```{r, echo=FALSE}
C=1
df_4a <- tibble(x = seq(-2*pi, 2*pi, length.out = 1000),
                y=tan(x+C)) %>% 
  mutate(y = ifelse(abs(y) < 20, y, NA))

ggplot(df_4a, aes(x=x, y=y)) + 
  geom_line()+
  coord_cartesian(ylim=c(-10,10), xlim = c(-5,5))+
  theme_bw()+
  labs(x="t", y="x")+
  ggtitle("Characteristic curves for Problem 4a", subtitle = "with C=1")
```

Since the range of $\arctan$ is limited to $\left(\frac{-\pi}{2},\frac{\pi}{2}\right)$, as $\lim\limits_{t \to \infty}$ that means $u(t,x)=\arctan(x)-t$ will go to $-\infty$.

```{r, echo=FALSE}

x=seq(-10,10,length.out = 1000)
ggplot() + 
  geom_line(aes(x=x, y=atan(x)-1, color="1"))+
  geom_line(aes(x=x, y=atan(x)-2, color="2"))+
  geom_line(aes(x=x, y=atan(x)-3, color="3"))+
  theme_bw()+
  labs(x="x", y="u(x,t)", color="Value of t")+
  ggtitle("Solution curves for Problem 4a")
```

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
Since $u_t=-1, u_x=\frac{1}{1+x^2}$, that means $u_t+\left(1+x^2\right)u_x=-1+1=0$
\end{tcolorbox}

#### (b)
$u_t-xu_x=0, \, u(0,x)=\frac{1}{1+x^2}, \, t=1,2,3$, and
what is $\lim\limits_{t \to \infty}u(t,x)$?

The directional derivative of $u$ in the direction of the vector $(1, -x)$ is zero. The curves in the $tx$ plane with $(1, -x)$ as a tangent vector have slopes $-x$:

$$\frac{dx}{dt}=-x$$
$$\frac{1}{x}dx=-dt$$
Solving this ODE gives the equations for the characteristic lines:
$$x=Ce^{-t}$$

```{r, warning = FALSE, echo=FALSE}
curves_4b <- tibble(C=c(1:5), seq=list(x)) %>% 
  rowwise() %>% 
  mutate(y=list(C*exp(-seq))) %>% 
  unnest(c(seq,y))

ggplot(data=curves_4b)+
  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
  coord_cartesian(ylim=c(NA, 100), xlim=c(-5, 9))+
  labs(x="t", y="x", color="C")+
  theme_light()+
  ggtitle("Characteristic curves for Problem 4b")
```

On each of the curves, $u(x,t)$ is constant because $$\frac{d}{dt}u(t,Ce^{-t}) = \pdv {u}{t}-Ce^{-t}\pdv{u}{x}=u_t-xu_x$$  
and we know that $u_t-xu_x=0$.

Set $\xi=xe^t$ and $\eta=x$. Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
$$\pdv{u}{\eta}+\pdv{u}{\xi}e^t $$

and similarly
$$u_t= \pdv{u}{t}=\pdv{u}{\eta}\pdv{\eta}{t}+\pdv{u}{\xi}\pdv{\xi}{t}=\pdv{u}{\xi}xe^t $$

Which means that 
$$u_t-xu_x=\pdv{u}{\xi}xe^t -x\left(\pdv{u}{\eta}+\pdv{u}{\xi}e^t \right)=0$$
Assuming $x\neq0$, this means
$$u_\eta=0$$
$$u=\int u_\eta \, d\eta \> + f(\xi)=0+ f(\xi)$$

So we have $u(t,x)=f(\xi)=f(xe^t)$ and we know that $u(0,x)=\frac{1}{1+x^2}$, which means $f(xe^0)=f(x)=\frac{1}{1+x^2}$ and therefore
\begin{tcolorbox}[colback=white, title=Solution]
$$u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$u_t=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)xe^t$$
$$u_x=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)e^t$$
And we see that $u_t-xe_t$ indeed equals zero.
\end{tcolorbox}

Plotting the solution $u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$ as a function of $x$ for $t=1,2,3$, we get:

```{r, warning=FALSE, echo=FALSE}

x=seq(-10,10,length.out = 1000)
ggplot() + 
  geom_line(aes(x=x, y=1/(1+(exp(1)*x)^2), color="1"))+
  geom_line(aes(x=x, y=1/(1+(exp(2)*x)^2), color="2"))+
  geom_line(aes(x=x, y=1/(1+(exp(3)*x)^2), color="3"))+
  theme_bw()+
  labs(x="x", y="u(x,t)", color="Value of t")+
  xlim(-5,5)+
  ggtitle("Solution curves for Problem 4b")
```

We can see that $u(t,x)$ will equal 1 at $x=0$ regardless of the value of $t$, but at all other values of $x$, $\lim\limits_{t \to \infty}u(t,x)=0$. This can be written as
\begin{equation}
\lim\limits_{t \to \infty}u(t,x)=
    \begin{cases}
        1 & \text{if } x=0\\
        0 & \text{if } x\neq0
    \end{cases}
\end{equation}

#### (c)
$u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2, \, u(0,y)=e^y$

$\frac{dy}{dx}=x$
Solving this ODE yields the equation for the characteristic curves: $y=\frac{x^2}{2}+C$

```{r, warning = FALSE, echo=FALSE}
curves_4c <- tibble(C=c(1:5), seq=list(x)) %>% 
  rowwise() %>% 
  mutate(y=list(0.5*(seq)^2+C)) %>% 
  unnest(c(seq,y))

ggplot(data=curves_4c)+
  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
  coord_cartesian(ylim=c(NA, 45), xlim=c(-9.5, 9.5))+
  labs(x="x", y="y", color="C")+
  theme_light()+
  ggtitle("Characteristic curves for Problem 4c")
```

Now we define our new coordinate system as $\xi=y-\frac{x^2}{2}, \, \eta=x$

Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
$$\pdv{u}{\eta}-x\pdv{u}{\xi}$$

and similarly
$$u_y= \pdv{u}{y}=\pdv{u}{\eta}\pdv{\eta}{y}+\pdv{u}{\xi}\pdv{\xi}{y}=$$
$$\pdv{u}{\xi}$$

$$u_x+xu_y=\pdv{u}{\eta}-x\pdv{u}{\xi}+x\pdv{u}{\xi}=\pdv{u}{\eta}$$

And since we know $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, this means we have
$$\pdv{u}{\eta}=\left(y-\frac{x^2}{2}\right)^2=\xi^2$$
Now we integrate with respect to $\eta$ to find
$$u=\eta\xi^2+F(\xi)$$

At $u(0,y)$, this becomes $u=0y^2+F(y)$, or $u(0,y)=F(y)$.
On the other hand, we are told that $u(0,y)=e^y$, which means that $F(y)=e^y$.
$$u=\eta\xi^2+e^\xi$$
Substitute back in $x$ and $y$:
\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=x\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}$$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$u_x=-2x^2\left(y-\frac{x^2}{2}\right)+\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}(-x)$$
$$u_y=2x\left(y-\frac{x^2}{2}\right)+\exp{\left(y-\frac{x^2}{2}\right)}$$
Which means that $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, as we wanted.
\end{tcolorbox}

#### (d)
$2xu_x+(x+1)u_y=y$, for $x>0, u=2y$ on $x=1$

$$\frac{dx}{dy}=\frac{x+1}{2x}=\frac{1}{2}+\frac{1}{2x}$$
Solving this ODE yields the characteristic curves:
$$y=\frac{x}{2}+\frac{\ln(x)}{2}+C$$
$$C=2y-x-\ln(x)$$

```{r, warning = FALSE, echo=FALSE}
curves_4d <- tibble(C=c(1:5), seq=list(x)) %>% 
  rowwise() %>% 
  mutate(y=list((seq/2)+(log(seq)/2)+C)) %>% 
  unnest(c(seq,y))

ggplot(data=curves_4d)+
  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
  coord_cartesian(ylim=c(-1, 12), xlim=c(-1, 9.5))+
  labs(x="x", y="y", color="C")+
  theme_light()+
  ggtitle("Characteristic curves for Problem 4d")
```