### 2024-02-09 PDEs/math245_hw2.Rmd
math hw proiblem 2

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-09"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1
Determine the region in which the given equation is hyperbolic, parabolic, elliptic, or
singular.

a) $$u_{xx}+y^2u_{yy}+u_x-u+x^2=0 $$

b) $$u_{xx}-yu_{yy}+xu_x+yu_y+u=0 $$

## Question 2
Using a factorization similar to the wave equation, solve the following IVP:
\begin{equation}
    \begin{cases}
      u_{xx}+2u_{xy}-3u_{yy}=0 & x\subset \mathbb{R},\; y> 0\\
      u(0,x)=\sin{x} & x\subset \mathbb{R}\\
      u_y(0,x) = x & x\subset \mathbb{R}
    \end{cases}       
\end{equation}

First, we can factor the equation as follows:
$$\left( \pdv{}{x}+3\pdv{}{y}\right) \left(\pdv{}{x}-\pdv{}{y}\right)u=0$$
or $$\left( \partial_{x}+3\partial_{y}\right) \left(\partial_{x}-\partial_{y}\right)u=0$$

Then set $\left(\pdv{}{x}-\pdv{}{y}\right)u=v$, giving us

$\left( \pdv{}{x}+3\pdv{}{y}\right)v= v_x+3v_t=0$

which we know has the solution $v(x,y)=f(3x-y)$, so
$$u_x-u_y=f(3x-y)$$

On (characteristic) lines with the slope $y=-x+c$, or $y+x=\text{constant}$, we must have $u_x-u_y=f(3x-y)=0$.
Set $\eta=x+y$ and $\xi=x$. 
Then by the chain rule,
$$u_x= u_{\eta}+u_{\xi}, \; u_y=u_{\eta}$$
And let's rewrite $y$ as $y=\eta-x=\eta-\xi$.

So
$$u_x-u_y=f(3x-y) \quad\longrightarrow\quad u_{\xi}=f(3\xi-\eta+\xi)$$

$$u_{\xi}=f(4\xi-\eta)$$
Now integrate with respect to $\xi$:
$$u(\eta, \xi)=F(4\xi-\eta)+G(\eta)$$
where $F$ is the antiderivative of $f$ with respect to $\xi$.

Now convert back to our original variables:
$$u(x,y)=F(3x-y)+G(x+y)$$

Using the fact that $u(0,x)=\sin{x}$,

 \begin{equation} \label{eq:1} u(0,x)=\sin{x}=F(3x)+G(x) \end{equation}
now replace $x$ with a new neutral variable, $\alpha$ and differentiate:
 $$\sin{\alpha}=F(3\alpha)+G(\alpha) $$
\begin{equation} \label{eq:2} \cos{\alpha}=3F'(3\alpha)+G'(\alpha) \end{equation}

But we can also differentiate $u(x,y)=F(3x-y)+G(x+y)$ with respect to $y$ to get
$$u_y(x,y)=-F'(3x-y)+G'(x+y)$$
but from our initial conditions, we know
$$u_y(0,x)=-F'(3x-0)+G'(x+0)=x$$

Let's replace $x$ by our neutral variable $\alpha$ and solve for $F'$:
$$F'(\alpha)=G'(3\alpha)-\alpha$$
Now plug this into \ref{eq:2}:
$$\cos{\alpha}=3G'(\alpha)-3\alpha+G'(\alpha)$$

$$G(\alpha)= \frac{1}{4} \int {\cos{\alpha}+3\alpha} = \frac{\sin{\alpha}}{4}+ \frac{3 \alpha^2}{8} $$
So that means \ref{eq:1} becomes:
$$\sin{\alpha}=F(3\alpha)+\frac{\sin{\alpha}}{4}+ \frac{3 \alpha^2}{8}$$

$$F(\alpha)=\frac{3\sin{(\frac{\alpha}{3})}}{4}- \frac{\alpha^2}{24}$$

Which means $u(x,y)=F(3x-y)+G(x+y)$ becomes

$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}- \frac{(3x-y)^2}{24}+  \frac{\sin{(x+y)}}{4}+ \frac{3(x+y)^2}{8}=$$

\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}+\frac{\sin{(x+y)}}{4}+ xy+  \frac{y^2}{3} $$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]

 $$u_y=\frac{-1}{4}\cos{\left(x-\frac{y}{3}\right)}+\frac{\cos{(x+y)}}{4}+ x+  \frac{2y}{3} $$
  $$u_{yy}=\frac{-1}{12}\sin{\left(x-\frac{y}{3}\right)}-\frac{\sin{(x+y)}}{4}+ \frac{2}{3} $$

$$u_x=\frac{3}{4}\cos{\left(x-\frac{y}{3}\right)}+\frac{\cos{(x+y)}}{4}+y$$

$$u_{xx}=\frac{-3}{4} \sin{\left( x-\frac{y}{3} \right)} - \frac{\sin{(x+y)}}{4}$$

$$u_{xy}=\frac{1}{4}\sin{\left(x-\frac{y}{3}\right)}-\frac{\sin{(x+y)}}{4}+1$$

Check that $u_{xx}+2u_{xy}-3u_{yy}=0$
$$
\left(\frac{-3}{4}+ \frac{2}{4}+\frac{1}{4}\right)\sin{\left(x-\frac{y}{3}\right)}+
\left(\frac{-1}{4}+ \frac{-2}{4}+\frac{3}{4}\right)\sin{(x+y)}+(0+2-3)=0
$$

\end{tcolorbox}

## Question 3