### 2024-02-13 PDEs/math245_hw2.Rmd
more PDEs

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-13"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Done, feel OK: 1, 2
## Done, very low confidence: 4, 6
## Mostly done: 7
## Not started: 3, 5, 8

## Question 1
Determine the region in which the given equation is hyperbolic, parabolic, elliptic, or
singular.

a) $$u_{xx}+y^2u_{yy}+u_x-u+x^2=0 $$

$a=1, b=0, c=-y^2$, so we have $b^2-ac=0-(-y^2)=y^2$. This will be positive everywhere except for $y=0$, so the equation is hyperbolic where $y \neq 0$ and parabolic for $y=0$.

b) $$u_{xx}-yu_{yy}+xu_x+yu_y+u=0 $$

$a=1, b=0, c=-y$, so we have $b^2-ac=0-(-y)=y$. Thus, the equation will be hyperbolic where $y>0$, parabolic where $y=0$, and elliptic where $y<0$.

## Question 2
Using a factorization similar to the wave equation, solve the following IVP:
\begin{equation}
    \begin{cases}
      u_{xx}+2u_{xy}-3u_{yy}=0 & x\subset \mathbb{R},\; y> 0\\
      u(0,x)=\sin{x} & x\subset \mathbb{R}\\
      u_y(0,x) = x & x\subset \mathbb{R}
    \end{cases}       
\end{equation}

First, we can factor the equation as follows:
$$\left( \pdv{}{x}+3\pdv{}{y}\right) \left(\pdv{}{x}-\pdv{}{y}\right)u=0$$
or $$\left( \partial_{x}+3\partial_{y}\right) \left(\partial_{x}-\partial_{y}\right)u=0$$

Then set $\left(\pdv{}{x}-\pdv{}{y}\right)u=v$, giving us

$\left( \pdv{}{x}+3\pdv{}{y}\right)v= v_x+3v_t=0$

which we know has the solution $v(x,y)=f(3x-y)$, so
$$u_x-u_y=f(3x-y)$$

On (characteristic) lines with the slope $y=-x+c$, or $y+x=\text{constant}$, we must have $u_x-u_y=f(3x-y)=0$.
Set $\eta=x+y$ and $\xi=x$. 
Then by the chain rule,
$$u_x= u_{\eta}+u_{\xi}, \; u_y=u_{\eta}$$
And let's rewrite $y$ as $y=\eta-x=\eta-\xi$.

So
$$u_x-u_y=f(3x-y) \quad\longrightarrow\quad u_{\xi}=f(3\xi-\eta+\xi)$$

$$u_{\xi}=f(4\xi-\eta)$$
Now integrate with respect to $\xi$:
$$u(\eta, \xi)=F(4\xi-\eta)+G(\eta)$$
where $F$ is the antiderivative of $f$ with respect to $\xi$.

Now convert back to our original variables:
$$u(x,y)=F(3x-y)+G(x+y)$$

Using the fact that $u(0,x)=\sin{x}$,

 \begin{equation} \label{eq:1} u(0,x)=\sin{x}=F(3x)+G(x) \end{equation}
now replace $x$ with a new neutral variable, $\alpha$ and differentiate:
 $$\sin{\alpha}=F(3\alpha)+G(\alpha) $$
\begin{equation} \label{eq:2} \cos{\alpha}=3F'(3\alpha)+G'(\alpha) \end{equation}

But we can also differentiate $u(x,y)=F(3x-y)+G(x+y)$ with respect to $y$ to get
$$u_y(x,y)=-F'(3x-y)+G'(x+y)$$
but from our initial conditions, we know
$$u_y(0,x)=-F'(3x-0)+G'(x+0)=x$$

Let's replace $x$ by our neutral variable $\alpha$ and solve for $F'$:
$$F'(\alpha)=G'(3\alpha)-\alpha$$
Now plug this into \ref{eq:2}:
$$\cos{\alpha}=3G'(\alpha)-3\alpha+G'(\alpha)$$

$$G(\alpha)= \frac{1}{4} \int {\cos{\alpha}+3\alpha} = \frac{\sin{\alpha}}{4}+ \frac{3 \alpha^2}{8} $$
So that means \ref{eq:1} becomes:
$$\sin{\alpha}=F(3\alpha)+\frac{\sin{\alpha}}{4}+ \frac{3 \alpha^2}{8}$$

$$F(\alpha)=\frac{3\sin{(\frac{\alpha}{3})}}{4}- \frac{\alpha^2}{24}$$

Which means $u(x,y)=F(3x-y)+G(x+y)$ becomes

$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}- \frac{(3x-y)^2}{24}+  \frac{\sin{(x+y)}}{4}+ \frac{3(x+y)^2}{8}=$$

\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}+\frac{\sin{(x+y)}}{4}+ xy+  \frac{y^2}{3} $$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]

 $$u_y=\frac{-1}{4}\cos{\left(x-\frac{y}{3}\right)}+\frac{\cos{(x+y)}}{4}+ x+  \frac{2y}{3} $$
  $$u_{yy}=\frac{-1}{12}\sin{\left(x-\frac{y}{3}\right)}-\frac{\sin{(x+y)}}{4}+ \frac{2}{3} $$

$$u_x=\frac{3}{4}\cos{\left(x-\frac{y}{3}\right)}+\frac{\cos{(x+y)}}{4}+y$$

$$u_{xx}=\frac{-3}{4} \sin{\left( x-\frac{y}{3} \right)} - \frac{\sin{(x+y)}}{4}$$

$$u_{xy}=\frac{1}{4}\sin{\left(x-\frac{y}{3}\right)}-\frac{\sin{(x+y)}}{4}+1$$

Check that $u_{xx}+2u_{xy}-3u_{yy}=0$
$$
\left(\frac{-3}{4}+ \frac{2}{4}+\frac{1}{4}\right)\sin{\left(x-\frac{y}{3}\right)}+
\left(\frac{-1}{4}+ \frac{-2}{4}+\frac{3}{4}\right)\sin{(x+y)}+(0+2-3)=0
$$

\end{tcolorbox}

## Question 3

Solve the Neumann boundary value problem for the wave equation on half line:
\begin{equation}
    \begin{cases}
      u_{tt}=c^2u_{xx}+f(t,x) & 0<x<\infty\\
      u(0,x)=\phi{x} & 0<x<\infty\\
      u_t(0,x) = \psi{x} & 0<x<\infty\\
      u_x(t,0) = h(t) & t>0
    \end{cases}       
\end{equation}

## Question 4

Consider the 3D wave equation for $u(t,x,y,z)$:

$$u_{tt}=c^2\Delta u \quad\quad (x,y,z) \in \mathbb{R}^3, \quad t>0$$

Change the coordinates to spherical coordinates. Assume the solution is spherically symmetric, so that $u(t,x,y,z)=u(t,r)$ and does not depend on $\theta$ and $\phi$. Find the solution for $u(0,r)=0$ and

\begin{equation} u_t(0,r) = \begin{cases}
  1  & \lvert{r}\rvert \leq 1 \\
  0 &  \lvert{r}\rvert > 1
\end{cases}
\end{equation}

Hint: use the substitution $u(t,r)= \frac{1}{r} w(t,r)$.

First, we need to derive the formula for the Laplacian in spherical coordinates.

We know the equation for the Laplacian in polar coordinates is:

$$\frac{\partial^2}{\partial r^2}+\frac{1}{r}\frac{\partial}{\partial r}+ \frac{1}{r^2}\frac{\partial^2}{\partial \theta^2}$$

Now let's convert to spherical coordinates:

$$r=\sqrt{x^2+y^2+z^2}=\sqrt{s^2+z^2}$$

$$x=s \cos{\phi}$$
$$y=s \sin{\phi}$$
$$z=r \cos{\theta}$$
$$s=r \sin{\theta}$$

By the two-dimensional Laplacian, we have

$$u_{zz}+u_{ss}=u_{rr}+\frac{1}{r}u_r+\frac{1}{r^2}u_{\theta\theta}$$ and

$$u_{xx}+u_{yy}=u_{ss}+\frac{1}{s}u_s+\frac{1}{s^2}u_{\phi\phi}$$
We add these two equations and cancel $u_s$ to get

$$\Delta u = u_{rr}+\frac{1}{r}u_r+\frac{1}{r^2}u_{\theta\theta}+\frac{1}{s}u_s+\frac{1}{s^2}u_{\phi\phi}$$

Now since $u$ doesn't depend on $\theta$ or $\phi$, we have
$$\Delta u = u_{rr}+\frac{1}{r}u_r+\frac{1}{s}u_s=u_{rr}+\frac{1}{r}u_r+\frac{1}{r \sin{\theta}}u_s$$

$$ \frac{\partial u}{\partial s} = \frac{\partial u}{\partial r} \frac{\partial r}{\partial s}+\frac{\partial u}{\partial\theta}\frac{\partial\theta}{\partial s}+\frac{\partial u}
{\partial\phi}\frac{\partial\phi}{\partial s}=u_r\frac{1}{ \sin{\theta}}+0+0=u_r\frac{s}{ r}$$

So with our change of variables, we have

$$u_{tt}=c^2 \left(u_{rr} + \frac{2}{r}u_r \right)$$

Now set $w=ru$, or $u= \frac{w}{r}$. Then 
$$w_t=ru_t, \quad w_{tt}=ru_{tt}, \quad u_{tt}= \frac{w_{tt}}{r} $$
$$w_t=ru_t, \quad w_{tt}=ru_{tt}, \quad u_{tt}= \frac{w_{tt}}{r} $$
$$u_r = \frac{w_r}{r}-\frac{w}{r^2} $$

$$u_{rr} = \frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} $$
So $$u_{tt}=c^2 \left(u_{rr} + \frac{2}{r}u_r \right)$$ becomes

$$\frac{w_{tt}}{r} =c^2 \left(\frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} + \frac{2}{r} \left(\frac{w_r}{r}-\frac{w}{r^2} \right) \right)$$
which simplifies to
$$w_{tt}=c^2 w_{rr}$$
but this is just the wave equation, which we know has the solution
 $$w(t,r)= \frac{\varphi(r+ct)+\varphi(r-ct)}{2}+\frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$
 Since $\varphi=0$, 
 $$w(t,r)= \frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$

Now we have 4 cases:

Case 1: $r-ct \geq -1, r+ct \leq 1$
$$w(t,r)= \frac{1}{2c} \int_{r-ct}^{r+ct}{s\;ds}$$

Case 2:$r-ct < -1, r+ct>1$
$$w(t,r)= \frac{1}{2c} \int_{-1}^{1}{s\;ds}$$

Case 3:$r-ct < -1, r+ct \leq 1$
$$w(t,r)= \frac{1}{2c} \int_{-1}^{r+ct}{s\;ds}$$

Case 4:$r-ct \geq -1, r+ct>1$
$$w(t,r)= \frac{1}{2c} \int_{r-ct}^{1}{s\;ds}$$

Since $u=\frac{w}{r}$, this means we have
\begin{equation}u(t,r)=
    \begin{cases} 
      \frac{1}{2c} \int_{r-ct}^{r+ct}{s\;ds} & r-ct \geq -1, r+ct \leq 1\\
      \\
      \frac{1}{2c} \int_{-1}^{1}{s\;ds} & r-ct < -1, r+ct > 1\\
      \\
      \frac{1}{2c} \int_{-1}^{r+ct}{s\;ds} & r-ct < -1, r+ct \leq 1\\
      \\
      \frac{1}{2c} \int_{r-ct}^{1}{s\;ds} & r-ct \geq -1, r+ct > 1\\

    \end{cases}       
\end{equation}

## Question 5
Consider the following Dirichlet boundary value problem:
\begin{equation} 
\begin{cases}
  u_{tt}+x(t,x)u_t=u_{xx} & 0<x<1 \\
  u(0,x)=\phi(x) & 0<x<1 \\
  u_t(0,x)=\psi(x) & 0<x<1 \\
  u(t,0) = u(t,1)=0 & t \geq 0 \\
\end{cases}
\end{equation}

Assume that $\lvert a(t,x)\rvert \leq m$ for some constant $m$ and all $0<x<1$ and $t\geq 0$. Let 
$$E(t)= \frac{1}{2} \int_{0}^{1}{ \left(u_t(t,x)^2+u_x(t,x)^2\right)}\:dx $$

(1) Show that $\frac{dE(t)}{dt} \leq 2mE(t)$ for $t \geq 0$.
(2) Use part (a) and show that $\frac{d}{dt} \left(e^{-2mE(t)}\right) \leq0$ for 
all $t \geq 0$. Hence, by integration from $[0,t]$, we get that

$$E(t) \leq e^{2mt}E(0) \quad \text{for all } t \geq 0$$

(3) If $\phi(x)=\psi(x)=0$ for all $0<x<1$, what does this say about $E(t)$ for $t \geq 0$ and hence about $u(t,x)$ for $t \geq 0$?

(4) Use the previous part to prove uniqueness of the following problem:
\begin{equation} 
\begin{cases}
  u_{tt}+a(t,x)u_t=u_{xx} & 0<x<1, t>0 \\
  u(0,x)=\phi(x) & 0<x<1 \\
  u_t(0,x)=\psi(x) & 0<x<1 \\
  u(t,0) = f(t) & t \geq 0 \\
  u(t,1) = g(t) & t \geq 0 \\
\end{cases}
\end{equation}

## Problem 6

Does the D’Alembert method work for the wave equation $u_{tt} = c(x)^2u_{xx}$? What about $u_{tt} = c(t)^2u_{xx}$? Why?

Let's try the factorization method if $c=c(x)$:
$$\left( \partial_{t}+c(x)\partial_{x}\right) \left(\partial_{t}-c(x)\partial_{x}\right)u=0$$
Set $\xi=x+c(x)t$ and $\eta=x-c(x)t$.

By the chain rule,

$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}$$

$$=\partial_{\xi}(1+c')+\partial_{\eta}\,(1-c')$$

$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}$$
$$=c\partial_{\xi}-c\partial_{\eta}$$
When we plug these back into
$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
we get

$$ \left(c\partial_{\xi}-c\partial_{\eta}+c\partial_{\xi}(1+c')+c\partial_{\eta}\,(1-c')\right)
\left(c\partial_{\xi}-c\partial_{\eta} -c\partial_{\xi}(1+c')-c\partial_{\eta}\,(1-c')\right)u=0$$

When $c$ is constant, $c'=0$ and this simplifies to $-4c^2 u_{\xi\eta}=0$, allowing us to integrate to find $u(x,t)=F(x+ct)+G(x-ct)$. But we cannot do the same simplification when we have the $c'$ terms, meaning we cannot solve the wave equation by the same method when the wave speed is not constant.

Similarly,

Let's try the factorization method if $c=c(t)$:
$$\left( \partial_{t}+c(t)\partial_{x}\right) \left(\partial_{t}-c(t)\partial_{x}\right)u=0$$
Set $\xi=x+c(t)t$ and $\eta=x-c(t)t$.

By the chain rule,

$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}=
\partial_{\xi}+\partial_{\eta}$$

$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}=\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)
$$

When we plug these back into
$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
we get

$$\left( \partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)+c\partial_{\xi}+c\partial_{\eta} \right)
\left(\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)-c\partial_{\xi}-c\partial_{\eta}\right)u=0$$
Where again we cannot cancel terms to simplify our equation as we could with constant $c$.

## Problem 7 (The Poisson-Darboux Equation)   
Solve the initial value problem

\begin{equation} 
\begin{cases}
  u_{tt}-u_{xx}- \frac{2}{x}u_x=0 & -\infty <x<\infty, t>0 \\
  u(0,x)=0 & -\infty <x<\infty \\
  u_t(0,x)=g(x) & -\infty <x<\infty \\
\end{cases}
\end{equation}

where $g(x)=g(-x)$ is an even function. Hint: set $w=xu$.

Using the results from when we set $w=ru$ in Problem 4,
$$u_t=\frac{w_t}{x}, u_{tt}=\frac{w_{tt}}{x} $$
$$u_x= \frac{w_x}{x}- \frac{w}{x^2} $$

$$u_{xx}=  \frac{w_{xx}}{x}-\frac{2w_x}{x^2}+\frac{2w}{x^3}$$
So $u_{tt}-u_{xx}- \frac{2}{x}u_x=0$ becomes

$$\frac{w_{tt}}{x} -\frac{w_{xx}}{x}+\frac{2w_x}{x^2}- \frac{2w}{x^3}- \frac{2}{x} \left(\frac{w_x}{x}- \frac{w}{x^2}\right)=0 $$
Which simplifies to $w_{tt}-w_{xx}=0$, the wave equation with $c=1$.
By the D'Alembert formula, since we have $\varphi=0$,
$$w(t,x)=\frac{1}{2c} \int_{x-t}^{x+t}{\psi(s)ds}$$
$w_t=xu_t$
Since $y=x$ is an odd function and $u_t(0,x)=g(x)$ is an even function, $w_t=xg(x)=h(x)$ must be an odd function.

## Problem 8   
Solve the following characteristic initial value problem:
\begin{equation} 
\begin{cases}
  y^3u_{xx}-yu_{yy}+u_y=0 & 0<x<4, \quad \lvert y \rvert \leq 2 \sqrt{2} \\
  u(x,y)= f(x) & x+ \frac{y^2}{2}=4 \text{ for } 2 \leq x \leq 4 \\
  u(x,y)=g(x) & x- \frac{y^2}{2}=0 \text{ for } 0 \leq x \leq 2 \\
\end{cases}
\end{equation}

where $f(2)=g(2)$. Hint: Use the transformation $\eta=x-\frac{y^2}{2}$ and $\xi=x+\frac{y^2}{2}$ and express the PDE in the coordinates $(\xi, \eta)$.