### 2024-02-16 PDEs/math245_hw2.Rmd
finished PDE prob 8

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-16"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Done, feel OK: 1, 2, 5, 7, 8
## Done, low confidence: 4
## Redo: 6
## Not started: 3

## Question 1
Determine the region in which the given equation is hyperbolic, parabolic, elliptic, or
singular.

a) $$u_{xx}+y^2u_{yy}+u_x-u+x^2=0 $$

$a=1, b=0, c=-y^2$, so we have $b^2-ac=0-(-y^2)=y^2$. This will be positive everywhere except for $y=0$, so the equation is hyperbolic where $y \neq 0$ and parabolic for $y=0$.

b) $$u_{xx}-yu_{yy}+xu_x+yu_y+u=0 $$

$a=1, b=0, c=-y$, so we have $b^2-ac=0-(-y)=y$. Thus, the equation will be hyperbolic where $y>0$, parabolic where $y=0$, and elliptic where $y<0$.

## Question 2
Using a factorization similar to the wave equation, solve the following IVP:
\begin{equation*}
    \begin{cases}
      u_{xx}+2u_{xy}-3u_{yy}=0 & x\subset \mathbb{R},\; y> 0\\
      u(0,x)=\sin{x} & x\subset \mathbb{R}\\
      u_y(0,x) = x & x\subset \mathbb{R}
    \end{cases}       
\end{equation*}

First, we can factor the equation as follows:
$$\left( \pdv{}{x}+3\pdv{}{y}\right) \left(\pdv{}{x}-\pdv{}{y}\right)u=0$$
or $$\left( \partial_{x}+3\partial_{y}\right) \left(\partial_{x}-\partial_{y}\right)u=0$$

Then set $\left(\pdv{}{x}-\pdv{}{y}\right)u=v$, giving us

$\left( \pdv{}{x}+3\pdv{}{y}\right)v= v_x+3v_t=0$

which we know has the solution $v(x,y)=f(3x-y)$, so
$$u_x-u_y=f(3x-y)$$
Now we can reorder our equation as $$\left(\partial_{x}-\partial_{y}\right)
\left( \partial_{x}+3\partial_{y}\right)u=0$$
and set $w=\left( \partial_{x}+3\partial_{y}\right)u$

Then $$w_x-w_y=0 $$

Which we know has the solution $w(x,y)=g(x+y)$. So $u_x+3u_y=g(x+y)$, which gives us a system of two equations:

\begin{equation*}
\begin{cases}
u_x-u_y=f(3x-y)
u_x+3u_y=g(x+y)
\end{cases}
\end{equation*}

Subtract the first equation from the second:
$$4u_y=-f(3x-y)+g(x+y) $$

Now we can integrate with respect to $y$ to find that:
$$u(x,y)=F(3x-y)+G(x+y)$$
where $F$ is the antiderivative of $-f$ with respect to $y$ and $G$ is the antiderivative of $g$ with respect to $y$.

Using the fact that $u(0,x)=\sin{x}$,

 \begin{equation} \label{eq:1} u(0,x)=\sin{x}=F(3x)+G(x) \end{equation}
now replace $x$ with a new neutral variable, $\alpha$ and differentiate:
 $$\sin{\alpha}=F(3\alpha)+G(\alpha) $$
\begin{equation} \label{eq:2} \cos{\alpha}=3F'(3\alpha)+G'(\alpha) \end{equation}

But we can also differentiate $u(x,y)=F(3x-y)+G(x+y)$ with respect to $y$ to get
$$u_y(x,y)=-F'(3x-y)+G'(x+y)$$
but from our initial conditions, we know
$$u_y(0,x)=-F'(3x-0)+G'(x+0)=x$$

Let's replace $x$ by our neutral variable $\alpha$ and solve for $F'$:
$$F'(\alpha)=G'(3\alpha)-\alpha$$
Now plug this into (\ref{eq:2}):
$$\cos{\alpha}=3G'(\alpha)-3\alpha+G'(\alpha)$$

$$G(\alpha)= \frac{1}{4} \int {\cos{\alpha}+3\alpha} = \frac{\sin{\alpha}}{4}+ \frac{3 \alpha^2}{8} $$
So that means (\ref{eq:1}) becomes:
$$\sin{\alpha}=F(3\alpha)+\frac{\sin{\alpha}}{4}+ \frac{3 \alpha^2}{8}$$

$$F(\alpha)=\frac{3\sin{(\frac{\alpha}{3})}}{4}- \frac{\alpha^2}{24}$$

Which means $u(x,y)=F(3x-y)+G(x+y)$ becomes

$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}- \frac{(3x-y)^2}{24}+  \frac{\sin{(x+y)}}{4}+ \frac{3(x+y)^2}{8}=$$

\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}+\frac{\sin{(x+y)}}{4}+ xy+  \frac{y^2}{3} $$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]

 $$u_y=\frac{-1}{4}\cos{\left(x-\frac{y}{3}\right)}+\frac{\cos{(x+y)}}{4}+ x+  \frac{2y}{3} $$
  $$u_{yy}=\frac{-1}{12}\sin{\left(x-\frac{y}{3}\right)}-\frac{\sin{(x+y)}}{4}+ \frac{2}{3} $$

$$u_x=\frac{3}{4}\cos{\left(x-\frac{y}{3}\right)}+\frac{\cos{(x+y)}}{4}+y$$

$$u_{xx}=\frac{-3}{4} \sin{\left( x-\frac{y}{3} \right)} - \frac{\sin{(x+y)}}{4}$$

$$u_{xy}=\frac{1}{4}\sin{\left(x-\frac{y}{3}\right)}-\frac{\sin{(x+y)}}{4}+1$$

Check that $u_{xx}+2u_{xy}-3u_{yy}=0$
$$
\left(\frac{-3}{4}+ \frac{2}{4}+\frac{1}{4}\right)\sin{\left(x-\frac{y}{3}\right)}+
\left(\frac{-1}{4}+ \frac{-2}{4}+\frac{3}{4}\right)\sin{(x+y)}+(0+2-3)=0
$$

\end{tcolorbox}

## Question 3

Solve the Neumann boundary value problem for the wave equation on half line:
\begin{equation*}
    \begin{cases}
      u_{tt}=c^2u_{xx}+f(t,x) & 0<x<\infty\\
      u(0,x)=\phi{x} & 0<x<\infty\\
      u_t(0,x) = \psi{x} & 0<x<\infty\\
      u_x(t,0) = h(t) & t>0
    \end{cases}       
\end{equation*}

## Question 4

Consider the 3D wave equation for $u(t,x,y,z)$:

$$u_{tt}=c^2\Delta u \quad\quad (x,y,z) \in \mathbb{R}^3, \quad t>0$$

Change the coordinates to spherical coordinates. Assume the solution is spherically symmetric, so that $u(t,x,y,z)=u(t,r)$ and does not depend on $\theta$ and $\phi$. Find the solution for $u(0,r)=0$ and

\begin{equation*} u_t(0,r) = \begin{cases}
  1  & \lvert{r}\rvert \leq 1 \\
  0 &  \lvert{r}\rvert > 1
\end{cases}
\end{equation*}

Hint: use the substitution $u(t,r)= \frac{1}{r} w(t,r)$.

First, we need to derive the formula for the Laplacian in spherical coordinates.

We know the equation for the Laplacian in polar coordinates is:

$$\frac{\partial^2}{\partial r^2}+\frac{1}{r}\frac{\partial}{\partial r}+ \frac{1}{r^2}\frac{\partial^2}{\partial \theta^2}$$

Now let's convert to spherical coordinates:

$$r=\sqrt{x^2+y^2+z^2}=\sqrt{s^2+z^2}$$

$$x=s \cos{\phi}, \quad y=s \sin{\phi}, \quad z=r \cos{\theta}$$
$$s=r \sin{\theta}$$

By the two-dimensional Laplacian, we have

$$u_{zz}+u_{ss}=u_{rr}+\frac{1}{r}u_r+\frac{1}{r^2}u_{\theta\theta},$$

$$u_{xx}+u_{yy}=u_{ss}+\frac{1}{s}u_s+\frac{1}{s^2}u_{\phi\phi}$$
We add these two equations and cancel $u_s$ to get

$$\Delta u = u_{rr}+\frac{1}{r}u_r+\frac{1}{r^2}u_{\theta\theta}+\frac{1}{s}u_s+\frac{1}{s^2}u_{\phi\phi}$$

Now since $u$ doesn't depend on $\theta$ or $\phi$, we have
$$\Delta u = u_{rr}+\frac{1}{r}u_r+\frac{1}{s}u_s=u_{rr}+\frac{1}{r}u_r+\frac{1}{r \sin{\theta}}u_s$$

$$ \frac{\partial u}{\partial s} = \frac{\partial u}{\partial r} \frac{\partial r}{\partial s}+\frac{\partial u}{\partial\theta}\frac{\partial\theta}{\partial s}+\frac{\partial u}
{\partial\phi}\frac{\partial\phi}{\partial s}=u_r\frac{1}{ \sin{\theta}}+0+0=u_r\frac{s}{ r}$$

So with our change of variables, we have

$$u_{tt}=c^2 \left(u_{rr} + \frac{2}{r}u_r \right)$$

Now set $w=ru$, or $u= \frac{w}{r}$. Then 
$$w_t=ru_t, \quad w_{tt}=ru_{tt}, \quad u_{tt}= \frac{w_{tt}}{r} $$
$$w_t=ru_t, \quad w_{tt}=ru_{tt}, \quad u_{tt}= \frac{w_{tt}}{r} $$
$$u_r = \frac{w_r}{r}-\frac{w}{r^2} $$

$$u_{rr} = \frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} $$
So $u_{tt}=c^2 \left(u_{rr} + \frac{2}{r}u_r \right)$ becomes

$$\frac{w_{tt}}{r} =c^2 \left(\frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} + \frac{2}{r} \left(\frac{w_r}{r}-\frac{w}{r^2} \right) \right),$$
which simplifies to
$$w_{tt}=c^2 w_{rr}.$$
But this is just the wave equation, and we can use d'Alembert's formula to find the solution:
 $$w(t,r)= \frac{\varphi(r+ct)+\varphi(r-ct)}{2}+\frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$
 Since $\varphi=0$, 
 $$w(t,r)= \frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$

Now we have 4 cases:

Case 1: $r-ct \geq -1, r+ct \leq 1$
$$w(t,r)= \frac{1}{2c} \int_{r-ct}^{r+ct}{s\;ds}$$

$$= \frac{1}{4c} ((r+ct)^2-(r-ct)^2)$$
$$= \frac{1}{4c}(r^2+2crt+c^2t^2-r^2+2crt-c^2t^2)= \frac{4crt}{4c}=rt  $$

Case 2:$r-ct < -1, r+ct>1$
$$w(t,r)= \frac{1}{2c} \int_{-1}^{1}{s\;ds}=$$
$$w(t,r)= \frac{1}{4c} (1-1)=0$$

Case 3:$r-ct < -1, r+ct \leq 1$
$$w(t,r)= \frac{1}{2c} \int_{-1}^{r+ct}{s\;ds}$$

$$= \frac{1}{4c} ((r+ct)^2-1)$$

Case 4:$r-ct \geq -1, r+ct>1$
$$w(t,r)= \frac{1}{2c} \int_{r-ct}^{1}{s\;ds}$$
$$= \frac{1}{4c} (1-(r-ct)^2)$$

Since $u=\frac{w}{r}$, this means we have
\begin{tcolorbox}[colback=white, title=Solution]
\begin{equation*}u(t,r)=
    \begin{cases} 
      t & \text{ if } r-ct \geq -1, r+ct \leq 1\\
      0 & \text{ if } r-ct < -1, r+ct > 1\\
      \frac{1}{4rc} ((r+ct)^2-1) & \text{ if } r-ct < -1, r+ct \leq 1\\
      \frac{1}{4rc} (1-(r-ct)^2) &\text{ if }  r-ct \geq -1, r+ct > 1\\
    \end{cases}       
\end{equation*}
\end{tcolorbox}

## Question 5
Consider the following Dirichlet boundary value problem:
\begin{equation*} 
\begin{cases}
  u_{tt}+x(t,x)u_t=u_{xx} & 0<x<1 \\
  u(0,x)=\phi(x) & 0<x<1 \\
  u_t(0,x)=\psi(x) & 0<x<1 \\
  u(t,0) = u(t,1)=0 & t \geq 0 \\
\end{cases}
\end{equation*}

Assume that $\lvert a(t,x)\rvert \leq m$ for some constant $m$ and all $0<x<1$ and $t\geq 0$. Let 
$$E(t)= \frac{1}{2} \int_{0}^{1}{ \left(u_t(t,x)^2+u_x(t,x)^2\right)}\:dx $$

(1) Show that 
\begin{equation}\label{eqn:E} 
\frac{dE(t)}{dt} \leq 2mE(t)
\end{equation}
for $t \geq 0$.

First differentiate $E(t)$:

$$ \frac{dE}{dt}= \frac{d}{dt} \left[\frac{1}{2} \int_{0}^{1}{ \left(u_t^2+u_x^2\right)}\:dx \right]$$ 
\begin{equation} \label{eqn:prob5a}
=\frac{1}{2} \int_{0}^{1}{ \frac{\partial}{\partial t} \left(u_t^2+u_x^2\right)} \:dx
\end{equation}

\begin{equation} \label{eqn:prob5b}
=\frac{1}{2} \int_{0}^{1}{ 2u_t u_{tt}+ 2u_x u_{xt}}\:dx
\end{equation}

\begin{equation}\label{eqn:prob5c}
=\int_{0}^{1}{ u_t u_{tt}}\: dx + \int_{0}^{1}{2u_x u_{xt}}\:dx =: I+J.
\end{equation}

The equality (\ref{eqn:prob5a}) follows from differentiation under the integral sign, while (\ref{eqn:prob5b}) follows from the chain rule for partial derivatives.

Consider the integral $J$ in (\ref{eqn:prob5c}). By integrating by parts, we can move one of the partials $\frac{\partial}{\partial x}$ to the other factor, at the cost of introducing a minus sign and a boundary term. Hence $J$ becomes

\begin{equation} \label{eqn:J}
J=\int_{0}^{1}{ u_x u_{xt}}\: dx =- \int_{0}^{1}{u_xx u_t}\:dx + 
\left[ u_x u_t \right]_{x=0}^{x=1}.
\end{equation}

The boundary term vanishes, since $u(t,0) \equiv u(t,1)  \equiv 0$ for $t>0$ implies that $u_t$ is identically zero at $x=0,1$. So, substituting (\ref{eqn:J}) for $J$, equation (\ref{eqn:prob5c}) becomes
 $$ I+J=\int_{0}^{1}{u_t u_{tt}} \;dx - \int_{0}^{1}{u_{xx} u_t} \;dx$$
 $$= \int_{0}^{1} {u_t (u_{tt}-u_{xx})} \;dx$$
\begin{equation} \label{eqn:J2}
=\int_{0}^{1}{ u_t(-au_t) }\:dx
\end{equation}

\begin{equation} \label{eqn:J3}
=\int_{0}^{1}{ (-a)u_t^2 }\: dx.
\end{equation}

Here, equality (\ref{eqn:J2}) just uses the PDE. Since $u_t^2 \geq 0$ and $-a \leq \lvert -a \rvert \leq m$, we see that $(-a)u_t^2 \leq mu_t^2$. Hence, the expression in (\ref{eqn:J3}) satisfies the inequality

\begin{equation} \label{eqn:IJ}
I+J= \int_{0}^{1}{ (-a)u_t^2 }\:dx \leq \int_{0}^{1}{ mu_t^2 }\:dx \leq m \int_{0}^{1}{ u_t^2 + u_x^2 }\:dx =2mE.
\end{equation}

Where we also used the fact that $u_x^2 \geq 0$ means that $m\int_{0}^{1}{ u_t^2 }\:dx \leq m\int_{0}^{1}{ (u_t^2 + u_x^2) }\:dx$.

The desired inequality (\ref{eqn:E}) follows from (\ref{eqn:IJ}) and the fact that $\frac{dE}{dt}=I+J.$

(2) Use part (1) and show that $\frac{d}{dt} \left( e^{-2mE(t)} \right) \leq 0$ for 
all $t \geq 0$. Hence, by integration from $[0,t]$, we get that

\begin{equation}\label{eqn:leq}
E(t) \leq e^{2mt} E(0) \text{ for all } t \geq 0.
\end{equation}

By the product rule, 
$$ \frac{d}{dt} \left(e^{-2mt}E\right)=-2me^{-2mt}E+e^{-2mt}\frac{dE}{dt} $$
$$=e^{-2mt}\left(\frac{dE}{dt}-2mE\right) $$
$$\leq e^{-2mt} \cdot 0=0. $$

(3) If $\phi(x)=\psi(x)=0$ for all $0<x<1$, what does this say about $E(t)$ for $t \geq 0$ and hence about $u(t,x)$ for $t \geq 0$?

Since $u(0,x) \equiv \varphi(x) \equiv 0$ for $0<x<1$, we see that $u_x$ is identically zero at time $t=0$. Similarly, $u_t(0,x) \equiv \psi(x) \equiv 0$ for $0<x<1$. Thus at $t=0$,

\begin{equation}\label{eqn:E0}
E(0)= \frac{1}{2} \int_{0}^{1}{ \left(u_t(t,0)^2+u_x(t,0)^2\right)}\:dx =\frac{1}{2} \int_{0}^{1}{0^2+0^2 \:dx=0.}
\end{equation}

But (\ref{eqn:E0}) together with the inequality (\ref{eqn:leq}) implies that

$$0 \leq E(t)  \leq e^{2mt}E(0) \leq 0$$
for all $t \geq 0$. Hence the energy $E$ is identically zero. Since the integrand $u_t^2+u_x^2$ is nonnegative, this is only possible if $u_t \equiv u_x \equiv 0$ for $t>0, 0<x<1$, meaning that $u$ varies with neither time nor position. But this implies that $u$ must be identically zero everywhere.

(4) Use the previous part to prove uniqueness of the following problem:
\begin{equation*} 
\begin{cases}
  u_{tt}+a(t,x)u_t=u_{xx} & 0<x<1, t>0 \\
  u(0,x)=\phi(x) & 0<x<1 \\
  u_t(0,x)=\psi(x) & 0<x<1 \\
  u(t,0) = f(t) & t \geq 0 \\
  u(t,1) = g(t) & t \geq 0 \\
\end{cases}
\end{equation*}

Let $u$ and $v$ be two solutions, and define $w:=u-v$. Observe that $w$ satisfies the original boundary value problem along with the conditions specified in part (3), implying that $w \equiv 0$.  Hence $u-v\equiv 0$ and so any solution $u$ must be unique. $\square$

## Problem 6

Does the D’Alembert method work for the wave equation $u_{tt} = c(x)^2u_{xx}$? What about $u_{tt} = c(t)^2u_{xx}$? Why?

Let's try the factorization method if $c=c(x)$:
$$\left( \partial_{t}+c(x)\partial_{x}\right) \left(\partial_{t}-c(x)\partial_{x}\right)u=0$$
Set $\xi=x+c(x)t$ and $\eta=x-c(x)t$.

By the chain rule,

$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}$$

$$=\partial_{\xi}(1+c')+\partial_{\eta}\,(1-c')$$

$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}$$
$$=c\partial_{\xi}-c\partial_{\eta}$$
When we plug these back into
$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
we get

$$ \left(c\partial_{\xi}-c\partial_{\eta}+c\partial_{\xi}(1+c')+c\partial_{\eta}\,(1-c')\right)
\left(c\partial_{\xi}-c\partial_{\eta} -c\partial_{\xi}(1+c')-c\partial_{\eta}\,(1-c')\right)u=0$$

When $c$ is constant, $c'=0$ and this simplifies to $-4c^2 u_{\xi\eta}=0$, allowing us to integrate to find $u(x,t)=F(x+ct)+G(x-ct)$. But we cannot do the same simplification when we have the $c'$ terms, meaning we cannot solve the wave equation by the same method when the wave speed is not constant.

Similarly,

Let's try the factorization method if $c=c(t)$:
$$\left( \partial_{t}+c(t)\partial_{x}\right) \left(\partial_{t}-c(t)\partial_{x}\right)u=0$$
Set $\xi=x+c(t)t$ and $\eta=x-c(t)t$.

By the chain rule,

$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}=
\partial_{\xi}+\partial_{\eta}$$

$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}=\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)
$$

When we plug these back into
$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
we get

$$\left( \partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)+c\partial_{\xi}+c\partial_{\eta} \right)
\left(\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)-c\partial_{\xi}-c\partial_{\eta}\right)u=0$$
Where again we cannot cancel terms to simplify our equation as we could with constant $c$.

## Problem 7 (The Poisson-Darboux Equation)   
Solve the initial value problem

\begin{equation*} 
\begin{cases}
  u_{tt}-u_{xx}- \frac{2}{x}u_x=0 & -\infty <x<\infty, t>0 \\
  u(0,x)=0 & -\infty <x<\infty \\
  u_t(0,x)=g(x) & -\infty <x< \infty \\
\end{cases}
\end{equation*}

where $g(x)=g(-x)$ is an even function. Hint: set $w=xu$.

Using the results from when we set $w=ru$ in Problem 4,
$$u_t=\frac{w_t}{x}, u_{tt}=\frac{w_{tt}}{x} $$
$$u_x= \frac{w_x}{x}- \frac{w}{x^2} $$

$$u_{xx}=  \frac{w_{xx}}{x}-\frac{2w_x}{x^2}+\frac{2w}{x^3}$$
So $u_{tt}-u_{xx}- \frac{2}{x}u_x=0$ becomes

$$\frac{w_{tt}}{x} -\frac{w_{xx}}{x}+\frac{2w_x}{x^2}- \frac{2w}{x^3}- \frac{2}{x} \left(\frac{w_x}{x}- \frac{w}{x^2}\right)=0 $$
Which simplifies to $w_{tt}-w_{xx}=0$, and so $w$ solves the wave equation when $-\infty <x<\infty, t>0$.

Now observe that $w(0,x)=xu(0,x)=0$ and $w_t(0,x)=xu_t(0,x)=xg(x)$. Hence, $w$ solves the initial value problem:

\begin{equation} \label{eqn:prob7}
    \begin{cases}
      w_{tt}-w_{xx} & -\infty <x<\infty, t>0 \\
      w(0,x)=0 & -\infty <x<\infty, t>0 \\
      w_t(0,x) = xg(x) & -\infty <x<\infty, t>0
    \end{cases}       
\end{equation}

whre $g(x)$ is an even function.

By d'Alembert's formula, the solution to (\ref{eqn:prob7}) is given by:
$$w(t,x)=\frac{1}{2} \int_{x-t}^{x+t}{sg(s)ds}$$
Hence since $w_t=xu_t$, 
\begin{equation} \label{eqn:prob7a}
w(t,x)=\frac{1}{2x} \int_{x-t}^{x+t}{sg(s)ds,}
\end{equation}
assuming $x \neq 0$.

To handle the case when $x=0$, assume that $u$ is continuous in $x$, so that $u(t,0)=\lim_{x\to 0} u(t,x)$. Define

$$I(y)=\int_{0}^{y}sg(s)\;ds. $$

This allows us to rewrite (\ref{eqn:prob7a}) as

$$u(t,x)=\frac{I(x+t)-I(x-t)}{2x}= \frac{I(x+t)-I(t-x)}{2x},$$
Noting that the evenness of $g(s)$ means that $sg(s)$ is odd, and so $I(y)$ is even (one can check by substitution that $I(y)=I(-y)$). Hence, letting $x \rightarrow 0$, we see that

\begin{equation}\label{eqn:prob7b}
u(t,0)= \lim_{x\to 0} \frac{I(t+x)-I(t-x)}{2x}.
\end{equation}

But this is just the form of a derivative, namely the symmetric derivative of $I$ at $t$. By the Fundamental Theorem of Calculus, assuming $g$ is continuous, we know the limit (\ref{eqn:prob7b}) exists and is necessarily equal to $I'(t)=tg(t)$.

Thus $u(t,x)$ is given for any $t \geq 0, x \in \mathbb{R}$ by

\begin{equation} \label{eqn:prob7c} 
u(t,x)=
\begin{cases}
      \frac{1}{2x} \int_{x-t}^{x+t}{sg(s)ds,} & x \neq 0, \\
      tg(t), & x=0 \\
    \end{cases} 
\end{equation}

## Problem 8   
Solve the following characteristic initial value problem:
\begin{equation*} 
\begin{cases}
  y^3u_{xx}-yu_{yy}+u_y=0 & 0<x<4, \quad \lvert y \rvert \leq 2 \sqrt{2} \\
  u(x,y)= f(x) & x+ \frac{y^2}{2}=4 \text{ for } 2 \leq x \leq 4 \\
  u(x,y)=g(x) & x- \frac{y^2}{2}=0 \text{ for } 0 \leq x \leq 2 \\
\end{cases}
\end{equation*}

where $f(2)=g(2)$. Hint: Use the transformation $\eta=x-\frac{y^2}{2}$ and $\xi=x+\frac{y^2}{2}$ and express the PDE in the coordinates $(\xi, \eta)$.

Set $\eta=x-\frac{y^2}{2}$ and $\xi=x+\frac{y^2}{2}$. Then

$$\frac{\partial u}{\partial x} = \frac{\partial u}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial u}{\partial\eta}\frac{\partial\eta}{\partial x} = \frac{\partial u}{\partial \xi}+\frac{\partial u}{\partial\eta}$$

$$\frac{\partial^2 u}{\partial x^2} = \frac{\partial }{\partial x} \left(\frac{\partial u}{\partial \xi}\right) +\frac{\partial }{\partial x} \left(\frac{\partial u}{\partial \eta}\right)$$

$$=\frac{\partial^2 u}{\partial \xi^2} \frac{\partial \xi}{\partial x}+
\frac{\partial^2 u}{\partial\eta \partial \xi}\frac{\partial \xi}{\partial x} +
\frac{\partial^2 u}{\partial\eta \partial \xi}\frac{\partial\eta}{\partial x}+
\frac{\partial^2 u}{\partial \eta^2}\frac{\partial \eta}{\partial x}  $$

Which we can simplify to

$$u_{xx}= u_{\xi \xi}+2 u_{\xi \eta}+u_{\eta \eta}$$

Now we do the same for $y$:

$$\frac{\partial u}{\partial y} = \frac{\partial u}{\partial \xi} \frac{\partial \xi}{\partial y}+\frac{\partial u}{\partial\eta}\frac{\partial\eta}{\partial y} = y\frac{\partial u}{\partial \xi}-y\frac{\partial u}{\partial\eta}$$

But since $\eta=x-\frac{y^2}{2}$ and $\xi=x+\frac{y^2}{2}$, we can rewrite $y$ as $\sqrt{\xi-\eta}$.

$$\frac{\partial u}{\partial y} = \sqrt{\xi-\eta} \left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)$$

Now using the product rule,
$$\frac{\partial^2 u}{\partial y^2} = \sqrt{\xi-\eta}\; \frac{\partial}{\partial y}\left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)+\left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)$$

Where the second term has been simplified because $\frac{\partial}{\partial y}$ of $\sqrt{\xi-\eta}$ is simply $\frac{\partial}{\partial y}y=1$.

Now we use the chain rule again for the first term:

Now using the product rule,
$$\frac{\partial^2 u}{\partial y^2} = \sqrt{\xi-\eta}\left(\frac{\partial^2 u}{\partial \xi^2} \frac{\partial \xi}{\partial y}+
\frac{\partial^2 u}{\partial\eta \partial \xi}\frac{\partial\eta}{\partial y} -
\frac{\partial^2 u}{\partial \eta^2}\frac{\partial \eta}{\partial y}-
\frac{\partial^2 u}{\partial \eta \partial \xi}\frac{\partial \xi}{\partial y}\right)
+
\left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)$$

Which simplifies to

$$u_{yy}= \left(\xi-\eta \right)\left(u_{\xi \xi}+u_{\eta \eta} -2 u_{\xi \eta}\right) +u_\xi-u_\eta$$

This means $y^3u_{xx}-yu_{yy}+u_y=0$ becomes

$$\left(\xi-\eta \right)^{\frac{3}{2}} \left(u_{\xi \xi}+2 u_{\xi \eta}+u_{\eta \eta}\right) - \left(\xi-\eta \right)^{\frac{1}{2}} \left[\left(\xi-\eta \right)\left(u_{\xi \xi}-2 u_{\xi \eta}+u_{\eta \eta} \right) +u_\xi-u_\eta \right]+\left(\xi-\eta \right)^{\frac{1}{2}} \left(u_\xi-u_\eta\right)=0  $$

$$\left(\xi-\eta \right)^{\frac{3}{2}} \left(u_{\xi \xi}+2 u_{\xi \eta}+
u_{\eta \eta}\right) - \left(\xi-\eta \right)^{\frac{3}{2}} \left(u_{\xi \xi}-2 u_{\xi \eta}+
u_{\eta \eta}\right)-
\left(\xi-\eta \right)^{\frac{1}{2}} \left(u_\xi-u_\eta\right)+
\left(\xi-\eta \right)^{\frac{1}{2}} \left(u_\xi-u_\eta\right)=0  $$

$$\left(\xi-\eta \right)^{\frac{3}{2}} \left(4\, u_{\xi \eta}\right)=0$$

But by integrating with respect to both variables, we recognize $u_{\xi \eta}=0$ as having the solution

$$u(\xi, \eta)=h_1(\xi)+h_2(\eta),$$

or in our original coordinates,

$$u(x,y)=h_1\left(x+\frac{y^2}{2}\right)+h_2\left(x-\frac{y^2}{2}\right) $$

Now we use our given conditions. When  $2 \leq x \leq 4$, we have $g(x)=h_1(2x)+h_2(0)$, which means $h_1(z_1)=g\left(\frac{z_1}{2}\right)-h_2(0)$.

Similarly, when $0 \leq x \leq 2$, we have $f(x)=h_1(4)+h_2(2x-4)$, which means $h_2(z_2)=f\left(\frac{z_2}{2}+2\right)-h_1(4)$.

Thus,
$$h_1(z_1)+h_2(z_2)=g\left(\frac{z_1}{2}\right)-h_2(0)+f\left(\frac{z_2}{2} + 2\right)-h_1(4)$$

But $h_2(0)+h_1(4)$ is just $g(x)$ when $x=2$ (or equivalently, $f(x)$ when $x=2$),

so after replacing $z_1$ by $x+\frac{y^2}{2}$ and $z_2$ by $x-\frac{y^2}{2}$, we get

\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=g\left(\frac{x}{2}+\frac{y^2}{4}\right)+f\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f(2)$$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$u_y(x,y)=\frac{y}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{y}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)$$

$$u_{yy}(x,y)=\frac{y^2}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+
\frac{1}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+
\frac{y^2}{4}f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)+
\frac{1}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)$$

$$u_x(x,y)=\frac{1}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f'(2)$$

$$u_{xx}(x,y)=\frac{1}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{4}f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f''(2)$$

This means $y^3u_{xx}-yu_{yy}+u_y=0$ becomes

\begin{equation*}
  \begin{aligned}
  & \left[\frac{y^3}{4} g''\left(\frac{x}{2} +\frac{y^2}{4}\right) + \frac{y^3}{4} f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)\right]- \\
  & \left[\frac{y^3}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+
\frac{y}{2} g' \left(\frac{x}{2}+\frac{y^2}{4} \right)+
\frac{y^3}{4} f'' \left( \frac{x}{2}+ \frac{y^2}{4}+2 \right)+
\frac{y}{2}f' \left( \frac{x}{2}+ \frac{y^2}{4}+2 \right) \right] + \\
  & \left[ \frac{y}{2} g' \left( \frac{x}{2}+ \frac{y^2}{4} \right)+ \frac{y}{2}f' \left( \frac{x}{2}+ \frac{y^2}{4}+2 \right) \right]=0
 \end{aligned}
\end{equation*}

\end{tcolorbox}