### 2024-02-23 PDEs/math245_hw3.Rmd
started pde hw3

---
title: "MATH 245 Homework 3"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-23"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1  (Parity of Solution)
Prove the following.

**(a)** If initial conditions of the wave equation on the whole line are even(odd), the solution is
even(odd).
\newline

  Consider the IVP
\begin{equation*}
    \begin{cases}
      u_{tt}=c^2u_{xx}\ & -\infty <x<\infty\\
      u(0,x)=\varphi(x) & -\infty<x<\infty\\
      u_t(0,x) = \psi(x) & -\infty<x<\infty\\
    \end{cases}       
\end{equation*}

where $\varphi(x)$ and $\psi(x)$ are even functions, i.e., $\varphi(-x)=\varphi(x)$ and $\psi(-x)=\psi(x)$. Then using D'Alembert's formula, we know the solution $u$ is

$$u(x,t)= \frac{\varphi(x+ct)+\varphi(x-ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(s)ds}$$
Now plug in $-x$:
$$u(-x,t)= \frac{\varphi(-x+ct)+\varphi(-x-ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(s)ds}$$
$$ = \frac{\varphi(x-ct)+\varphi(x+ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(-s)ds}$$
After setting $y=-s, dy=-ds$ in the integral of $\psi$:

$$ u(-x,t)= \frac{\varphi(x-ct)+\varphi(x+ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(y)dy}=u(x,t).$$
Therefore, even initial conditions imply that the solution of the wave equation is even.

Similarly, assume $\varphi(x)$ and $\psi(x)$ are odd functions, i.e., $\varphi(-x)=-\varphi(x)$ and $\psi(-x)=-\psi(x)$. Then using D'Alembert's formula, we know $u(-x, t)$ is

$$u(-x,t)= \frac{\varphi(-x+ct)+\varphi(-x-ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(s)ds}$$
$$= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(-s)ds}$$

After setting $y=-s, dy=-ds$ in the integral of $\psi$:
$$u(-x,t)= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(-y)dy}=-u(x,t).$$
Therefore, odd initial conditions imply that the solution of the wave equation is odd.

**(b)** If the initial condition of the heat equation on the whole line is even(odd), the solution is even(odd).

## Question 2  (Speed of Heat vs Wave)
Consider the traveling wave $u(x, t) = f(x-at)$ where $f$ is a given function of one variable.

(a) If it is a solution of the wave equation, show that the speed must be $a =\pm c$ (unless $f$ is a
linear function).

(b) If it is a solution of the diffusion equation, find $f$ and show that the speed $a$ is arbitrary.

## Question 4  (Maximum Principle)

Consider two solutions $u(t,x)$ and $v(t,x)$ of the diffusion equation in $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$

(a) Let $M(T)=$ the maximum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$. Does $M(T)$ increase or decrease as a function of $T$? Explain.

(b) Let $m(T)=$ the minimum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$. Does $m(T)$ increase or decrease as a function of $T$? Explain.

(c) **Comparison Principle**
If $u \leq v$ for $t=0,$ for $x=0, $ for $x=l$, then $u \leq v$ for $0 \leq t \leq \infty$ and $0 \leq x \leq l$.

## Question 5 (Diffusion Equation with Dissipation). 
Solve the following IVP for constant dissipation $b>0$.

\begin{equation}\label{eqn:dissipation}
\begin{cases}
u_t-ku_{xx}+bu=0 & -\infty \leq x \leq \infty, \; t>0 \\
u(0,x)=\psi(x), & -\infty \leq x \leq \infty

\end{cases}
\end{equation}

We will use the change of variables $u(t,x)=e^{-bt}v(t,x)$

## Question 6
Prove that there is no maximum principle for the wave equation.