### 2024-02-23 PDEs/math245_hw3.Rmd
pde problem 6

---
title: "MATH 245 Homework 3"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-23"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
##Done: 1, 6
##Not started: 2-5

## Question 1  (Parity of Solution)

**(a)** Prove that if initial conditions of the wave equation on the whole line are even(odd), the solution is
even(odd).
\newline
\newline
Consider the IVP
\begin{equation*}
    \begin{cases}
      u_{tt}=c^2u_{xx}\ & -\infty <x<\infty\\
      u(0,x)=\varphi(x) & -\infty<x<\infty\\
      u_t(0,x) = \psi(x) & -\infty<x<\infty\\
    \end{cases}       
\end{equation*}

where $\varphi(x)$ and $\psi(x)$ are even functions, i.e., $\varphi(-x)=\varphi(x)$ and $\psi(-x)=\psi(x)$. Then using D'Alembert's formula, we know the solution $u$ is

$$u(x,t)= \frac{\varphi(x+ct)+\varphi(x-ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(s)ds}$$
Now plug in $-x$:
$$u(-x,t)= \frac{\varphi(-x+ct)+\varphi(-x-ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(s)ds}$$
$$= \frac{\varphi(x-ct)+\varphi(x+ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(-s)ds}$$
After setting $y=-s, dy=-ds$ in the integral of $\psi$:

$$ u(-x,t)= \frac{\varphi(x-ct)+\varphi(x+ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(y)dy}=u(x,t).$$
Therefore, even initial conditions imply that the solution of the wave equation is even.

Similarly, assume $\varphi(x)$ and $\psi(x)$ are odd functions, i.e., $\varphi(-x)=-\varphi(x)$ and $\psi(-x)=-\psi(x)$. Then using D'Alembert's formula, we know $u(-x, t)$ is

$$u(-x,t)= \frac{\varphi(-x+ct)+\varphi(-x-ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(s)ds}$$
$$= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(-s)ds}$$

After setting $y=-s, dy=-ds$ in the integral of $\psi$:
$$u(-x,t)= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(-y)dy}=-u(x,t).$$
Therefore, odd initial conditions imply that the solution of the wave equation is odd.

**(b)** Prove that if the initial condition of the heat equation on the whole line is even(odd), the solution is even(odd).

  Consider the IVP
\begin{equation*}
    \begin{cases}
      u_{t}=k u_{xx} & -\infty <x<\infty\\
      u(0,x)=f(x)\\
    \end{cases}       
\end{equation*}

Then we know

$$u(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy $$
Which means, when $x=-x$:
$$u(-x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(-x-y)^2/4kt}f(y)\;dy $$
Now use the change of variable $y=-s, dy=-ds$:
$$u(-x,t)= \frac{-1}{\sqrt{4 \pi kt}} \int_{\infty}^{-\infty} e^{-(-x+s)^2/4kt}f(-s)\;ds $$

$$= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-((-1)(x-s))^2/4kt}f(-s)\;ds $$
But $e^{-((-1)(x-s))^2/4kt}$ is equivalent to $e^{-(x-s)^2/4kt}$, so we have

$$u(-x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds $$

Now assume $f$ is odd. By definition we have $f(-s)=-f(s)$, so

$$u(-x,t)=\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds = 
\frac{-1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(s)\;ds = -u(x,t) $$

And therefore our solution is odd when our initial $f(x)$ is an odd function. Similarly, if we assume $f(x)$ is even, by definition we have $f(-s)=f(s)$, so
$$u(-x,t)=\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds = 
\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(s)\;ds = u(x,t), $$
meaning our solution is also even.

## Question 2  (Speed of Heat vs Wave)
Consider the traveling wave $u(x, t) = f(x-at)$ where $f$ is a given function of one variable.

(a) If it is a solution of the wave equation, show that the speed must be $a =\pm c$ (unless $f$ is a
linear function).

(b) If it is a solution of the diffusion equation, find $f$ and show that the speed $a$ is arbitrary.

## Question 3 (Heat Equation with Robin boundary)
Consider the following problem with a Robin boundary condition:
\begin{equation*}
    \begin{cases}
      u_{t}=k u_{xx} & 0 <x<\infty,0 <t<\infty \\
      u(0,x)=x & t=0,0 <x<\infty \\
      u_x(0,t)-2u(0,t)=0 & x=0
    \end{cases}       
\end{equation*}

## Question 4  (Maximum Principle)

Consider two solutions $u(t,x)$ and $v(t,x)$ of the diffusion equation in $\{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$

(a) Let $M(T)=$ the maximum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq T\}$. Does $M(T)$ increase or decrease as a function of $T$?

The maximum principle says that if $u(x, t)$ satisfies the diffusion equation in a rectangle $\{ 0 \leq x \leq x, 0 \leq t \leq T \}$, then the maximum value of $u(x, t)$ is assumed either initially $(t = 0)$ or on the lateral sides $(x = 0$ or $x = l)$.

By the definition of maximum, if $u$ assumes a higher value on the lateral sides $x = 0$ or $x = l$ when $T$ is extended from some $T_1$ to a new $T$, $T_2 > T_1$, then the maximum of $u$ will increase to that new value; otherwise, the maximum of $u$ will remain the same. In other words, $M$ is non-decreasing (increases, but not strictly) as a function of $T$.

(b) Let $m(T)=$ the minimum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq T \}$. Does $m(T)$ increase or decrease as a function of $T$?

By the definition of minimum, if $u$ assumes a lower value on the lateral sides $x = 0$ or $x = l$ when $T$ is extended from some $T_1$ to a new $T$, $T_2 > T_1$, then the minimum of $u$ will decrease to that new value; otherwise, the minimum of $u$ will remain the same. In other words, $M$ is non-increasing (decreases, but not strictly) as a function of $T$.

(c) Comparison Principle: Show that if $u \leq v$ for $t=0,x=0,$ and $x=l$, then $u \leq v$ for $0 \leq t \leq \infty$ and $0 \leq x \leq l$.

## Question 5 (Diffusion Equation with Dissipation)
Solve the following IVP for constant dissipation $b>0$.

\begin{equation}\label{eqn:dissipation}
\begin{cases}
u_t-ku_{xx}+bu=0 & -\infty \leq x \leq \infty, \; t>0 \\
u(0,x)=\psi(x), & -\infty \leq x \leq \infty

\end{cases}
\end{equation}

We will use the change of variables $u(t,x)=e^{-bt}v(t,x)$

## Question 6

Suppose that there is a maximum principle for the wave equation and that $u(x,t)$ is a solution to the wave equation in the closed rectangle $\Omega_T = \{ 0 \leq x \leq \pi, \; 0 \leq t \leq \pi/c \}$. Then the maximum of $u$ is assumed either initially (at $t=0$) or on the lines $x=0$ or $x=\pi$.

Consider the particular solution $u(x, t) = \sin x \sin ct$. This clearly satisfies the wave equation:
$$u_x=\cos x \sin ct $$
$$u_{xx}=-\sin x \sin ct $$
$$u_t=\sin x (c \cos ct) $$
$$u_{tt}=-\sin x (c^2 \sin ct)= c^2 u_{xx}$$

According to our presumed maximum principle, the maximum of $u(x,t)$ would have to be one of the following:
$$u(x,0)= \sin x \sin 0=0 $$
$$u(0, t)=\sin 0 \sin ct = 0 $$
$$u(\pi, t)=\sin \pi \sin ct = 0 $$

But $u(x,t)$ has a higher value when $x=\pi/2$ and $t=\pi/2c$:  
$$u\left(\frac{\pi}{2}, \frac{\pi}{2c}\right)=\sin \left(\frac{\pi}{2}\right) \sin \left(\frac{\pi}{2}\right)=1$$

This contradicts the definition of maximum, so we have found a counterexample and therefore there is no maximum principle for the wave equation.