### 2024-02-23 PDEs/math245_hw3.Rmd
pde hw3 mostly done

---
title: "MATH 245 Homework 3"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-23"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
##Done: 1, 2, 4, 5, 6.

##Not Done: 3

## Question 1  (Parity of Solution)

**(a)** Prove that if initial conditions of the wave equation on the whole line are even(odd), the solution is
even(odd).
\newline
\newline
Consider the IVP
\begin{equation*}
    \begin{cases}
      u_{tt}=c^2u_{xx}\ & -\infty <x<\infty\\
      u(0,x)=\varphi(x) & -\infty<x<\infty\\
      u_t(0,x) = \psi(x) & -\infty<x<\infty\\
    \end{cases}       
\end{equation*}

where $\varphi(x)$ and $\psi(x)$ are even functions, i.e., $\varphi(-x)=\varphi(x)$ and $\psi(-x)=\psi(x)$. Then using D'Alembert's formula, we know the solution $u$ is

$$u(x,t)= \frac{\varphi(x+ct)+\varphi(x-ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(s)ds}$$
Now plug in $-x$:
$$u(-x,t)= \frac{\varphi(-x+ct)+\varphi(-x-ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(s)ds}$$
$$= \frac{\varphi(x-ct)+\varphi(x+ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(-s)ds}$$
After setting $y=-s, dy=-ds$ in the integral of $\psi$:

$$ u(-x,t)= \frac{\varphi(x-ct)+\varphi(x+ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(y)dy}=u(x,t).$$
Therefore, even initial conditions imply that the solution of the wave equation is even.

Similarly, assume $\varphi(x)$ and $\psi(x)$ are odd functions, i.e., $\varphi(-x)=-\varphi(x)$ and $\psi(-x)=-\psi(x)$. Then using D'Alembert's formula, we know $u(-x, t)$ is

$$u(-x,t)= \frac{\varphi(-x+ct)+\varphi(-x-ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(s)ds}$$
$$= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(-s)ds}$$

After setting $y=-s, dy=-ds$ in the integral of $\psi$:
$$u(-x,t)= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(-y)dy}=-u(x,t).$$
Therefore, using the uniqueness of the wave equation, odd initial conditions imply that the solution of the wave equation is odd.

**(b)** Prove that if the initial condition of the heat equation on the whole line is even(odd), the solution is even(odd).

  Consider the IVP
\begin{equation*}
    \begin{cases}
      u_{t}=k u_{xx} & -\infty <x<\infty\\
      u(0,x)=f(x)\\
    \end{cases}       
\end{equation*}

Then we know

$$u(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy $$
Which means, when $x=-x$:

$$u(-x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(-x-y)^2/4kt}f(y)\;dy $$
Now use the change of variable $y=-s, dy=-ds$:

$$u(-x,t)= \frac{-1}{\sqrt{4 \pi kt}} \int_{\infty}^{-\infty} e^{-(-x+s)^2/4kt}f(-s)\;ds $$

$$= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-((-1)(x-s))^2/4kt}f(-s)\;ds $$
But $e^{-((-1)(x-s))^2/4kt}$ is equivalent to $e^{-(x-s)^2/4kt}$, so we have

$$u(-x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds $$

Now assume $f$ is odd. By definition, we have $f(-s)=-f(s)$, so

$$u(-x,t)=\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds = 
\frac{-1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(s)\;ds = -u(x,t) $$

And therefore our solution is odd when our initial $f(x)$ is an odd function. Similarly, if we assume $f(x)$ is even, by definition we have $f(-s)=f(s)$, so
$$u(-x,t)=\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds = 
\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(s)\;ds = u(x,t), $$
meaning (using the uniqueness of the heat equation), our solution is also even.

## Question 2  (Speed of Heat vs Wave)
Consider the traveling wave $u(x, t) = f(x-at)$ where $f$ is a given function of one variable.

(a) If it is a solution of the wave equation, show that the speed must be $a =\pm c$ (unless $f$ is a linear function).

\begin{equation}\label{eqn:tderiv}
u_t = -af'(x-at), \quad u_{tt}=a^2 f''(x-at)
\end{equation}

\begin{equation}\label{eqn:xderiv}
u_x = f'(x-at), \quad u_{xx}=f''(x-at)
\end{equation}

If $u(x, t) = f(x-at)$ is a solution to the wave equation, then we must have $u_{tt}=c^2 u_{xx}$. By substituting (\ref{eqn:tderiv}) and (\ref{eqn:xderiv}) into the PDE, we get

$$a^2 f''=c^2f'' ,$$
which implies at $a^2=c^2$, or $a= \pm c$. However, if $f$ is a linear function, then the second derivative $f''=0$ and we cannot conclude that $a^2=c^2$.

(b) If it is a solution of the diffusion equation, find $f$ and show that the speed $a$ is arbitrary.

If $u(x, t) = f(x-at)$ is a solution to the wave equation, then we must have $u_{t}=k u_{xx}$. By substituting (\ref{eqn:tderiv}) and (\ref{eqn:xderiv}) into the PDE, we get

$$-af'=kf'' $$
$$-af'=kf'' $$
$$ \frac{-a}{k}= \frac{f''}{f'} $$
Now if we set $\xi=x-at$, 
$$\frac{-a}{k}=\frac{f''}{f'}= \frac{d}{d\xi} \ln(f') $$
So integrating with respect to $\xi$,
$$\frac{-a}{k}\xi+C= \ln (f')$$
$$ f'(\xi) = e^{ \frac{-a}{k} \xi+C_1} = C_2 e^{\frac{-a}{k} \xi} $$
Now if we integrate again with respect to $\xi$,
$$f(\xi)=\frac{-k}{a}C_2 e^{\frac{-a}{k}\xi}+C_3 $$

Set a new constant $C_4=\frac{-k}{a}C_3$,
$$f(\xi)=C_4 e^{\frac{-a}{k}\xi}+C_3 $$

If we change back to the original variables,
$$f(x-at)=C_4 e^ \frac{-a(x-at)}{k}+C_3 $$

Thus, we found the solution
\begin{equation} \label{eqn:findf}
u(x, t) = f(x-at)=A e^{\frac{-a}{k}(x-at)}+B,
\end{equation}
where $A$ and $B$ are constants.

We can check that this satisfies the heat equation:
$$u_t=A\frac{a^2}{k} e^ {\frac{-a}{k}(x-at)}$$

$$u_x=A\frac{-a}{k} e^ {\frac{-a}{k}(x-at)}$$

$$u_{xx}=A\frac{a^2}{k^2} e^ {\frac{-a}{k}(x-at)}$$
$$u_t-ku_{xx}=A\frac{a^2}{k} e^ {\frac{-a}{k}(x-at)}-kA\frac{a^2}{k^2} e^ {\frac{-a}{k}(x-at)}=0$$
However, (\ref{eqn:findf}) will be a solution regardless of the value of $a$; i.e., $a$ is arbitrary.

## Question 3 (Heat Equation with Robin boundary)
Consider the following problem with a Robin boundary condition:
\begin{equation*} (\ast)
    \begin{cases}
      u_{t}=k u_{xx} & 0 <x<\infty,0 <t<\infty \\
      u(0,x)=x & t=0,0 <x<\infty \\
      u_x(0,t)-2u(0,t)=0 & x=0
    \end{cases}       
\end{equation*}

The purpose of this exercise is to verify the solution formula for $(\ast)$. Let $f(x)=x$ for $x>0$ and let $f(x)=x+1-e^{2x}$ for $x<0$, and let

$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy. $$
**(a)** What PDE and initial conditions does $v(x,t)$ satisfy for $-\infty < x < \infty$?

$v(x,t)$ satisfies

\begin{equation*}
    \begin{cases}
      u_{t}=k u_{xx} & 0 <x<\infty,0 <t<\infty \\
      u(0,x)=x & t=0,0 <x<\infty \\
    \end{cases}       
\end{equation*}

**(b)** Let $w=v_x-2v$. What PDE and initial condition does $w(x,y)$ satisfy for $-\infty < x < \infty$?

**(c)** Show that $f'(x)-2f(x)$ is an odd function for $x \neq 0$.
\begin{equation*}  f(x)=
\begin{cases}
   x & \text{ if } x >0 \\
   x+1-e^{2x} & \text{ if } x<0\\
\end{cases}
\end{equation*}

\begin{equation*}  f'(x)=
\begin{cases}
   1 & \text{ if } x >0 \\
   1-2e^{2x} & \text{ if } x<0\\
\end{cases}
\end{equation*}

\begin{equation*}  f'(x)-2f(x)=
\begin{cases}
   1-2x & \text{ if } x >0 \\
   -1-2x & \text{ if } x<0\\
\end{cases}
\end{equation*}

Which means that
\begin{equation*}  f'(-x)-2f(-x)=
\begin{cases}
   1-2(-x) & \text{ if } (-x) >0 \\
   -1-2(-x) & \text{ if } (-x) <0\\
\end{cases}
\end{equation*}

But this is equivalent to
\begin{equation*}
f'(-x)-2f(-x)=
\begin{cases}
   1+2x & \text{ if } x <0 \\
    -1+2x & \text{ if } x>0\\
\end{cases}
\end{equation*}

which is exactly $-(f'(x)-2f(x))$. Therefore, $f'(x)-2f(x)$ is an odd function of $x$.

**(d)** Use Exercise 2.4.11 to show that $w$ is an odd function of $x$.

**(e)** Deduce that $v(x, t)$ satisfies $(\ast)$ for $x>0$. Assuming uniqueness, deduce that the solution of $(\ast)$ is given by
$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy. $$

## Question 4  (Maximum Principle)

Consider two solutions $u(t,x)$ and $v(t,x)$ of the diffusion equation in $\{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$

(a) Let $M(T)=$ the maximum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq T\}$. Does $M(T)$ increase or decrease as a function of $T$?

The maximum principle says that if $u(x, t)$ satisfies the diffusion equation in a rectangle $\{ 0 \leq x \leq x, 0 \leq t \leq T \}$, then the maximum value of $u(x, t)$ is assumed either initially $(t = 0)$ or on the lateral sides $(x = 0$ or $x = l)$.

By the definition of maximum, if $u$ assumes a higher value on the lateral sides $x = 0$ or $x = l$ when $T$ is extended from some $T_1$ to a new $T$, $T_2 > T_1$, then the maximum of $u$ will increase to that new value; otherwise, the maximum of $u$ will remain the same. In other words, $M$ is non-decreasing (increases, but not strictly) as a function of $T$.

(b) Let $m(T)=$ the minimum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq T \}$. Does $m(T)$ increase or decrease as a function of $T$?

By the definition of minimum, if $u$ assumes a lower value on the lateral sides $x = 0$ or $x = l$ when $T$ is extended from some $T_1$ to a new $T$, $T_2 > T_1$, then the minimum of $u$ will decrease to that new value; otherwise, the minimum of $u$ will remain the same. In other words, $M$ is non-increasing (decreases, but not strictly) as a function of $T$.

(c) Comparison Principle: Show that if $u \leq v$ for $t=0,x=0,$ and $x=l$, then $u \leq v$ for $0 \leq t \leq \infty$ and $0 \leq x \leq l$.

Define $w(t,x)=u(t,x)-v(t,x)$ and note that $w$ solves the heat equation in $\{ 0 \leq x \leq l, 0 \leq t \leq T \}$:
$$w_t-kw_{xx}=u_t-v_t-k(u_{xx}-v_{xx})=(u_t-ku_{xx})-(v_t-kv_{xx})=0-0=0 $$
Since we are assuming $u \leq v$ for $t=0,x=0,$ and $x=l$, then $w \leq 0$ for $t=0,x=0,$ and $x=l$. But by the Maximum Principle, $w$ assumes its maximum value on one of those three lines. This means we can take $T$ to be arbitrarily large and still have $w \leq 0$, i.e., $w \leq 0$ on $\{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$ and therefore $u \leq v$ for $0 \leq t \leq \infty$ and $0 \leq x \leq l$.

## Question 5 (Diffusion Equation with Dissipation)
Solve the following IVP for constant dissipation $b>0$.

\begin{equation}\label{eqn:dissipation}
\begin{cases}
u_t-ku_{xx}+bu=0 & -\infty \leq x \leq \infty, \; t>0 \\
u(0,x)=\psi(x), & -\infty \leq x \leq \infty

\end{cases}
\end{equation}

We will use the change of variables $u(t,x)=e^{-bt}v(t,x)$, and differentiate to find that:

$$u_t=e^{-bt}v_t-be^{-bt}v$$
$$u_x=e^{-bt}v_x$$
$$u_{xx}=e^{-bt}v_{xx}$$

So $u_t-ku_{xx}+bu=0$ becomes

$$e^{-bt}v_t-be^{-bt}v-k \left(e^{-bt}v_{xx}\right) + be^{-bt}v=0 $$
After canceling the $be^{-bt}v$ terms and dividing by $e^{-bt}$ (which we can do because it cannot equal zero), we are left with

$$v_t-kv_{xx}=0, \quad v(0,x)=u(0,x)=\psi(x) $$
Hence, the general solution of $v$ is

$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy, $$
and the general solution of $u$ is therefore

$$u(x,t)= \frac{e^{-bt}}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy. $$

## Question 6

Suppose that there is a maximum principle for the wave equation and that $u(x,t)$ is a solution to the wave equation in the closed rectangle $\Omega_T = \{ 0 \leq x \leq \pi, \; 0 \leq t \leq \pi/c \}$. Then the maximum of $u$ is assumed either initially (at $t=0$) or on the lines $x=0$ or $x=\pi$.

Consider the particular solution $u(x, t) = \sin x \sin ct$. This clearly satisfies the wave equation:
$$u_x=\cos x \sin ct $$
$$u_{xx}=-\sin x \sin ct $$
$$u_t=\sin x (c \cos ct) $$
$$u_{tt}=-\sin x (c^2 \sin ct)= c^2 u_{xx}$$

According to our presumed maximum principle, the maximum of $u(x,t)$ would have to be one of the following:
$$u(x,0)= \sin x \sin 0=0 $$
$$u(0, t)=\sin 0 \sin ct = 0 $$
$$u(\pi, t)=\sin \pi \sin ct = 0 $$

But $u(x,t)$ has a higher value when $x=\pi/2$ and $t=\pi/2c$:  
$$u\left(\frac{\pi}{2}, \frac{\pi}{2c}\right)=\sin \left(\frac{\pi}{2}\right) \sin \left(\frac{\pi}{2}\right)=1$$

This contradicts the definition of maximum, so we have found a counterexample and therefore there is no maximum principle for the wave equation.