### 2024-02-28 bio_dynamics/bio274_hw5.Rmd
started bio hw 5

---
title: "BIOL 274 Homework 5"
author: "Ruby Krasnow"
date: "2024-02-27"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Define each of the following scalars, vectors, and matrices:
$$s=3, r=-2, q= \frac{1}{4} $$

\begin{equation*}
    v = \begin{bmatrix} -8 \\ 4 \\ -16 \end{bmatrix},\quad
    w = \begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix},\quad
    A = \begin{bmatrix} 
    -1 & 3 & 2 \\ 
    2 & 0 & -1 \\ 
    3 & 5 & -4 \end{bmatrix},
\end{equation*}

\begin{equation*}
    B = \begin{bmatrix} 
    0 & -2 & 7 & 4 \\ 
    6 & 0 & -5 & 1 \\
    0 & 1 & 4 & -2 \end{bmatrix},\quad \text{and} \quad
    C = \begin{bmatrix} 
    11 & -3 & 1 & 0 \\
    -2 & 5 & 0 & -1 \\
    1 & 3 & 0 & 4 \end{bmatrix}
\end{equation*}

We will also define them in R, for use later in checking our manual computations
```{r}
#scalars
s=3
r=-2
q=(1/4)

#vectors
v=c(-8, 4, -16)
w=c(2,-1,3,4)

#matrices
A = matrix(c(-1, 2, 3, 3, 0, 5, 2, -1, -4), nrow=3, ncol=3)
B = matrix(c(0,6,0,-2,0,1,7,-5,4,4,1,-3), nrow=3)
C = matrix(c(11,-2,1,-3,5,3,1,0,0,0,-1,-4), nrow=3)
```

\newcommand{\negA}{
    \begin{bmatrix} 
    1 & -3 & -2 \\ 
    -2 & 0 & 1 \\ 
    -3 & -5 & 4 \end{bmatrix}}

\newcommand{\B}{
    \begin{bmatrix} 
    0 & -2 & 7 & 4 \\ 
    6 & 0 & -5 & 1 \\ 
    0 & 1 & 4 & -3 \end{bmatrix}}

\newcommand{\C}{
    \begin{bmatrix} 
    11 & -3 & 1 & 0 \\ 
    -2 & 5 & 0 & -1 \\ 
    1 & 3 & 0 & -4 \end{bmatrix}}

Use these to determine each of the following, if possible. If not possible, state why.

## 1. $sw$

\begin{equation*}
    sw = 3\begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix} = 
    \begin{bmatrix} 3(2) \\ 3(-1) \\ 3(3) \\ 3(4) \end{bmatrix} =
    \begin{bmatrix} 6 \\ -3 \\ 9 \\ 12 \end{bmatrix}
\end{equation*}

Check our manual computation:
```{r}
as.vector(s*w)
```

## 2. $qv$

\begin{equation*}
    qv = \frac{1}{4} \begin{bmatrix} -8 \\ 4 \\ -16 \end{bmatrix} =
    \begin{bmatrix} -8/4 \\ 4/4 \\ -16/4 \end{bmatrix} =
     \begin{bmatrix} -2 \\ 1 \\ -4 \end{bmatrix}
\end{equation*}

Check our manual computation:
```{r}
as.vector(q*v)
```

##3. $v+w$

This is not possible, because $v \in \mathbb{R}^3$ while $w \in \mathbb{R}^4$, and in order to add two vectors, they must be in the same n-dimensional space $\mathbb{R}^n$, where $n$ is the same for both vectors. In other words, $v$ has three elements while $w$ has four. To add vectors, they must be the same size.

When we try to do this in R, we should get an error:
```{r, error=TRUE}
# the "error=TRUE" argument allows it to show errors without halting R
as.vector(v+w)
```

##4. $(s+r)w$

$$(s+r)w=(3-2)w=w=\begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix}$$

Check our manual computation:
```{r}
as.vector((s+r) %*% w)
```

##5. $Aw$

This is not possible, because $A$ is a 3 by 3 matrix while $w$ is a vector in $\mathbb{R}^4$. In order for this to be possible, $w$ would need to have three elements, not four.

When we try to do this in R, we should get an error:
```{r, error=TRUE}
A %*% w
```

##6. $(-A)v$

$$ -A = \negA $$
$$ -Av = \negA \begin{bmatrix} -8 \\ 4 \\ -16 \end{bmatrix} =
\begin{bmatrix}
(1)(-8)+(-3)(4)+(-2)(-16) \\
(-2)(-8)+(0)(4)+(1)(-16) \\
(-3)(-8)+(-5)(4)+(4)(-16)
\end{bmatrix} = \begin{bmatrix} 12 \\ 0 \\ -60 \end{bmatrix} $$

Now we will check our manual computation:
```{r}
negA = -1*A #Define -A

negA %*% v #multiply A by v
```

##7. $Bw$

$$Bw=\B \begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix} =
\begin{bmatrix}
(0)(2)+(-2)(-1)+(7)(3)+(4)(4) \\
(6)(2)+(0)(-1)+(-5)(3)+(1)(4) \\
(0)(2)+(1)(-1)+(4)(3)+(-3)(4)
\end{bmatrix} = \begin{bmatrix} 39 \\ 1 \\ -1 \end{bmatrix}$$

Now we will check our manual computation:
```{r}
B %*% w #multiply B by w
```

##8. $B+C$

$$B+C=\B+\C=
\begin{bmatrix}
0+11 & -2-3 & 7+1 & 4+0 \\
6-2 & 0+5 & -5+0 & 1-1 \\
0+1 & 1+3 & 4+0 & -3-4 \end{bmatrix} =
\begin{bmatrix}
11 & -5 & 8 & 4 \\
4 & 5 & -5 & 0 \\
1 & 4 & 4 & -7 \end{bmatrix}$$

Now we will check our manual computation:
```{r}
B + C #add B and C
```

##9. $C-B$

$$C-B=\C-\B=
\begin{bmatrix}
11-0 & -3-2 & 1-7 & 0-4 \\
-2-6 & 5-0 & 0-(-5) & -1-1 \\
1-0 & 3-1 & 0-4 & -4-(-3) \end{bmatrix} =
\begin{bmatrix}
11 & -5 & -6 & -4 \\
-8 & 5 & 5 & -2 \\
1 & 2 & -4 & -1 \end{bmatrix}$$

Now we will check our manual computation:
```{r}
C-B #add B and C
```

##10. $Cv$

This is not possible, because $C$ is a 3 by 4 matrix while $v$ is a vector in $\mathbb{R}^3$. In order for this to be possible, $v$ would need to have four elements, not three.

When we try to do this in R, we should get an error:
```{r, error=TRUE}
C %*% v
```

##11. $(s-r)B$