### 2024-03-15 bio_dynamics/bio274_hw6.Rmd
bio hw 6

---
title: "BIOL 274 Homework 6"
author: "Ruby Krasnow"
date: "2024-03-15"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1  
Write each of the following systems in matrix/vector notation:

(a)
\begin{align*}
      x' &= 4x-6y \\
      y' &= 0.7x + y
\end{align*}

\begin{equation*}
\begin{bmatrix} x' \\ y' \end{bmatrix} = \begin{bmatrix} 
    4 & -6 \\
    0.7 & 1 \end{bmatrix} \begin{bmatrix} x \\ y \end{bmatrix}, \quad \text{ or }
  \end{equation*}

\begin{equation*}
\dot{X} = \begin{bmatrix} 
    4 & -6 \\
    0.7 & 1 \end{bmatrix} X
  \end{equation*}

(b)
\begin{align*}
    \frac{dx_1}{dt}-16x_2 &=-2x_1 \\
    \frac{dx_2}{dt}+4x_1 & =0
\end{align*}

\begin{equation*}
\begin{bmatrix} x_1' \\ x_2' \end{bmatrix} = \begin{bmatrix} 
    -2 & 16 \\
    -4 & 0 \end{bmatrix} \begin{bmatrix} x_1 \\ x_2 \end{bmatrix}, \quad \text{ or }
  \end{equation*}

\begin{equation*}
\dot{X} = \begin{bmatrix} 
    -2 & 16 \\
    -4 & 0 \end{bmatrix} X
  \end{equation*}

(c)
\begin{align*}
    \dot{v}=v-u+\sin{t} \\
    \dot{u}=2v+u-\cos{t}
\end{align*}

\begin{equation*}
\begin{bmatrix} \dot{v} \\ \dot{u} \end{bmatrix} = \begin{bmatrix} 
    1 & -1 \\
    2 & 1 \end{bmatrix} \begin{bmatrix}v \\ u \end{bmatrix} + \begin{bmatrix} \sin{t} \\ -\cos{t} \end{bmatrix}, \quad \text{ or }
  \end{equation*}

\begin{equation*}
\dot{V} = \begin{bmatrix} 
    1 & -1 \\
    2 & 1 \end{bmatrix} V + \begin{bmatrix} \sin{t} \\ -\cos{t} \end{bmatrix}
  \end{equation*}

## Question 2

Find the general solution for each of the following systems:  
(a)
\begin{align*}
      \dot{x} &= 9x-8y \\
      \dot{y} &= 24x-19y
\end{align*}

\begin{equation*}
A = \begin{bmatrix} 
    9 & -8 \\
    24 & -19 \end{bmatrix}
  \end{equation*}

\begin{equation*}
\det{\begin{bmatrix} 
    9-\lambda & -8 \\
    24 & -19\lambda \end{bmatrix}}=0 \\
\end{equation*}

$$ \left(9-\lambda\right) \left(-19-\lambda\right)+192=0 $$
\begin{equation*}
-171-9\lambda+19\lambda+ \lambda^2+192 = \lambda^2+10\lambda+21 =(\lambda+7)(\lambda+3)
\end{equation*}

So we have found two distinct real eigenvalues, $\lambda_1=-7, \; \lambda_2=-3$. Now we need to solve $AX=\lambda X$, or $(A-\lambda I)X=0$ with these two values of $\lambda$. Let's start with the first eigenvalue, -7.

\begin{equation*}
A+7I=\begin{bmatrix} 
    9+7 & -8 \\
    24 & -19+7 \end{bmatrix}= \begin{bmatrix} 
    16 & -8 \\
    24 & -12 \end{bmatrix} \end{equation*}

Now we use Gauss-Jordan elimination:
 \begin{equation*}   
   \begin{bmatrix} 
  16 & -8 \\  
  24 & -12   \end{bmatrix} \longrightarrow{}
 \begin{bmatrix} 
  -2 & 1 \\  
  -2 & 1
 \end{bmatrix}
\longrightarrow{}
 \begin{bmatrix} 
  -2 & 1 \\  
  0 & 0 \end{bmatrix} \end{equation*}

 \begin{equation*} 
  -2x_1+x_2=0 \quad \rightarrow{} \quad x_2=2x_1 \quad \rightarrow{} \quad 
  v_1=
   \begin{bmatrix} 1 \\ 2  \end{bmatrix}
\end{equation*}

And repeat for the second eigenvalue, -3:
\begin{equation*} A+3I=\begin{bmatrix} 
    9+3 & -8 \\
    24 & -19+3 \end{bmatrix}= \begin{bmatrix} 
    12 & -8 \\
    24 & -16 \end{bmatrix}
\end{equation*}

 \begin{equation*}   
   \begin{bmatrix} 
  12 & -8 \\  
  24 & -16   \end{bmatrix} \longrightarrow{}
 \begin{bmatrix} 
  -3 & 2 \\  
  0 & 0
 \end{bmatrix} \end{equation*}

 \begin{equation*} 
  -3x_1+2x_2=0 \quad \rightarrow{} \quad 2x_2=3x_1 \quad \rightarrow{} \quad v_2=
   \begin{bmatrix} 2 \\ 3  \end{bmatrix}
\end{equation*}

So we have found $\lambda_1=-7$ with eigenvector $\begin{bmatrix} 1 \\ 2  \end{bmatrix}$ and $\lambda_2=-3$ with eigenvector $\begin{bmatrix} 2 \\ 3  \end{bmatrix}$. We can plug these values into our formula for the general solution of a system of linear first-order ODEs with constant coefficients:

$$y=C_1e^{\lambda_1 t}v_1 +C_2e^{\lambda_2 t}v_2$$
\begin{tcolorbox}[colback=white, title=Solution]
$$y=C_1e^{-7 t}\begin{bmatrix} 1 \\ 2  \end{bmatrix} +C_2e^{-3 t}\begin{bmatrix} 2 \\ 3 \end{bmatrix}$$
\end{tcolorbox}

\newpage
Let's also do a quick check using R:
```{r}
A <- matrix(c(9, -8, 24, -19), 2, 2, byrow=TRUE)

lambda1 <- eigen(A)$values[1] # eigenvalue 1
v1<- eigen(A)$vectors[,1] / eigen(A)$vectors[1,1] #eigenvector 1
# divided by the first element to give integer values

lambda2 <- eigen(A)$values[2] # eigenvalue 2
v2 <- eigen(A)$vectors[,2] / eigen(A)$vectors[1,2]*2 #eigenvector 2
# divided by the first element and multiplied by 2 to give integer values
```

```{r}
paste("lambda 1 = ", lambda1, " with eigenvector [", v1[1], ",", v1[2], "]", sep="")
paste("lambda 2 = ", lambda2, " with eigenvector [", v2[1], ",", v2[2], "]", sep="")
```

(b)

$$y=C_1e^{-t} \begin{bmatrix} \sin{2t} \\ \cos{2t} \end{bmatrix}+
C_2e^{-t} \begin{bmatrix} -\cos{2t} \\ \sin{2t} \end{bmatrix} $$