### 2024-03-16 bio_dynamics/bio274_filing_cabinet.Rmd
finished bio hw6

```{r}
x=c(100:5000)

fish1 <- tibble(x=x,
                orig=0.3*(1-(x/2500)),
                constant=0.3*(1-(x/2500))-100,
                third=0.3*(1-(x/2500))-x/3,
                prop=0.3*(1-(x/2500))-0.5*sqrt(x))
fish1 <- fish1 %>% pivot_longer(cols=2:5, names_to = "eqn", values_to = "dP_dt")

ggplot()+
 # geom_line(data=fish1, aes(x=x, y=dP_dt, color=eqn))+
  geom_line(data=fish1 %>% filter(eqn=="orig"), aes(x=x, y=dP_dt, color=eqn))+
  theme_bw()+
  labs(x="P", y="dP/dt", color="type")
```

\newpage
Let's first do a quick check of our eigenvalues and eigenvectors using R:
```{r}
A <- matrix(c(9, -8, 24, -19), 2, 2, byrow=TRUE)

lambda1 <- eigen(A)$values[1] # eigenvalue 1
v1<- eigen(A)$vectors[,1] / eigen(A)$vectors[1,1] #eigenvector 1
# divided by the first element to give integer values

lambda2 <- eigen(A)$values[2] # eigenvalue 2
v2 <- eigen(A)$vectors[,2] / eigen(A)$vectors[1,2]*2 #eigenvector 2
# divided by the first element and multiplied by 2 to give integer values
```

```{r}
paste("lambda 1 = ", lambda1, " with eigenvector [", v1[1], ",", v1[2], "]", sep="")
paste("lambda 2 = ", lambda2, " with eigenvector [", v2[1], ",", v2[2], "]", sep="")
```