### 2024-03-18 PDEs/math245_hw4.Rmd
started math hw4

---
title: "MATH 245 Homework 4"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-03-18"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1: Find eigenvalues and eigenfunctions

**(a)**

$- \frac{d^2}{dx^2}X(x)=\lambda X(x)$ in $0 < x < l$ with boundary conditions $X'(0) = 0 = X(l)$.

**(b)**

$x^2 X''(x) + xX'(x)+\lambda X(x) = 0$ in $1 < x < e$ with boundary conditions $X(1) = 0 = X(e)$.

**(c)**

On the interval $0 \leq x \leq 1$ of length one, consider the eigenvalue problem
$$-X'' = \lambda X,\quad  X'(0) + X(0) = 0,\quad X(1) = 0$$

(i) Find an eigenfunction with eigenvalue zero. Call it $X_0(x)$.

(ii) Find an equation for the positive eigenvalues $\lambda = \beta^2$.

(iii) Show graphically from part (b) that there are an infinite number of positive eigenvalues.

(iv) Is there a negative eigenvalue?

## Question 2  
Find the Fourier-series of $f(x)$. Does the Fourier-series converge (i) pointwise, or (ii)
uniformly?

**(a)**
\begin{equation*} f(x)=
    \begin{cases}
      1-\lvert x \rvert & \lvert x \rvert \leq 1\\
      1 & 1 < \lvert x \rvert \leq \pi\\
    \end{cases}       
\end{equation*}

**(b)**
\begin{equation*} f(x)=\lvert x \rvert=
    \begin{cases}
      -x & -\pi \leq x \leq 0 \\
      x &  0<x\leq \pi \\
    \end{cases}       
\end{equation*}

**(c)**
$$f(x)=x+x^2, \quad -\pi \leq x \leq \pi $$

## Question 3

**(a)** Find the Fourier-sine-series of
\begin{equation*} f(x)=
    \begin{cases}
      1 & 0 < x < \pi /2 \\
      2 &   \pi /2 < x < \pi \\
    \end{cases}       
\end{equation*}

**(b)** Find the Fourier-cosine-series of $f(x)= \lvert \sin{x} \rvert$. Then find
$$\sum^\infty_{k=1} \frac{1}{4k^2-1} $$

**(c)** The Riemann Zeta function is defined for $s > 1$ by
$$ \zeta(s)= \sum^\infty_{n=1} \frac{1}{n^s}$$

By computing the Fourier series of $x^2$ over $-\pi<x<\pi$ and using Parseval’s identity, compute 
$\zeta(4)$.

**(d)** Use the Fourier series in 2c and the pointwise convergence theorem to find $\zeta(2)$. Then find
$$\sum^\infty_{n=1} \frac{1}{(2n-1)^2} $$

## Question 4
Compute the complex Fourier series of the following functions:

**(a)** Compute the complex Fourier series of $f(x) = e^x$ and show that

$$\coth{\pi}= \frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} + \;...\right) $$

**(b)** Find the complex Fourier series of $xe^{ix}$. Then use your result to find the Fourier series of $x \cos{x}$ and $x \sin{x}$.

## Question 5.

Find the function represented by the new series which is obtained by termwise integration of the following series from $0$ to $x$.