### 2024-03-22 bio_dynamics/bio274_hw7.Rmd

---
title: "BIOL 274 Homework 7"
author: "Ruby Krasnow"
date: "2024-03-22"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---
# Visualizing Solutions of Linear Systems with Constant Coefficients

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

For each of the following systems, state the general solution and sketch the phase portrait (with at least 4 different solutions / ICs shown):

## Question 1

## Question 2

## Question 3

## Question 4

## Question 5

## Question 6

## Question 7

## Question 8