### 2024-03-25 PDEs/math245_hw4.Rmd
worked on pdes

---
title: "MATH 245 Homework 4"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-03-24"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1: Find eigenvalues and eigenfunctions

### (a)

$- \frac{d^2}{dx^2}X(x)=\lambda X(x)$ in $0 < x < l$ with boundary conditions $X'(0) = 0 = X(l)$

\underline{Case 1:} Positive eigenvalues, $\lambda = \beta^2 >0$

Re-writing as $X''+\lambda X=0$, we will get the characteristic equation: $r^2+\beta^2=0$. Since our characteristic equation has complex roots $r= \pm i\beta$, our solutions take the form 
$$X(x)=A\sin{(\beta x)}+B\cos{(\beta x)}$$
Differentiating, we find that 
$$X'(x)=\beta A\cos{(\beta x)}-\beta B\sin{(\beta x)}$$
Now plugging in our initial condition $X'(0) = 0$, we get
$X'(0)=\beta A=0$. And since we are in a case where $\beta \neq 0$, this means $B=0$ so $X(x)=A\sin{(\beta x)}$. Now we use our boundary condition $X(l)=0$ to get $X(l)=A\sin{(\beta l)}=0$. If $A=0$, then $X(0)=0$ and this contradicts the definition of an eigenfunction. Therefore, we must have $\sin{(\beta l)}=0$, which can only occur if $\beta = \frac{(2n+1)\pi}{l}$. Therefore, this case gives us eigenvalues

$$\lambda_n= \left(\frac{(2n+1)\pi}{l}\right)^2 \quad n=1,2,3... $$
with eigenfunctions
$$X_n(x)=\sin{\left(\frac{(2n+1)\pi x}{l}\right)}$$

\underline{Case 2:} Zero eigenvalues, $\lambda=0$
$X''=0$ implies that $X(x)$ is of the form $Ax+B$, with derivative $X'(x)=A$. Now plugging in our initial condition $X'(0) = 0$, we get $A=0$, which means $X(x)=B$. But from the boundary condition $X(l)=0$, we get $B=0$ and so $X(0)=0$. Therefore, there are no eigenfunctions $X(x)$ that satisfy $- \frac{d^2}{dx^2}X(x)=\lambda X(x)$ when $\lambda=0$ and hence no zero eigenvalues.

\underline{Case 3:} Negative eigenvalues, $\lambda = -\beta^2 <0$
This case gives us the characteristic equation: $r^2-\beta^2=0$. Since our characteristic equation has distinct real roots $r= \pm \beta$, our solutions take the form 
$$X(x)=Ae^{\beta x}+Be^{-\beta x}$$
Differentiating, we find that
$$X'(x)=\beta Ae^{\beta x}-\beta Be^{-\beta x}$$

$$X'(0)=\beta A-\beta B=0$$
Since we are in a case where $\beta \neq 0$, this means $A-B=0$, or $A=B$. Then the boundary condition gives
$$X(l)=Be^{\beta l}+Be^{-\beta l}=0$$

Since $e^{\beta l}$ and $e^{-\beta l}$ are nonzero for all values of $l$, we must have $B=0$, implying that again $X(x)=0$.

Thus, this problem has only positive eigenvalues and their associated eigenfunctions as found in Case 1.

### (b)

$x^2 X''(x) + xX'(x)+\lambda X(x) = 0$ in $1 < x < e$ with boundary conditions $X(1) = 0 = X(e)$.

We recognize this equation as having the same form as a second-order Cauchy-Euler equation, a linear homogeneous ODE of the form $ax^2y+bxy'+cy=0$ with the auxiliary equation $ar(r-1)+br+c=0$. Here, $a=b=1$ and $c=\lambda$, so we have

$$r(r-1)+r+\lambda=0$$
$$r^2-r+r+\lambda=0$$
$$r^2+\lambda=0$$
\underline{Case 1:} Positive eigenvalues, $\lambda = \beta^2 >0$

$$X(x)=A\sin{(\beta \ln{x})}+B\cos{(\beta \ln{x})}$$
$$X(1)=A\sin{(0)}+B\cos{(0)}=0 \quad \longrightarrow \quad B=0$$
$$X(e)=A\sin{(\beta \ln{e})}=0$$
$$A\sin{(\beta)}=0$$
If $A=0$, then $X(0)=0$ and this contradicts the definition of an eigenfunction. Therefore, we must have $\sin{(\beta)}=0$, which can only occur if $\beta = (2n+1)\pi$. Therefore, this case gives us eigenvalues 
$$\lambda_n= (2n+1)^2 \pi^2 \quad n=1,2,3... $$
with eigenfunctions
$$X_n(x)=\sin{\left((2n+1)\pi \ln{x}\right)}$$

\underline{Case 2:} Zero eigenvalues, $\lambda=0$
$X''=0$ implies that $X(x)$ is of the form $Ax+B$. Now plugging in our initial condition $X(1) = 0$, we get $A+B=0$, which means $X(x)=Ax-A$. But from the boundary condition $X(e)=0$, we get $A(e-1)=0$ which is only possible if $A=0$ and accordingly $X(0)=0$. Therefore, there are no eigenfunctions $X(x)$ that satisfy the boundary conditions when $X''=0$ and hence no zero eigenvalues.

\underline{Case 3:} Negative eigenvalues, $\lambda = -\beta^2 <0$
This case gives us the characteristic equation: $r^2-\beta^2=0$. Since our characteristic equation has distinct real roots $r= \pm \beta$, our solutions take the form 
$$X(x)=Ax^{\beta }+Bx^{-\beta}$$

$$X(1)=A+B=0 \quad \longrightarrow\quad A=-B$$
$$X(x)=Ax^{\beta }-Ax^{-\beta}$$
Then the other boundary condition gives $X(e)=Ae^{\beta}-Ae^{-\beta }=0$. Since $e^{\beta}$ and $e^{-\beta}$ are always nonzero, we must have $A=0$, implying that again $X(x)=0$.Thus, this problem has only positive eigenvalues and their associated eigenfunctions as found in Case 1.

### (c)

On the interval $0 \leq x \leq 1$ of length one, consider the eigenvalue problem
$$-X'' = \lambda X,\quad  X'(0) + X(0) = 0,\quad X(1) = 0$$

(i) Find an eigenfunction with eigenvalue zero. Call it $X_0(x)$.

$X''=0$ implies that $X(x)$ is of the form $Ax+B$ and $X'(x)=A$. Then $X(1)=0$ becomes $A+B=0$, or $A=-B$. Now we look at our second condition, 
$$ X'(0) + X(0) = 0$$
$X'(0)$ is just $A$ and $X(0)$ is just $B$, so again this gives us $A=-B$. Thus, we have $X(x)=Ax-A=A(x-1)$, so we have found
$$X_0(x)=x-1 $$

(ii) Find an equation for the positive eigenvalues $\lambda = \beta^2$.

$$X(x)=A\sin{(\beta x)}+B\cos{(\beta x)}$$
$$X'(x)=\beta A\cos{(\beta x)}-\beta B\sin{(\beta x)}$$

$$X(1)=A\sin{(\beta)}+B\cos{(\beta)}=0$$
$$A\sin{(\beta)}=-B\cos{(\beta)}$$

\begin{equation}\label{aciii}
-\frac{B}{A}=\frac{\sin{(\beta)}}{\cos{(\beta)}}=\tan{\beta}
\end{equation}

$$ X'(0) =\beta A, \quad X(0)=B$$
$$ X'(0) + X(0) = \beta A+B=0 $$
\begin{equation}\label{aciv}\beta=-\frac{B}{A}\end{equation}

Combining (\ref{aciii}) and (\ref{aciv}), we get the equation $\beta=\tan{(\beta)}$ for the positive eigenvalues.

(iii) Show graphically from part (b) that there are an infinite number of positive eigenvalues.

In part (b), we showed that $\beta$ is a positive eigenvalue if $\beta=\tan{(\beta)}$. However, plotting the equations $y=\beta$ and $y=\tan{\beta}$ reveals that these two functions intersect an infinite number of times (Fig. 1), meaning there are an infinite number of positive eigenvalues.

```{r, echo=FALSE,   fig.height=3, fig.width=5}
df_1c <- tibble(x = seq(-1*pi, 10*pi, length.out = 1000),
                y=tan(x)) %>% 
  mutate(y = ifelse(abs(y) < 30, y, NA))

ggplot(df_1c) + 
  geom_line(aes(x=x, y=y, color="A"))+
  geom_line(aes(x=x, y=x, color="B"))+
  geom_hline(yintercept=0, color="black")+
 coord_cartesian(ylim=c(-10,10), xlim=c(-1,11))+
  theme_bw()+
  labs(y="y", x=expression(beta), color="")+ scale_colour_discrete(labels=c(expression(paste("y=tan(", beta, ")")), expression(paste("y=", beta))))+ggtitle("Figure 1")

```

(iv) Is there a negative eigenvalue?

This case gives us the characteristic equation: $r^2-\beta^2=0$. Since our characteristic equation has distinct real roots $r= \pm \beta$, our solutions take the form 
$$X(x)=Ae^{\beta x}+Be^{-\beta x}$$
Differentiating, we find that
$$X'(x)=\beta Ae^{\beta x}-\beta Be^{-\beta x}$$

$$X'(0)=\beta A-\beta B$$
$$X(0)=A+B$$

$$ X'(0) + X(0) = \beta A-\beta B+A+B$$
$$0=A(\beta+1)+B(1-\beta)$$
$$-A(\beta+1)=B(1-\beta)$$
$$\frac{\beta+1}{1-\beta}=-\frac{B}{A}$$
Now turning to our other boundary condition,
$$X'(1)=\beta Ae^{\beta}-\beta Be^{-\beta}=0$$
$$\beta Ae^{\beta}=\beta Be^{-\beta}$$
$$-\frac{B}{A} = -e^{2\beta}$$
This would give us the following equation:
$$\frac{\beta+1}{1-\beta}=-e^{2\beta}$$
But the only time that this equality holds true is when $\beta=0$ (Fig. 2), but we are in the case when we defined $\beta$ to be nonzero, meaning that we have reached a contradiction and we cannot have any negative eigenvalues.

```{r, echo=FALSE,   fig.height=3, fig.width=5}
df_1d <- tibble(x = seq(-10, 10, length.out = 1000),
                y1=exp(2*x),
                y2=(x+1)/(1-x)) %>% 
  mutate(y2 = ifelse(abs(y2) < 30, y2, NA))

ggplot(df_1d) + 
  geom_hline(yintercept=0, color="black")+
  geom_vline(xintercept=0, color="black")+
  geom_line(aes(x=x, y=y1, color="A"))+
  geom_line(aes(x=x, y=y2, color="B"))+
  annotate(geom = "point",x=0,y=1)+
  coord_cartesian(ylim=c(-6,8), xlim=c(-4,4))+
  theme_bw()+
  labs(y="y", x=expression(beta), color="")+ scale_colour_discrete(labels=c(expression(paste("y=exp(2", beta, ")")), expression(paste("y=(", beta, "+1)/(1-", beta, ")"))))+ggtitle("Figure 2")

```

## Question 2  
Find the Fourier series of $f(x)$. Does the Fourier-series converge (i) pointwise, or (ii)
uniformly?

### (a)
\begin{equation*} f(x)=
    \begin{cases}
      1-\lvert x \rvert & \lvert x \rvert \leq 1\\
      1 & 1 < \lvert x \rvert \leq \pi\\
    \end{cases}       
\end{equation*}

The Fourier series for a function $f(x)$ on an interval $-l<x<l$ is
$$f(x)=a_0+ \sum_{n=1}^{\infty} \left( a_n \cos{  \left(\frac{n\pi x}{l}\right) }+b_n \sin{ \left(\frac{n\pi x}{l}\right)}\right)$$
where
$$a_0=\frac{1}{2l} \int_{-l}^{l}f(x)dx$$ 
$$ a_n=\frac{1}{l} \int_{-l}^{l}f(x)\cos{  \left(\frac{n\pi x}{l}\right) }dx$$
$$ b_n=\frac{1}{l} \int_{-l}^{l}f(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx$$

Let's start with $a_0$:
\begin{align*}
a_0 &=\frac{1}{2\pi} \int_{-\pi}^{\pi}f(x)dx \\
    &=\frac{1}{2\pi}\left[ \int_{-\pi}^{-1} 1\;dx + \int_{-1}^{1} (1-\lvert x \rvert) dx+ \int_{1}^{\pi} 1\;dx  \right] \\
    &= \frac{1}{2\pi}\int_{-\pi}^{\pi}1\;dx - \int_{-1}^{1} \lvert x \rvert\, dx \\
    &= 1- 2\int_{0}^{1} x \; dx \\
    &= 1- 1=0
\end{align*}

where we used the evenness of $\lvert x \rvert$ to rewrite $\int_{-1}^{1} \lvert x \rvert \,dx$ as $2\int_{0}^{1} x \; dx$. Thus, $a_0=0$. Now we perform a similar procedure to find $a_n$:

\begin{align*}
a_n &= \frac{1}{\pi} \int_{-\pi}^{\pi} f(x) \cos{(nx)}dx \\
    &= \frac{1}{\pi}\left[ \int_{-\pi}^{-1} \cos{(nx)}\;dx + \int_{-1}^{1} (1-\lvert x \rvert)\cos{(nx)}\; dx+ \int_{1}^{\pi} \cos{(nx)}\;dx  \right] \\
    &= \frac{1}{\pi} \int_{-\pi}^{\pi} \cos{(nx)}\;dx - \frac{1}{\pi} \int_{-1}^{1} \lvert x \rvert\cos{(nx)}\; dx
\end{align*}

Since $\sin{(\pi n)}$ and $\sin{(-\pi n)}$ are zero for any integer $n$, the first integral vanishes. We use the symmetry of $\lvert x \rvert \cos{(nx)}$ (both functions are even, and the product of two even functions is even) and rewrite the second integral as $$- \frac{2}{\pi} \int_{0}^{1}  x \cos{(nx)} \; dx $$

We then use integration by parts:
\begin{align*}
&- \frac{2}{\pi} \int_{0}^{1}  x \cos{(nx)}\; dx \\
&= - \frac{2}{n\pi} \left[x \sin{(nx)} \Bigr\rvert_0^1 - \int_{0}^{1}\sin{(nx)}\; dx \right] \\
&= - \frac{2}{n\pi} \left[\sin{(n)}+ \frac{1}{n} \cos{(nx)}\Bigr\rvert_0^1  \right] \\
&= - \frac{2}{n\pi} \left[\sin{(n)}+ \frac{1}{n} \cos{(n)} - \frac{1}{n} \right]
\end{align*}

$$ a_n= - \frac{2 \sin{n}}{n\pi} + \frac{2(\cos{n}-1)}{n^2 \pi} $$
\begin{align*}
b_n &=
\frac{1}{\pi}\left[ \int_{-\pi}^{-1} \sin{(nx)}\;dx + \int_{-1}^{1} (1-\lvert x \rvert)\sin{(nx)}\; dx+ \int_{1}^{\pi} \sin{(nx)}\;dx  \right] \\
    &=  \frac{1}{\pi} \int_{-\pi}^{\pi} \sin{(nx)}\;dx - \frac{1}{\pi} \int_{-1}^{1} \lvert x \rvert\sin{(nx)}\; dx
\end{align*}

But $\sin{x}$ is an odd function and so is $\lvert x \rvert\sin{(nx)}$, so both of these integrals vanish and we find $b_n=0$.

Now inserting the coefficients $a_n$ we found above, we get the Fourier series:
$$f(x)=\frac{2}{\pi} \sum_{n=1}^{\infty}  \left(\frac{\sin{n}}{n}+ \frac{\cos{n}+1}{n^2} \right) \cos{ \left( nx \right)}$$

### (b)
\begin{equation*} f(x)=\lvert x \rvert=
    \begin{cases}
      -x & -\pi \leq x \leq 0 \\
      x &  0<x\leq \pi
    \end{cases}
\end{equation*}

Let's start with $a_0$:
\begin{align*}
a_0 &=\frac{1}{2\pi} \int_{-\pi}^{\pi}\lvert x \rvert \, dx \\
    &=\frac{1}{\pi}\int_{0}^{\pi} x \;dx \\
    &= \frac{1}{2\pi} x^2 \Bigr\rvert_0^{\pi} = \frac{\pi}{2}
\end{align*}

\begin{align*}
a_n &= \frac{1}{\pi} \int_{-\pi}^{\pi}f(x)\cos{\left(nx\right) }dx \\
    &= \frac{2}{\pi} \int_{0}^{\pi}x\cos{\left(nx\right) }dx \\
    &= \frac{2}{n \pi} x \sin{(nx)} \Bigr\rvert_0^{\pi} - \frac{2}{n \pi}\int_{0}^{\pi}\sin{\left( nx \right) } dx \\
    &= 0 + \frac{2}{n^2 \pi} \cos{(nx)}\Bigr\rvert_0^{\pi} \\
    &= \frac{2}{n^2 \pi} (-1)^{n+1}
\end{align*}

$$b_n = \frac{1}{\pi} \int_{-\pi}^{\pi} f(x) \sin{(nx)}dx $$
But since $\lvert x \rvert$ is even and $\sin{(nx)}$ is odd, the product of the two functions is odd and so the integral vanishes over the symmetric interval.

Then we plug in our formulas for $a_0$ and $a_n$:
$$ f(x) = \frac{\pi}{2} + \frac{2}{\pi} \sum_{n=1}^{\infty} \left[ \frac{1}{n^2} (-1)^{n+1} \cos{(nx)} \right]$$

### (c)

$$f(x)=x+x^2, \quad -\pi \leq x \leq \pi$$
\begin{align*}
a_0 &=\frac{1}{2\pi} \int_{-\pi}^{\pi}x + x^2 dx \\
    &=\frac{1}{2\pi} \left( \frac{x^2}{2}+ \frac{x^3}{3} \right) \Bigr\rvert_{-\pi}^{\pi} \\
    &= \frac{1}{2\pi} \left( \frac{\pi^2}{2} + \frac{\pi^3}{3} - \frac{\pi^2}{2} + \frac{\pi^3}{3}\right) \\
    &= \frac{1}{2\pi} \left( \frac{2\pi^3}{3} \right) = \frac{\pi^2}{3}
\end{align*}

\begin{align*}
a_n &= \frac{1}{\pi} \int_{-\pi}^{\pi}f(x)\cos{(nx)}dx \\
    &= \frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}+ \frac{1}{\pi} \int_{-\pi}^{\pi}x^2\cos{(nx)}dx 
\end{align*}

Applying integration by parts to the first integral,
$$\frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}=\frac{1}{n\pi}x\sin{(nx)} \Bigr\rvert_{-\pi}^{\pi}+\frac{1}{n\pi} \int_{-\pi}^{\pi}\sin{(nx)}$$
$$=\frac{1}{n\pi}x\sin{(nx)} \Bigr\rvert_{-\pi}^{\pi}+\frac{-1}{n^2\pi} \cos{(nx)}\Bigr\rvert_{-\pi}^{\pi}=0+0=0$$

We then use the integration by parts table method demonstrated in class to find that

$$\frac{1}{\pi} \int_{-\pi}^{\pi}x^2\cos{(nx)}= \frac{x^2}{n}\sin{(nx)}\Bigr\rvert_{-\pi}^{\pi} +  \frac{2x}{n^2}\cos{(nx)}\Bigr\rvert_{-\pi}^{\pi}+\frac{2}{n^3}\sin{(nx)}\Bigr\rvert_{-\pi}^{\pi}$$
The first and last terms (with $\sin{(nx)}$) vanish, while the middle term evaluates to $\frac{4 \cos{(n\pi)}}{n^2}$.

$$\frac{4 \cos{(n\pi)}}{n^2}=\frac{4}{n^2} (-1)^{n}$$
Thus, we add this to the result of first integral to find that
$$a_n= \frac{4}{n^2} (-1)^{n} $$
\begin{align*}
b_n &= \frac{1}{\pi} \int_{-\pi}^{\pi}f(x)\sin{(nx)}dx \\
    &= \frac{1}{\pi} \int_{-\pi}^{\pi}x\sin{(nx)}+ \frac{1}{\pi} \int_{-\pi}^{\pi}x^2\sin{(nx)}dx 
\end{align*}

But since $x^2$ is even and $\sin{}$ is odd, the product under the second integral is odd functions and thus evaluates to zero over the symmetric interval $-\pi<x<\pi$. Applying integration by parts to the first integral, 
$$b_n=\frac{-1}{n\pi}x\cos{(nx)} \Bigr\rvert_{-\pi}^{\pi}+\frac{1}{n\pi} \int_{-\pi}^{\pi}\cos{(nx)}$$
But since $\cos$ is even and $\cos{\pi}=\cos{-\pi}$, both terms are zero. Thus, $b_n=0$ and

$$f(x)=a_0 + \sum_{n=1}^{\infty}  a_n \cos{  \left(\frac{n\pi x}{l}\right) } $$
$$f(x)=\frac{\pi^2}{3} + 4 \sum_{n=1}^{\infty} \frac{(-1)^{n}\cos{ (nx) }}{n^2}  $$

## Question 3

### a)  
Find the Fourier sine series of
\begin{equation*} f(x)=
    \begin{cases}
      1 & 0 < x < \pi /2 \\
      2 &   \pi /2 < x < \pi \\
    \end{cases}
\end{equation*}

The Fourier sine series of a function $f(x)$ is $$f(x)= \sum_{n=1}^{\infty} A_n \sin{  \left(\frac{n\pi x}{l}\right) }$$
where 
$$A_n = \frac{2}{l} \int_{0}^{l}f(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx $$
$$A_n = \frac{2}{\pi} \int_{0}^{\pi}f(x)\sin{(nx)}dx $$
\begin{align*}
A_n &= \frac{2}{\pi}\left[ \int_{0}^{\pi/2}\sin{(nx)}dx + 2 \int_{\pi/2}^{\pi}\sin{(nx)}dx \right] \\
    &= \frac{2}{\pi}\left[\frac{-1}{n} \cos{(nx)}\Bigr\rvert_{0}^{\pi/2}- \frac{2}{n} \cos{(nx)}\Bigr\rvert_{\pi/2}^{\pi}\right] \\
    &= \frac{2}{n \pi}\left[\cos{\left(\frac{n\pi}{2}\right)}+1-2\cos{(n\pi)} \right]
\end{align*}

 $$f(x)=  \frac{2}{\pi} \sum_{n=1}^{\infty} \frac{1}{n}\left[1+\cos{\left(\frac{n\pi}{2}\right)}-2\cos{(n\pi)} \right] \sin{(nx) }$$

### b)

Find the Fourier-cosine-series of $f(x)= \lvert \sin{x} \rvert$. Then find
$$\sum^\infty_{k=1} \frac{1}{4k^2-1} $$

The Fourier cosine series of a function $f(x)$ is $$f(x)=\frac{1}{2}A_0+  \sum_{n=1}^{\infty} A_n \cos{  \left(\frac{n\pi x}{l}\right) }$$
where 
$$A_n = \frac{2}{l} \int_{0}^{l}f(x)\cos{  \left(\frac{n\pi x}{l}\right) }dx $$
Assuming $l=\pi$, we have
\begin{equation}\label{eqn:3b}
A_n = \frac{2}{\pi} \int_{0}^{\pi}\lvert \sin{x} \rvert\cos{(nx)}dx
\end{equation}

But since $\sin$ is non-negative on the interval $0 \leq x \leq \pi$, we can drop the absolute value bars and use the following trigonometric identity to rewrite the integral: $\sin{\alpha}\cos{\beta} = \frac{1}{2} [ \sin{(\alpha + \beta)} + \sin{(\alpha - \beta)}]$

\begin{align*}
A_n &= \frac{1}{\pi} \int_{0}^{\pi} \sin{((1+n)x)} +\sin{((1-n)x)}dx \\
    &= \frac{1}{\pi}\left[ \frac{-1}{1+n} \cos{((1+n)x)}\Bigr\rvert_{0}^{\pi} -\frac{1}{1-n}\cos{((1-n)x)}\Bigr\rvert_{0}^{\pi} \right] \\
    &= \frac{1}{\pi}\left[ \frac{-1}{1+n} [ (-1)^{1+n}-1] -\frac{1}{1-n} [ (-1)^{1-n}-1]\right] \\
    & = \frac{1}{\pi}  \left[ (-1)^{1+n}-1) \right] \left(\frac{-1}{1+n} -\frac{1}{1-n} \right)
\end{align*}

Thus, we have found a formula for $A_n$,

$$A_n = \frac{-2}{\pi}  \frac{(-1)^{n}+1}{n^2-1}  $$
This formula works for all $n \neq 1$
For $n=0$, 
$$A_0 = \frac{-2}{\pi}  \frac{(-1)^{0}+1}{0-1}=  \frac{4}{\pi}  $$

For $n=1$, we can go back to (\ref{eqn:3b}) and plug in the appropriate value of $n$:
$$A_1 = \frac{1}{\pi} \int_{0}^{\pi} \sin{(2x)}\;dx = \frac{1}{2\pi} \cos{(2x)}\Bigr\rvert_{0}^{\pi} \;dx=0 $$
The Fourier cosine series is therefore
$$f(x)=\frac{2}{\pi}-\frac{2}{\pi}  \sum_{n=2}^{\infty}   \frac{(-1)^{n}+1}{n^2-1}\cos{(nx) }$$

Now, $(-1)^{n}+1$ will be zero for all odd $n$ and two for all even $n$. So substituting $n=2k$, we gwt

$$\lvert \sin{x} \rvert=\frac{2}{\pi}-\frac{4}{\pi}  \sum_{2k=2}^{\infty}   \frac{1}{4k^2-1}\cos{(2kx) }$$

Starting the index at $2k=2$ is equivalent to starting it at $k=1$, so after making that change and evaluating both sides at $x=0$, we can solve for the sum given in the question:
$$0=\frac{2}{\pi}-\frac{4}{\pi}  \sum_{k=1}^{\infty}   \frac{1}{4k^2-1}$$
$$\sum_{k=1}^{\infty}   \frac{1}{4k^2-1}=\frac{1}{2}$$

### c)

The Riemann Zeta function is defined for $s > 1$ by
$$ \zeta(s)= \sum^\infty_{n=1} \frac{1}{n^s}$$
By computing the Fourier series of $x^2$ over $-\pi<x<\pi$ and using Parseval’s identity, compute
$\zeta(4)$.

\begin{align*}
a_0 &=\frac{1}{2\pi} \int_{-\pi}^{\pi}x^2 dx \\
    &=\frac{1}{2\pi} \left(\frac{x^3}{3}\right) \Bigr\rvert_{-\pi}^{\pi} \\
    &= \frac{1}{2\pi} \left( \frac{\pi^3}{3} - \frac{-\pi^3}{3}\right) \\
    &= \frac{1}{2\pi} \left( \frac{2\pi^3}{3} \right) = \frac{\pi^2}{3}
\end{align*}

$$a_n = \frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}dx$$

But in part (2b), we found that this evaluates to zero.

<!-- **(d)** Use the Fourier series in 2c and the pointwise convergence theorem to find $\zeta(2)$. Then find -->
<!-- $$\sum^\infty_{n=1} \frac{1}{(2n-1)^2} $$ -->

<!-- ## Question 4 -->
<!-- Compute the complex Fourier series of the following functions: -->

<!-- **(a)** Compute the complex Fourier series of $f(x) = e^x$ and show that -->

<!-- $$\coth{\pi}= \frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} + \;...\right) $$ -->

<!-- **(b)** Find the complex Fourier series of $xe^{ix}$. Then use your result to find the Fourier series of $x \cos{x}$ and $x \sin{x}$. -->

<!-- ## Question 5 -->

<!-- Find the function represented by the new series which is obtained by termwise integration of the following series from $0$ to $x$. -->

<!-- $$ \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}=\log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }, \quad -\pi<x<\pi$$ -->