### 2024-04-03 bio_dynamics/bio274_hw8.Rmd
finished pde 4

---
title: "BIOL 274 Homework 8"
author: "Ruby Krasnow"
date: "2024-04-02"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---
# Visualizing Solutions of Linear Systems and Runga Kutta