### 2024-04-05 PDEs/math245_hw5.Rmd
started 274 hw8

---
title: "MATH 245 Homework 5"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-04-05"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1: