### 2024-04-09 PDEs/math245_old.Rmd

---
title: "MATH 245 old and unused content"
author: "Ruby Krasnow"
date: "2024-01-31"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Homework 2 Question 4 - old version
For each of the following IVPs, (i) find and plot the characteristic lines (curves), (ii) solve the IVP, and (iii) plot the solution of (a)-(b) for indicated time.

#### (a)
$u_t+\left(1+x^2\right)u_x=0, \, u(0,x)=\arctan{(x)}, \, t=1,2,3$, and
what is $\lim\limits_{t \to \infty}u(t,x)$?

$$\frac{dx}{dt}=1+x^2$$
$$\frac{1}{1+x^2}dx=dt$$

$$\arctan{(x)}=t+C$$
So $x=\tan{(t+C)}$ are the characteristic curves of $u_t+\left(1+x^2\right)u_x=0$.  
On each of the curves, $u(x,t)$ is constant because $$\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x} \sec^2{(t+c)}$$  
And since $1+\tan^2\theta=\sec^2\theta$, this means $\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x}\left( 1+\tan^2{(t+c)}\right)=u_t+\left(1+x^2\right)u_x$, which we know is 0.

Thus $u(t, \tan{(t+C)})=u(0, \tan{(0+C)})=u(0, C)$ is independent of $t$.  
Putting $x=\tan{(t+C)}$ and $C=\arctan{(x)}-t$, we have
$$u(t, x)=u(0,\arctan{(x)}-t)$$
It follows that $u(t,x)=f(\arctan{(x)}-t)$.  
And since we are given that $u(0,x)=\arctan{(x)}$, we have $f(\arctan{(x)}-0)=\arctan{(x)}$ so that $f(w)=w$ for any $w$, yielding our solution of

\begin{tcolorbox}[colback=white, title=Solution]
$$u(t,x)=\arctan{(x)}-t$$
\end{tcolorbox}

```{r, echo=FALSE}
C=1
df_4a <- tibble(x = seq(-2*pi, 2*pi, length.out = 1000),
                y=tan(x+C)) %>% 
  mutate(y = ifelse(abs(y) < 20, y, NA))

ggplot(df_4a, aes(x=x, y=y)) + 
  geom_line()+
  coord_cartesian(ylim=c(-10,10), xlim = c(-5,5))+
  theme_bw()+
  labs(x="t", y="x")+
  ggtitle("Characteristic curves for Problem 4a", subtitle = "with C=1")
```

Since the range of $\arctan$ is limited to $\left(\frac{-\pi}{2},\frac{\pi}{2}\right)$, as $\lim\limits_{t \to \infty}$ that means $u(t,x)=\arctan(x)-t$ will go to $-\infty$.

```{r, echo=FALSE}

x=seq(-10,10,length.out = 1000)
ggplot() + 
  geom_line(aes(x=x, y=atan(x)-1, color="1"))+
  geom_line(aes(x=x, y=atan(x)-2, color="2"))+
  geom_line(aes(x=x, y=atan(x)-3, color="3"))+
  theme_bw()+
  labs(x="x", y="u(x,t)", color="Value of t")+
  ggtitle("Solution curves for Problem 4a")
```

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
Since $u_t=-1, u_x=\frac{1}{1+x^2}$, that means $u_t+\left(1+x^2\right)u_x=-1+1=0$
\end{tcolorbox}

#### (b)
$u_t-xu_x=0, \, u(0,x)=\frac{1}{1+x^2}, \, t=1,2,3$, and
what is $\lim\limits_{t \to \infty}u(t,x)$?

The directional derivative of $u$ in the direction of the vector $(1, -x)$ is zero. The curves in the $tx$ plane with $(1, -x)$ as a tangent vector have slopes $-x$:

$$\frac{dx}{dt}=-x$$
$$\frac{1}{x}dx=-dt$$
Solving this ODE gives the equations for the characteristic lines:
$$x=Ce^{-t}$$

```{r, warning = FALSE, echo=FALSE}
curves_4b <- tibble(C=c(1:5), seq=list(x)) %>% 
  rowwise() %>% 
  mutate(y=list(C*exp(-seq))) %>% 
  unnest(c(seq,y))

ggplot(data=curves_4b)+
  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
  coord_cartesian(ylim=c(NA, 100), xlim=c(-5, 9))+
  labs(x="t", y="x", color="C")+
  theme_light()+
  ggtitle("Characteristic curves for Problem 4b")
```

On each of the curves, $u(x,t)$ is constant because $$\frac{d}{dt}u(t,Ce^{-t}) = \pdv {u}{t}-Ce^{-t}\pdv{u}{x}=u_t-xu_x$$  
and we know that $u_t-xu_x=0$.

Set $\xi=xe^t$ and $\eta=x$. Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
$$\pdv{u}{\eta}+\pdv{u}{\xi}e^t $$

and similarly
$$u_t= \pdv{u}{t}=\pdv{u}{\eta}\pdv{\eta}{t}+\pdv{u}{\xi}\pdv{\xi}{t}=\pdv{u}{\xi}xe^t $$

Which means that 
$$u_t-xu_x=\pdv{u}{\xi}xe^t -x\left(\pdv{u}{\eta}+\pdv{u}{\xi}e^t \right)=0$$
Assuming $x\neq0$, this means
$$u_\eta=0$$
$$u=\int u_\eta \, d\eta \> + f(\xi)=0+ f(\xi)$$

So we have $u(t,x)=f(\xi)=f(xe^t)$ and we know that $u(0,x)=\frac{1}{1+x^2}$, which means $f(xe^0)=f(x)=\frac{1}{1+x^2}$ and therefore
\begin{tcolorbox}[colback=white, title=Solution]
$$u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$u_t=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)xe^t$$
$$u_x=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)e^t$$
And we see that $u_t-xe_t$ indeed equals zero.
\end{tcolorbox}

Plotting the solution $u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$ as a function of $x$ for $t=1,2,3$, we get:

```{r, warning=FALSE, echo=FALSE}

x=seq(-10,10,length.out = 1000)
ggplot() + 
  geom_line(aes(x=x, y=1/(1+(exp(1)*x)^2), color="1"))+
  geom_line(aes(x=x, y=1/(1+(exp(2)*x)^2), color="2"))+
  geom_line(aes(x=x, y=1/(1+(exp(3)*x)^2), color="3"))+
  theme_bw()+
  labs(x="x", y="u(x,t)", color="Value of t")+
  xlim(-5,5)+
  ggtitle("Solution curves for Problem 4b")
```

We can see that $u(t,x)$ will equal 1 at $x=0$ regardless of the value of $t$, but at all other values of $x$, $\lim\limits_{t \to \infty}u(t,x)=0$. This can be written as
\begin{equation}
\lim\limits_{t \to \infty}u(t,x)=
    \begin{cases}
        1 & \text{if } x=0\\
        0 & \text{if } x\neq0
    \end{cases}
\end{equation}

#### (c)
$u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2, \, u(0,y)=e^y$

$\frac{dy}{dx}=x$
Solving this ODE yields the equation for the characteristic curves: $y=\frac{x^2}{2}+C$

```{r, warning = FALSE, echo=FALSE}
curves_4c <- tibble(C=c(1:5), seq=list(x)) %>% 
  rowwise() %>% 
  mutate(y=list(0.5*(seq)^2+C)) %>% 
  unnest(c(seq,y))

ggplot(data=curves_4c)+
  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
  coord_cartesian(ylim=c(NA, 45), xlim=c(-9.5, 9.5))+
  labs(x="x", y="y", color="C")+
  theme_light()+
  ggtitle("Characteristic curves for Problem 4c")
```

Now we define our new coordinate system as $\xi=y-\frac{x^2}{2}, \, \eta=x$

Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
$$\pdv{u}{\eta}-x\pdv{u}{\xi}$$

and similarly
$$u_y= \pdv{u}{y}=\pdv{u}{\eta}\pdv{\eta}{y}+\pdv{u}{\xi}\pdv{\xi}{y}=$$
$$\pdv{u}{\xi}$$

$$u_x+xu_y=\pdv{u}{\eta}-x\pdv{u}{\xi}+x\pdv{u}{\xi}=\pdv{u}{\eta}$$

And since we know $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, this means we have
$$\pdv{u}{\eta}=\left(y-\frac{x^2}{2}\right)^2=\xi^2$$
Now we integrate with respect to $\eta$ to find
$$u=\eta\xi^2+F(\xi)$$

At $u(0,y)$, this becomes $u=0y^2+F(y)$, or $u(0,y)=F(y)$.
On the other hand, we are told that $u(0,y)=e^y$, which means that $F(y)=e^y$.
$$u=\eta\xi^2+e^\xi$$
Substitute back in $x$ and $y$:
\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=x\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}$$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$u_x=-2x^2\left(y-\frac{x^2}{2}\right)+\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}(-x)$$
$$u_y=2x\left(y-\frac{x^2}{2}\right)+\exp{\left(y-\frac{x^2}{2}\right)}$$
Which means that $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, as we wanted.
\end{tcolorbox}

#### (d)
$2xu_x+(x+1)u_y=y$, for $x>0, u=2y$ on $x=1$

$$\frac{dx}{dy}=\frac{x+1}{2x}=\frac{1}{2}+\frac{1}{2x}$$
Solving this ODE yields the characteristic curves:
$$y=\frac{x}{2}+\frac{\ln(x)}{2}+C$$
$$C=2y-x-\ln(x)$$

```{r, warning = FALSE, echo=FALSE}
curves_4d <- tibble(C=c(1:5), seq=list(x)) %>% 
  rowwise() %>% 
  mutate(y=list((seq/2)+(log(seq)/2)+C)) %>% 
  unnest(c(seq,y))

ggplot(data=curves_4d)+
  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
  coord_cartesian(ylim=c(-1, 12), xlim=c(-1, 9.5))+
  labs(x="x", y="y", color="C")+
  theme_light()+
  ggtitle("Characteristic curves for Problem 4d")
```

Then set $\xi=x+3y, \; \eta=x-y$, giving us

$$\partial_x= \pdv{}{\xi} \pdv{\xi}{x}+\pdv{}{\eta} \pdv{\eta}{x}=\partial_{\xi}+\partial_{\eta}$$
$$\partial_y= \pdv{}{\xi} \pdv{\xi}{y}+\pdv{}{\eta} \pdv{\eta}{y}= 3\partial_{\xi}-\partial_{\eta}$$
Which means that 
$$\partial_{x}+3 \partial_{y}= 10 \partial_{\xi}-2\partial_{\eta}$$

$$\partial_{x}-\partial_{y}=-2\partial_{\xi}+2\partial_{\eta} $$

Which means we can rewrite the original equation as 
$$\left( \partial_{x}+3\partial_{y}\right) \left(\partial_{x}-\partial_{y}\right)u=0 \quad\longrightarrow\quad -8\partial$$
\begin{equation}\label{eq:sphere1}
x=r \sin{\theta} \cos{\phi}, \quad y=r \sin{\theta} \sin{\phi}, \quad z=r \cos{\theta}
\end{equation}

\begin{equation}\label{eq:sphere2}
r=\sqrt{x^2+y^2+z^2}, \quad \theta=\arccos{(\frac{z}{r})}, \quad \phi=\arctan{(\frac{y}{x})} \end{equation}

$$\theta=\arccos{(\frac{z}{r})}$$

Using the chain rule:
$$\Delta u= \frac{{\partial}^{2}f}{\partial x^{2}}+\frac{{\partial}^{2}f}{\partial y^{2}}+\frac{{\partial}^{2}f}{\partial z^{2}}$$

Now we find the following values from (\ref{eq:sphere1}):
$$\frac{\partial r}{\partial x}=\sin(\theta)\cos(\phi),
\qquad\frac{\partial\theta}{\partial x}=\frac{1}{r}\cos(\theta)\cos(\phi),
\qquad\frac{\partial\phi}{\partial x}=-\frac{1}{r}\frac{\sin(\phi)}{\sin(\theta)}$$
$$\frac{\partial r}{\partial y}=\sin(\theta)\sin(\phi),\qquad\frac{ \partial\theta}{\partial y}=\frac{1}{r}\cos(\theta)\sin(\phi),\qquad\frac{\partial\phi}{\partial y}=\frac{1}{r}\frac{\cos(\phi)}{\sin(\theta)} $$

$$\frac{\partial r}{\partial z}=\cos(\theta),\qquad\frac{\partial \theta}{\partial z}=-\frac{1}{r}\sin(\theta),\qquad\frac{\partial\phi}{\partial z}=0$$
$$ \frac{\partial u}{\partial x} = \frac{\partial u}{\partial r} \frac{\partial r}{\partial x}+\frac{\partial u}{\partial\theta}\frac{\partial\theta}{\partial x}+\frac{\partial u}
{\partial\phi}\frac{\partial\phi}{\partial x}=u_r\sin(\theta)\cos(\phi)$$

$$\frac{\partial u_x}{\partial x} = \frac{\partial u_x}{\partial r} \frac{\partial r}{\partial x}+\frac{\partial u_x}{\partial\theta}\frac{\partial\theta}{\partial x}+\frac{\partial u_x}{\partial\phi}\frac{\partial\phi}{\partial x}= $$

$$u_{rr}\sin^2(\theta)\cos^2(\phi)+ \frac{u_r}{r} \cos^2(\theta)\cos^2(\phi)- $$

$$u(x,t)= \frac{\varphi(x+ct)+\varphi(x-ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(s)ds}$$
Assume $c$ is not a constant, but instead a function of $x$.

Since $u=\frac{w}{r}$, this means we have
\begin{equation}u(t,r)=
    \begin{cases} 
      \frac{1}{2cr} \int_{r-ct}^{r+ct}{s\;ds} & r-ct \geq -1, r+ct \leq 1\\
      \\
      \frac{1}{2cr} \int_{-1}^{1}{s\;ds} & r-ct < -1, r+ct > 1\\
      \\
      \frac{1}{2cr} \int_{-1}^{r+ct}{s\;ds} & r-ct < -1, r+ct \leq 1\\
      \\
      \frac{1}{2cr} \int_{r-ct}^{1}{s\;ds} & r-ct \geq -1, r+ct > 1\\

    \end{cases}       
\end{equation}

On (characteristic) lines with the slope $y=-x+c$, or $y+x=\text{constant}$, we must have $u_x-u_y=f(3x-y)=0$.
Set $\eta=x+y$ and $\xi=x$. 
Then by the chain rule,
$$u_x= u_{\eta}+u_{\xi}, \; u_y=u_{\eta}$$
And let's rewrite $y$ as $y=\eta-x=\eta-\xi$.

So
$$u_x-u_y=f(3x-y) \quad\longrightarrow\quad u_{\xi}=f(3\xi-\eta+\xi)$$

$$u_{\xi}=f(4\xi-\eta)$$
Now integrate with respect to $\xi$:
$$u(\eta, \xi)=F(4\xi-\eta)+G(\eta)$$
where $F$ is the antiderivative of $f$ with respect to $\xi$.

## HW 2 Problem 6

Does the D’Alembert method work for the wave equation $u_{tt} = c(x)^2u_{xx}$? What about $u_{tt} = c(t)^2u_{xx}$? Why?

Let's try the factorization method if $c=c(x)$:
$$\left( \partial_{t}+c(x)\partial_{x}\right) \left(\partial_{t}-c(x)\partial_{x}\right)u=0$$
Set $\xi=x+c(x)t$ and $\eta=x-c(x)t$.

By the chain rule,

$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}$$

$$=\partial_{\xi}(1+c')+\partial_{\eta}\,(1-c')$$

$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}$$
$$=c\partial_{\xi}-c\partial_{\eta}$$
When we plug these back into
$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
we get

$$ \left(c\partial_{\xi}-c\partial_{\eta}+c\partial_{\xi}(1+c')+c\partial_{\eta}\,(1-c')\right)
\left(c\partial_{\xi}-c\partial_{\eta} -c\partial_{\xi}(1+c')-c\partial_{\eta}\,(1-c')\right)u=0$$

When $c$ is constant, $c'=0$ and this simplifies to $-4c^2 u_{\xi\eta}=0$, allowing us to integrate to find $u(x,t)=F(x+ct)+G(x-ct)$. But we cannot do the same simplification when we have the $c'$ terms, meaning we cannot solve the wave equation by the same method when the wave speed is not constant.

Similarly,

Let's try the factorization method if $c=c(t)$:
$$\left( \partial_{t}+c(t)\partial_{x}\right) \left(\partial_{t}-c(t)\partial_{x}\right)u=0$$
Set $\xi=x+c(t)t$ and $\eta=x-c(t)t$.

By the chain rule,

$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}=
\partial_{\xi}+\partial_{\eta}$$

$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}=\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)
$$

When we plug these back into
$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
we get

$$\left( \partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)+c\partial_{\xi}+c\partial_{\eta} \right)
\left(\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)-c\partial_{\xi}-c\partial_{\eta}\right)u=0$$
Where again we cannot cancel terms to simplify our equation as we could with constant $c$.

<!-- paste0("(", "x[0]","','","t[0]", ")")) -->

### HW5

Then the time problem has repeated real roots $r=\frac{-a}{2}$, giving us $T(t)=C_1 e^{-a/2}+C_2te^{-a/2}$.

Now we can return to our solution (\ref{eqn:teleSol}). Since our new IBVP (\ref{eqn:teleW}) has $\phi(x)=\psi(x)=0$, our coefficients $A_n$ and $B_n$ will be zero everywhere, which means $w(t,x)=0$ and therefore $u=v$, so any solution to (\ref{eqn:tele}) is in fact the unique solution.

$$T''+\lambda T=0 $$

$$ r^2+\lambda=0$$

$(r^2-\beta^2)(r^2+\beta^2)=(r+\beta)(r-\beta)(r+\beta i)(r-\beta i)$.

\underline{Case 3:} Negative eigenvalues, $\lambda = -\beta^4 <0$
This case gives us the characteristic equation: $r^4+\beta^4=0$, or $r^2= \pm \beta^2i$.

From $e^{i\theta}=\cos{\theta}+i\sin{\theta}$, we get that $e^{\frac{i\pi}{2}}=i$. Thus, $$r^2=\beta^2e^{\frac{i\pi}{2}}, r^2=\beta^2e^{\frac{-i\pi}{2}}$$
$$r=\beta e^{\frac{i\pi}{4}},\;-\beta e^{\frac{i\pi}{4}},\;\beta e^{\frac{-i\pi}{4}},\;-\beta e^{\frac{-i\pi}{4}} $$
$e^{\frac{\pi}{4}i}=\cos{\frac{\pi}{4}}+i\sin{\frac{\pi}{4}}= \frac{\sqrt{2}}{2}+i\frac{\sqrt{2}}{2}$

$e^{ \frac{-\pi}{4}i}=\cos{\frac{-\pi}{4}}+i\sin{\frac{-\pi}{4}}= \frac{\sqrt{2}}{2}-i\frac{\sqrt{2}}{2}$

$$r = \frac{\beta\sqrt{2}}{2} \pm i\frac{\beta\sqrt{2}}{2}, \frac{-\beta\sqrt{2}}{2} \pm i\frac{\beta\sqrt{2}}{2} $$

Our solutions would then take the form
$$X(x)=e^{\frac{-\beta\sqrt{2}}{2}}\left[A\cos{\left(\frac{\beta\sqrt{2}}{2} x\right)}+B\sin{\left(\frac{\beta\sqrt{2}}{2} x\right)}\right] + e^{\frac{\beta\sqrt{2}}{2}}\left[C\cos{\left(\frac{\beta\sqrt{2}}{2} x\right)}+D\sin{\left(\frac{\beta\sqrt{2}}{2} x\right)}\right]$$

$$X(0)=e^{\frac{-\beta\sqrt{2}}{2}}A+e^{\frac{\beta\sqrt{2}}{2}}C=0 $$
$$A=-e^{\beta\sqrt{2}}C $$
Let $\frac{\beta\sqrt{2}}{2}=\alpha$

$$X(x)=e^{-\alpha}\left[-Ce^{2\alpha}\cos{\left(\alpha x\right)}+B\sin{\left(\alpha x\right)}\right] + e^{\alpha}\left[C\cos{\left(\alpha x\right)}+D\sin{\left(\alpha x\right)}\right]$$

$$X(x)=e^{-\alpha}B\sin{\left(\alpha x\right)}-e^{-\alpha} e^{2\alpha}C \cos{\left(\alpha x\right)} + e^{\alpha}C\cos{\left(\alpha x\right)}+e^{\alpha}D\sin{\left(\alpha x\right)}$$
$$X(x)=e^{-\alpha}B\sin{\left(\alpha x\right)}+e^{\alpha}D\sin{\left(\alpha x\right)}$$

$$X(L)=e^{-\alpha}B\sin{(\alpha L)}+e^{\alpha}D\sin{(\alpha L)}=0$$
$$X''(L)=-\alpha^2e^{-\alpha}B\sin{(\alpha L)}-\alpha^2e^{\alpha}D\sin{(\alpha L)}=0$$

Assume $\sin{(\alpha L)} \neq 0$.

$$e^{-\alpha}B+e^{\alpha}D=0 \quad \Longrightarrow\quad B=-De^{2\alpha}$$
$$X(x)=-De^{2\alpha}e^{-\alpha}\sin{\left(\alpha x\right)}+e^{\alpha}D\sin{\left(\alpha x\right)}$$

Then $X(x)=0$ for all $x$, which is just the trivial solution that doesn't satisfy our ICs. So we must have $\sin{(\alpha L)}=0, \alpha=\frac{\beta\sqrt{2}}{2}=\frac{n \pi}{L}, \beta=\frac{2 n \pi}{\sqrt{2}L}$

<!-- \underline{Case 3:} Positive eigenvalues, $\lambda = \beta^4 >0$ -->

<!-- This case gives us the characteristic equation: $r^4+\beta^4=0$.  -->
<!-- $$(r^2+\beta^2i)(r^2-\beta^2i)=(r+\beta\sqrt{i})(r-\beta\sqrt{i})(r+\beta i^{3/2})(r-\beta i^{3/2})$$ -->
<!-- $$X(x) = A e^{(-1)^{3/4} \beta x} + B e^{-(-1)^{3/4} \beta x} + C e^{(-1)^{1/4} \beta x} + D e^{-(-1)^{1/4} \beta x} $$ -->

<!-- $$X(L) = A e^{(-1)^{3/4} \beta L} + B e^{-(-1)^{3/4} \beta L} + C e^{(-1)^{1/4} \beta L} + D e^{-(-1)^{1/4} \beta L}=0 $$ -->
<!-- $$A+B+C+D=0 $$ -->

<!-- $$X''(x) = -iA\beta^2 e^{(-1)^{3/4} \beta x} -iB\beta^2 e^{-(-1)^{3/4} \beta x} + iC\beta^2 e^{(-1)^{1/4} \beta x} + iD\beta^2 e^{-(-1)^{1/4} \beta x}=0 $$ -->
<!-- $$i\beta^2(-A-B+C+D)=0 $$ -->
<!-- $A+B=C+D$ -->

<!-- $$-A e^{(-1)^{3/4} \beta L} -Be^{-(-1)^{3/4} \beta L} + C e^{(-1)^{1/4} \beta L} + D e^{-(-1)^{1/4} \beta L}=0 $$ -->

\underline{Case 2:} Negative eigenvalues, $\lambda = -\beta^2 <0$
This case gives us the characteristic equation: $r^2-\beta^2=0$. Since our characteristic equation has distinct real roots $r= \pm \beta$, our solutions take the form 
$$T(t)=Ae^{\beta t}+Be^{-\beta t}$$

$\beta A - \beta B = \psi(x)$
$A+B=\phi(x)$