### 2024-04-15 PDEs/math245_hw6.qmd
circles for pdes

---
title: "MATH 245 Homework 6"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-04-14"
format: 
  pdf:
    fig-width: 4
    fig-height: 3
execute: 
  echo: false
---

```{r}
#| label: load_packages
#| warning: false
#| message: false

library(tidyverse)
```

## Done: 1-4
## Not started: 5-9

## Problem 1

The function $u: D \subset \mathbb{R^2} \rightarrow \mathbb{R}$ is subharmonic if $\Delta u = u_{xx}+u_{yy} \geq 0$ and superharmonic if $\Delta u = u_{xx}+u_{yy} \leq 0$. Does the maximum and minimum principle hold for subharmonic functions or superharmonic on a connected bounded domain $D \subset \mathbb{R^2}$? If yes, state and prove it, otherwise give a counterexample.

### Subharmonic

**Minimum - No**  
Take $u(x,y)=x^2$ on the connected bounded domain $D=\{(x,y):-1\leq x \leq 1, -1\leq y \leq 1\}$. Then $\Delta u = 2 >0$, so $u$ is subharmonic. However, the minimum of $u$ on $D$ occurs at $x=0$ where $u(0,y)=0$ and not on the boundary of $D$, where $u=1$.

```{r}
#| label: counterexample1
#| fig-align: "center"

x<- seq(-1,1, by=0.1)
y<-x^2
ggplot()+
  geom_line(aes(x=x,y=y))+
  theme_light()
```

**Maximum - Yes**

Let $D$ be a bounded domain in $\mathbb{R^2}$ and $u: D \subset \mathbb{R^2} \rightarrow \mathbb{R}$. Assume $u$ is continuous on $\bar{D}$, where $\bar{D}=D \cup \partial D$, and is a continuous subharmonic function such that $\Delta u = u_{xx}+u_{yy} \geq 0$.

Then because $D$ is bounded and $u$ is continuous on $\bar{D}$, $u$ must attain its maximum and minimum in $\bar{D}$. Define $M=\max_{\partial D} u=u(x_0)$ on $\partial D$. Our goal is to show that $\max_{D}u \leq \max_{\partial D}u=M$.

For every $\varepsilon >0$, we define $v(x)=u(x)+\varepsilon \lvert x \rvert^2$. Then
\begin{align*}
\Delta v& = \Delta u + \varepsilon \Delta \lvert x \rvert^2 \\
        &= \Delta u+\varepsilon \sum_{i=1}^{n} \frac{\partial^2 \Delta \lvert x\rvert^2}{\partial x_{i}^2} \\
        &= \Delta u+ \varepsilon(2+2+...+2) \\
        & = \Delta u+ 2n\varepsilon
\end{align*}

And since $u$ is subharmonic, we have that
\begin{equation}\label{eqn:max_princ}
\Delta v = \Delta u+ 2n\varepsilon > 0
\end{equation}

Now assume the max of $v$ is inside $D$, at some $x_1 \in D$. Then by (\ref{eqn:max_princ}), we have $\Delta v(x_1)>0$. By generalization of the second derivative theorem, we know that at the maximum of a function $v:\mathbb{R^n} \rightarrow \mathbb{R}$, the Hessian of $v$, $\nabla^2 v(x_1)$, should be negative semi-definite, meaning that all eigenvalues $\lambda_i$ are less than or equal to zero. But $\lambda_i \leq 0$ implies that
$$\Delta v(x_1)=\text{tr}\left(\nabla^2 v(x_1)\right)=(\lambda_1+\lambda_2+...+\lambda_n)(x_1) \leq 0$$
which is a contradiction of (\ref{eqn:max_princ}). Therefore, the maximum of $v$ is at the boundary,
$$v(x) \leq \max_{\partial D}v \quad \forall \;x \in D $$
So, $\forall \;x \in D,$
\begin{align*}
u(x)&=v(x)-\varepsilon\lvert x\rvert^2 \\
&\leq \max_{\partial D}v-\varepsilon\lvert x\rvert^2
\end{align*}
But since we can write
\begin{align*}
\max_{\partial D}v-\varepsilon\lvert x\rvert^2& = \max_{\partial D} \left(u(x)+\varepsilon\lvert x\rvert^2  \right)-\varepsilon\lvert x\rvert^2 \\
&= \max_{\partial D} u+\varepsilon\left(R-\lvert x\rvert^2 \right),
\end{align*}
where $R$ is the largest radius of the domain, we must have
$$u(x)\leq \max_{\partial D} u+\varepsilon\left(R-\lvert x\rvert^2 \right). $$
Since this is true for any $\varepsilon$, we take $\varepsilon \rightarrow 0$ to find that 
$u(x) \leq \max_{\partial D} u.\quad \square$

### Superharmonic

**Maximum - No**  
Take $u(x,y)=-x^2$ on the connected bounded domain $D=\{(x,y):-1\leq x \leq 1, -1\leq y \leq 1\}$. Then $\Delta u = -2 < 0$, so $u$ is superharmonic. However, the maximum of $u$ on $D$ occurs at $x=0$ where $u(0,y)=0$ and not on the boundary of $D$, where $u=-1$.

```{r}
#| label: counterexample2
#| fig-align: "center"

y<-(-1)*(x^2)
ggplot()+
  geom_line(aes(x=x,y=y))+
  theme_light()
```

**Minimum - Yes**

Let $D$ be a bounded domain in $\mathbb{R^2}$ and $u: D \subset \mathbb{R^2} \rightarrow \mathbb{R}$. Assume $u$ is continuous on $\bar{D}$, where $\bar{D}=D \cup \partial D$, and is a continuous superharmonic function such that $\Delta u = u_{xx}+u_{yy} \leq 0$.

Then because $D$ is bounded and $u$ is continuous on $\bar{D}$, $u$ must attain its maximum and minimum in $\bar{D}$. Define $m=\min_{\partial D} u=u(x_0)$ on $\partial D$. Our goal is to show that $\min_{D}u \geq \min_{\partial D}u=m$. For every $\varepsilon >0$, we define $v(x)=u(x)-\varepsilon \lvert x \rvert^2$. Then
\begin{align*}
\Delta v& = \Delta u - \varepsilon \Delta \lvert x \rvert^2 \\
        &= \Delta u-\varepsilon \sum_{i=1}^{n} \frac{\partial^2 \Delta \lvert x\rvert^2}{\partial x_{i}^2} \\
        &= \Delta u- \varepsilon(2+2+...+2) \\
        & = \Delta u- 2n\varepsilon
\end{align*}
And since $u$ is superharmonic, we have that
\begin{equation}\label{eqn:min_princ}
\Delta v = \Delta u- 2n\varepsilon <0
\end{equation}

Now assume the min of $v$ is inside $D$, at some $x_1 \in D$. Then by (\ref{eqn:min_princ}), we have $\Delta v(x_1)<0$. By generalization of the second derivative theorem, we know that at the minimum of a function $v:\mathbb{R^n} \rightarrow \mathbb{R}$, the Hessian of $v$, $\nabla^2 v(x_1)$, should be positive semi-definite, meaning that all eigenvalues $\lambda_i$ are greater than or equal to zero. But $\lambda_i \geq 0$ implies that
$$\Delta v(x_1)=\text{tr}\left(\nabla^2 v(x_1)\right)=(\lambda_1+\lambda_2+...+\lambda_n)(x_1) \geq 0$$
which is a contradiction of (\ref{eqn:min_princ}). Therefore, the minimum of $v$ is at the boundary,
$$v(x) \geq \min_{\partial D}v \quad \forall \;x \in D $$
So, $\forall \;x \in D,$
\begin{align*}
u(x)&=v(x)+\varepsilon\lvert x\rvert^2 \\
&\leq \min_{\partial D}v+\varepsilon\lvert x\rvert^2
\end{align*}
But since we can write
\begin{align*}
\min_{\partial D}v+\varepsilon\lvert x\rvert^2& = \min_{\partial D} \left(u(x)-\varepsilon\lvert x\rvert^2  \right)+\varepsilon\lvert x\rvert^2 \\
&= \min_{\partial D} u+\varepsilon\left(\lvert x\rvert^2 -R\right),
\end{align*}

where $R$ is the largest radius of the domain, we must have
$$u(x) \leq \min_{\partial D} u+\varepsilon\left(\lvert x\rvert^2-R \right). $$
Since this is true for any $\varepsilon$, we take $\varepsilon \rightarrow 0$ to find that 
$u(x) \geq \min_{\partial D} u.\quad \square$

## Problem 2

Suppose $u$ is harmonic on the disk $D=\{(x, y):x^2+y^2< 4\}$ and $u = \sin{(\theta)} + 1$ on
$\partial D$. Without finding the solution $u$, find the maximum value of $u$ in $D \cup \partial D$ and value of $u$ at the origin.

By the maximum principle, $u$ obtains its maximum in $D \cup \partial D$ on the boundary $\partial D$ and not inside. The maximum of $u=\sin{\theta}+1$ on $\partial D$ is at $\theta=\frac{\pi}{2}$, where $\sin{\theta}$ is at its maximum. Therefore, the maximum value of $u$ in $D \cup \partial D$ is $u=\sin{\left(\frac{\pi}{2}\right)}+1=1+1=2$.

The mean value property says that for a harmonic function $u$ in a disk $D$, continuous in its closure $\bar{D}$, the value of $u$ at the center of $D$ equals the average of $u$ on its circumference. The average value of a continuous function $f(x)$ over the interval $[a,b]$ is given by 
$$f_{avg}(x)=\frac{1}{b-a}\int_a^b f(x)\;dx$$
Here,
\begin{align*}
u_{avg}(\theta)& =\frac{1}{2\pi}\int_0^{2\pi} (\sin\theta+1)\;d\theta \\
&=\frac{1}{2\pi}\Bigl[(\cos\theta+\theta) \Bigr]_0^{2\pi} \\
&=\frac{1}{2\pi}\Bigl[(\cos{2\pi}+2\pi)-\cos{0} \Bigr] \\
&= \frac{1}{2\pi}(1+2\pi-1)=1
\end{align*}
Therefore, the value of $u$ at the origin is 1.

## Problem 3

Use the maximum principle to show that the solution of the Dirichlet problem for $\Delta u=0$ depends continuously on the boundary data.

Define $u_1$ and $v$ to be two solutions to the Dirichlet problems
\begin{equation*}
    \begin{cases}
     \Delta u_1 =0         & \text{in }D \\
      u_1=h(x)       & \text{on }\partial D
    \end{cases}       
\end{equation*}

\begin{equation*}
    \begin{cases}
     \Delta u_2 =0         & \text{in }D \\
      u_2=h(x)+\varepsilon(x)      & \text{on }\partial D
    \end{cases}       
\end{equation*}

where $\varepsilon(x)$ is small $\forall x$ on $\partial D$. Now let $w=u_1-u_2$. Then $w$ solves the following Dirichlet problem:
\begin{equation*}
    \begin{cases}
     \Delta w =0         & \text{in }D \\
      w=\varepsilon(x)       & \text{on }\partial D
    \end{cases}       
\end{equation*}

By the maximum and minimum principles,
$$\min_{\partial D}\varepsilon(x)\leq w\leq \max_{\partial D}\varepsilon(x)$$
If $\varepsilon \rightarrow 0$ for all $x$ on $\partial D$, then $w=u_1-u_2$ approaches zero for all $x \in D$.

## Problem 4

Solve the following problem on semi-annulus domain:
\begin{equation}\label{eqn:semi}
    \begin{cases}
     u_{rr}+\frac{1}{r}u_r+ \frac{1}{r^2}u_{\theta\theta}=0  & 0<a<r<b, 0<\theta<\pi \\
     u(r,0)=u(r,\pi)=0       & 0<a<r<b \\
    u_r(a, \theta)=\theta       &  0<\theta<\pi\\
     u_r(b, \theta)=0      &  0<\theta<\pi\\
    \end{cases}       
\end{equation}

We look for a separated solution of the form $u=\Theta(\theta)R(r)$. First, we rewrite $u_{rr}+\frac{1}{r}u_r+ \frac{1}{r^2}u_{\theta\theta}=0$ as 
$$R''\Theta + \frac{1}{r}R'\Theta+\frac{1}{r^2}R\Theta''=0$$

Separating variables, we get
$$\frac{R''}{R}r^2 + \frac{R'}{R}r=-\frac{\Theta''}{\Theta}$$
Since the left-hand side is a function of only $R$ and the right-hand side is a function of only $\Theta$, we can set both sides equal to a constant $\lambda$ and separate this into two problems, one in $\Theta$ and one in $R$. Start with $\theta$:
$$\Theta''+\Theta\lambda=0$$
From our boundary conditions, 
$$R(r)\Theta(0)=R(r)\Theta(\pi)=0 \quad\Longrightarrow\quad  \Theta(\pi)=\Theta(0)=0$$
So we have an eigenvalue problem with homogenous Dirichlet BCs, which we have shown to have the eigenvalues and eigenvectors
$$\lambda_n= \left(\frac{n\pi}{l}\right)^2, \quad X_n(x)=\sin{\left(\frac{n\pi x}{l}\right)},\quad n=1,2,3...$$
Here, that means
$$\lambda_n=n^2, \quad X_n(x)=\sin{(n\theta)},\quad n=1,2,3... $$

Now onto the $R$ problem:
$$\frac{R''}{R}r^2 + \frac{R'}{R}r=\lambda$$
$$R''r^2 + R'r-\lambda R=0$$
This is an Euler ODE of the form $ax^2y''+bxy'+cy=0$, which has the characteristic equation $am(m-1)+bm+c$ and, if $m_1, m_2$ are distinct real roots, the general solution $y=. (Note that I have used $m$ instead of the standard $r$ to avoid confusion with the variable $r$ from (\ref{eqn:semi}).

Here, $a=1, b=1, c=-\lambda$, so the characteristic equation becomes $m(m-1)+m-\lambda$, which simplifies to $m^2-\lambda$. Thus, our roots are $\pm \sqrt{\lambda}=\pm n$. Thus, $R(r)=Cr^n+Dr^{-n}$.

Our separated solutions thus take the form
$$u_n(r,\theta)=\left(Cr^n+Dr^{-n}\right)\sin{(n\theta)}, \quad n=1,2,3... $$
$$u(r,\theta)=\sum_{n=1}^{\infty} \left(Cr^n+Dr^{-n}\right)\sin{(n\theta)} $$

Now we use our boundary conditions $u_r(a, \theta)=\theta$ and $u_r(b, \theta)=0$:
$$u_r(r,\theta)=\sum_{n=1}^{\infty} \left(nCr^{n-1}-nDr^{-n-1}\right)\sin{(n\theta)} $$
$$u_r(b,\theta)=0=\sum_{n=1}^{\infty} \left(nCb^{n-1}-nDb^{-n-1}\right)\sin{(n\theta)} $$
$$Cb^{n}=Db^{-n} \rightarrow C=Db^{-2n}$$
$$u_r(a,\theta)=\theta=\sum_{n=1}^{\infty} \left(nDb^{-2n}a^{n-1}-nDa^{-n-1}\right)\sin{(n\theta)} $$
$$\frac{nD}{a}\left(b^{-2n}a^{n}-a^{-n}\right)=\int_0^\pi \theta\sin{(n\theta)}d\theta$$
$$\int_0^\pi \theta\sin{(n\theta)}d\theta=\theta(-1/n\cos{n\theta})\Bigr\rvert^\pi_0=\pi(-1/n)(-1)^n=\frac{\pi(-1)^{n+1}}{n}$$
$$\frac{nD}{a}\left(b^{-2n}a^{n}-a^{-n}\right)=\frac{\pi(-1)^{n+1}}{n}$$

Therefore, our solution to (\ref{eqn:semi}) is
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
$$u(r,\theta)=\sum_{n=1}^{\infty} \left(Cr^n+Dr^{-n}\right)\sin{(n\theta)}$$
where 
$$D=\frac{a\pi(-1)^{n+1}}{n^2\left(b^{-2n}a^{n}-a^{-n}\right)}$$
$$C=Db^{-2n}$$
\end{tcolorbox}

## Problem 5

Consider the Neumann problem with periodic BC:
\begin{equation}\label{eqn:prob5}
    \begin{cases}
     \Delta u=0 & r^2>a^2 \\
     u_r(a,\theta)=g(\theta)     & r^2=a^2
    \end{cases}       
\end{equation}

Find the necessary condition for the existence of a solution and then solve the problem. Hint: Use Green's identity
$$\iint\limits_{D}(u\Delta v-v\Delta u)\,dA=\int\limits_{\partial D} \left(u\frac{\partial v}{\partial n} - v\frac{\partial u}{\partial n}\right)\,ds $$
where $D$ is a domain and $n$ is a unit normal vector. Note that for a 2-dimensional domain $D$, $dA=dx\,dy$ and $ds$ is arc length.

## Problem 6.

Prove the uniqueness of the Neumann problem on $D \subset \mathbb{R}$
\begin{equation}\label{eqn:unique}
    \begin{cases}
     \Delta u=f  & \text{in }D \\
     \frac{\partial u}{\partial n}=g    & \text{on }\partial D
    \end{cases}       
\end{equation}

up to a constant. Use the following energy and Green's identity:
$$E(u)=\frac{1}{2}\int_D\lvert \nabla u\rvert^2\,dA$$

Suppose we have two solutions $u$ and $v$ to (\ref{eqn:unique}) and define $w=u-v$. Then we have a new Neumann problem
\begin{equation}\label{eqn:unique2}
    \begin{cases}
     \Delta w=0  & \text{in }D \\
     \frac{\partial w}{\partial n}=0   & \text{on }\partial D
    \end{cases}       
\end{equation}