### 2024-04-29 PDEs/pres.Rmd
presentation

---
title: "PDE pres"
author: "Ruby Krasnow"
date: "2024-04-27"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]

bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

We represent the the population density of *B. violaceus* at time $t$ and spatial coordinate $x$ as $u(t,x)$.

Assume that the population occupies a domain of fixed length $L$, so that $0 < x < L$.
The environment outside the domain is hostile, so there is no population there: if
$x\leq 0$ or $x \geq L$, then $u(x,t)= 0$ for all $t$.

If we assumed simple linear population growth, our PDE would take the form

\begin{equation}\label{eqn:lin_PDE}
\frac{\partial{u}}{\partial{t}}=\text{D}\frac{\partial{u}^2}{\partial{x^2}}+ru, \quad 0<x<L
\end{equation}

with the Dirichlet BCs $u(0,t)=u(0,L)=0$.

This is the simplest version of the general reaction-diffusion model, for a homogeneous, unstructured population growing exponentially and dispersing in a uniform, one-dimensional environment [@andow1990; @skellam1951].

We know how to solve this! The solution is
$$u(t,x)= \sum_{n=1}^{\infty} C_n e^{(r-\lambda_n)t} \sin{\left(\frac{n\pi x}{L}\right)}, $$
where

$$\lambda_n= \left(\frac{n\pi}{L}\right)^2D,\quad n=1,2,3... $$
and the coefficients $C_n$ are determined by the initial density $u(x,0)$.

$$\lambda_1= \left(\frac{\pi}{L}\right)^2D \quad \Longrightarrow \quad r= \left(\frac{\pi}{L}\right)^2D \quad \Longrightarrow \quad L_{cr}=\pi \sqrt{\frac{D}{r}}$$

A more realistic model considers the evolution of a compact initial
population distribution in a large open space. Rather than imposing the restriction that the population always occupy a fixed region, we instead allow the population to grow freely. Only the
initial condition is bounded, assumed for simplicity to be connected:
\begin{align}\label{eqn:IC_1} u(0,x) =
\begin{cases}
\phi(x)>0 & 0<x<L \\
0 & \text{otherwise}
\end{cases}
\end{align}

Here, $L$ is the size of the initially occupied area and the function $\phi$ describes the
initial population distribution. The equation describing the population dynamics,
that is,

\begin{equation}\label{eqn:PDE}
\frac{\partial{u}}{\partial{t}}=\text{D}\frac{\partial{u}^2}{\partial{x^2}}+F(u)
\end{equation}

should now be defined on a much larger domain than $L$; here, we assume $-\infty<x<\infty$. Note that $F(u)$ can now be a nonlinear function (like logistic growth), which is much more biologically realistic.

For biologically reasonable choices of $F(u)$, the growth function will be bounded by the linear function, such that $F(u) \leq ru$.

\begin{equation}\label{eqn:comp1}
\text{D}\frac{\partial{u}^2}{\partial{x^2}}+F(u) \leq \text{D}\frac{\partial{U}^2}{\partial{x^2}}+rU
\end{equation}

Now, note that for compact initial conditions like (\ref{eqn:IC_1}), there exist positive constants $G$ and $\sigma$ such that we bound the initial distribution from above by a Gaussian distribution: \begin{equation}\label{eqn:initGaus}
U(0,x)=\frac{G}{\sqrt{4\pi\sigma^2}}\exp{\left(-\frac{x^2}{4\sigma^2}\right)}
\end{equation}

\begin{equation}\label{eqn:comp2} 
\text{such that}\quad u(0,x) \leq U(0,x)\quad \text{for all }x
\end{equation}

\begin{equation}
\frac{\partial{U}}{\partial{t}}=\text{D}\frac{\partial{U}^2}{\partial{x^2}}+rU
\end{equation}

The linear equation in (1) can be readily solved, since the change of variables $U(t,x)= W(t,x)e^{rt}$ reduces it to the standard diffusion equation for $W$. Using forward and inverse Fourier transforms, we find that
$$W(t,x)=\frac{G}{\sqrt{4\pi(\sigma^2+Dt)}}\exp{\left(-\frac{x^2}{4(\sigma^2+Dt)}\right)}  $$

Converting back to our original variable, we conclude that the solution to (1) with the initial condition (5a) is a normal distribution, specifically
$$U(t,x)=\frac{G}{\sqrt{4\pi(\sigma^2+Dt)}}\exp{\left(-\frac{x^2}{4(\sigma^2+Dt)}+rt\right)}$$

Now, we can quantify the spatial spread of the population by considering how quickly a level set $U(x,t)=U_c$ moves in space, where $U_c$ could be considered a threshold of detection. Rearranging the above equation, we can get the following expression
$$\left(\frac{x}{t}\right) ^2=4rD+\frac{4r\sigma^2}{t}-\frac{4(\sigma^2+Dt)}{t^2}\ln{  \left(\frac{U_{c}\sqrt{4\pi(\sigma^2+Dt)}}{G}\right) }$$

As $t$ becomes large, this formula simplifies to $\frac{\lvert x \rvert}{t}\rightarrow \sqrt{4rD}$. Thus, the asymptotic rate of spread is $V=\sqrt{4rD}$, as we initially sought to show.

## General reaction-diffusion equation
$$\frac{\partial{u}}{\partial{t}}=
\text{D}\left(\frac{\partial{u}^2}{\partial{x^2}}\right)+f(u)$$

## Linear growth
$$\frac{\partial{u}}{\partial{t}}=
\text{D}\left(\frac{\partial{u}^2}{\partial{x^2}}\right)+ru$$

## Fisher's equation
$$\frac{\partial{u}}{\partial{t}}=
\text{D}\left(\frac{\partial{u}^2}{\partial{x^2}}\right)+ru\left(1-\frac{u}{K}\right)$$

## Harvest/fishing mortality
$$\frac{\partial{u}}{\partial{t}}=
\text{D}\left(\frac{\partial{u}^2}{\partial{x^2}}\right)+ru\left(1-\frac{u}{K}\right)-qE(x)u$$

## With advection
$$\frac{\partial{u}}{\partial{t}}=
\text{D}\left(\frac{\partial{u}^2}{\partial{x^2}}\right)-a\frac{\partial{u}}{\partial{x}}+f(u)$$

# References